### 2022-11-29 presentation.Rmd
first draft

---
title: "Transparency in Science"
subtitle: "Principled Approaches to Computational Reproducibility and Preregistration"
author: 
  - "Aaron Peikert"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
repro:
  packages:
    - xaringanthemer
    - showtext
    - svglite
    - patchwork
    - here
    - qrcode
---
class: center, middle
name: qrcode

```{r setup, include=FALSE}
library(xaringanthemer)
library(tidyverse)
library(patchwork)
library(qrcode)
options(htmltools.dir.version = FALSE)
knitr::read_chunk("R/plots.R")
knitr::opts_chunk$set(
  fig.width=9, fig.height=4, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  dev = "svglite"
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
primary_color <- "#024959"
secondary_color <- "#F28705"
style_duo(
  primary_color = primary_color,
  secondary_color = secondary_color,
  header_font_google = google_font("Didact Gothic", "400", "400i"),
  text_font_google   = google_font("IBM Plex Sans", "400", "400i"),
  text_font_size = "1.5rem",
  extra_fonts = list(google_font("Parisienne", "400", "400i"))
  )

style_extra_css(css = list(".todo" = list(color = "#17C25B"),
                           ".optional" = list(color = "#05BCE6"),
                           ".large" = list(`font-size` = "130%"),
                           ".small" = list(`font-size` = "80%"),
                           ".enormous" = list(`font-size` = "1000%"),
                           ".full" = list(padding = "0px",
                                          margin = "0px",
                                          border = "0px"),
                           ".vertical" = list(`writing-mode` = "vertical-rl",
                                              `text-orientation` = "mixed"),
                           ".blurry" = list(color = "transparent",
                                            `text-shadow` = "0 0 5px rgba(0,0,0,0.5)"),
                           ".blurry2" = list(filter = "blur(2px)",
                                             `-webkit-filter` = "blur(2px)"),
                           ".curly" = list("font-family" = "Parisienne"),
                           ".white" = list(color = "#FFFFFF")
                    )
                )
```

```{r, echo=FALSE}
link <- "github.com/aaronpeikert/bayes-prereg"
```

```{r, echo=FALSE, out.width = "30%"}
fs::dir_create("images")
generate_svg(qr_code(link), here::here("images/", "qr_slides.svg"), foreground = secondary_color, background = primary_color, show = FALSE)
knitr::include_graphics("images/qr_slides.svg")
```

`r link`

.pull-right[![CC0](https://mirrors.creativecommons.org/presskit/buttons/88x31/svg/cc-zero.svg)]

---
class: center, middle

# Transparency in Science

## Principled Approaches to Computational Reproducibility and Preregistration

### Aaron Peikert

### LCBC Berlin  | 22.11.2022

---
class: inverse, center, middle

# Trust in Science

---
class: center, middle

.enormous[?]

---

```{r, echo=FALSE}
crisis <- tribble(
  ~response, ~bin, ~prop,
  "Yes, a signicant crisis", "YES", .52,
  "Yes, a slight crisis", "YES", .38,
  "No, there is no crisis", "NO", .03,
  "Don’t know", "NA", 0.07
)

crisis_plot <- crisis %>%
  group_by(bin) %>%
  summarise(prop = sum(prop)) %>%
  ungroup() %>% 
  ggplot(aes(
    "response",
    prop,
    fill = bin,
    label = bin,
    size = prop
  )) +
  geom_col() +
  geom_text(position = position_stack(vjust = 0.5)) +
  coord_flip() +
  theme_void() +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
  ) +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_discrete(expand=c(0,0), limits = NULL) +
  scale_fill_manual(values = c("YES" = "#E3350E", "NO" = "#F28705", "NA" = "lightgrey")) +
  scale_size_continuous(range = c(4, 20)) +
  NULL
```

## Is there a replication crisis?*

.center[

```{r, out.width = '70%'}
crisis_plot /
  (crisis_plot +
     aes(label = str_c(prop * 100, "%")) + 
     scale_size_continuous(range = c(3, 12))) + 
  plot_layout(heights = c(.7, .3)) + 
  plot_annotation(theme = theme_void() + theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
))

```
]

.small.right[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]

--

.small.right[***They call it "reproducibility" and fail to publish the raw data.**]

---
class: inverse, center, middle

# Transparency in Science*

*Statisticians' Edition

---
class: center, middle

# World → Theory

---
class: center, middle

# World ← Theory

---
class: center, middle

# World ↔ Theory

--

.curly.white.large[It's a match!]

---
class: center, middle, inverse

# Principled Approach

---
class: center, middle

# $L(\text{Theory}, \text{Data})$

---
class: center, middle

# ѴαЯi€𝜏y

### $L(\text{Theory}, \text{Data})$

---
class: center, middle

$$L_1(\text{Theory}, \text{Data})$$
$$\vdots$$
$$L_\infty(\text{Theory}, \text{Data})$$

---
class: center, middle, inverse

# Statistical Models

--

= automated induction

---
class: center, middle

# Theory

---
class: center, middle

# T█e█ry

---
class: middle

$$Model(\text{Theory})$$
---
class: middle

$$Model(\text{Theory}, \text{Data})$$

---

### How statisticians judge a model:

Sampling the world and compare:

$$
L(\text{Model}(\text{Data}), \text{Data}) + \mathcal{C}(\text{Model})
$$
e.g.: Adjusted R², Stein's Unbiased Risk Estimator, Information Criteria, etc.

.center[

**We account for peeking at the data**

]

---

### Researcher have a harder job:

Sampling the world and compare:

$$
L(\text{Model}(\text{Data}), \text{Data}) + \mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})
$$
.center[

**We must account for the models and humans ability to make sense of any data.**

]

---
class: center, middle

.left[What happens if]

.center[

$\mathcal{C}(\text{Model})$ or $\mathcal{C}(\text{Human})$

]

.right[is unknown?]

---
class: center, middle, inverse

# .blurry[Uncertainty]

---

The goal:

.center[

$L(\text{Model}(\text{Data}), \text{Data}) + \mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$

]

--

The enemy:

.pull-left[

.blurry2[$$\mathcal{C}(\text{Model})$$]

]

.pull-right[

.blurry2[$$\mathcal{C}(\text{Human})$$]

]

--

The solution:

.pull-left.center[

### Computational Reproducibility

]

.pull-right.center[

### Preregistration

]

---

.pull-left.center[

### Computational Reproducibility

Conceptually trivial

]

.pull-right.center[

### Preregistration

Conceptually difficult

]

--

.pull-left.center[

technologically difficult

low social hurdles

]

.pull-right.center[

technologically trivial

high social hurdles

]

---

.pull-left.center[

### Computational Reproducibility

Conceptually trivial 😀

technologically difficult 😒

low social hurdles 😀

]

.pull-right.center[

### Preregistration

Conceptually difficult 😒

technologically trivial 😀

high social hurdles 😒

]

--

Reducing uncertainty about:

.center[

$\mathcal{C}(\text{Model})$ &emsp; vs &emsp; $\mathcal{C}(\text{Human})$

]

---
class: center, middle, inverse

# Computational Reproducibility

---
class: center, middle

“Insanity is doing the same thing over and over again and expecting different results.”

.right[—Albert Einstein (disputed)]

--

As it turns out, doing the **same** thing is pretty complicated.

---
class: middle

1. Multiple inconsistent versions of code and data
--

2. Copy-and-paste errors
--

3. Ambiguous order of code execution
--

4. Broken dependencies

---
class: center, middle

| Problem       | Solution            | Implementation    |
|---------------|---------------------|-------------------|
| versions?     | version control     | Git               |
| copy&paste?   | dynamic documents   | R Markdown        |
| order?        | workflow automation | Make / GH Actions |
| dependencies? | software management | Docker            |

---
class: center, middle
background-color: white

```{r, out.width='80%'}
knitr::include_graphics("https://raw.githubusercontent.com/aaronpeikert/repro-tutorial/main/images/nutshell.svg")
```

---
class: center, middle

.small[<https://github.com/aaronpeikert/repro-talk/releases/download/marbach/repro-talk.pdf>]

---
class: center, middle, inverse

# Preregistration

---

## Is there a replication crisis?*

.center[

```{r, out.width = '70%'}
crisis_plot /
  (crisis_plot +
     aes(label = str_c(prop * 100, "%")) + 
     scale_size_continuous(range = c(3, 12))) + 
  plot_layout(heights = c(.7, .3)) + 
  plot_annotation(theme = theme_void() + theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
))
```
]

.small.right[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]

---
class: center, middle
layout: true

---

## Because researchers are *fraudsters*?

--

# No

???
Some say to prevent intentional wrongdoing. 
However, preregistrations are easily circumvented with bad intentions.
We must assume researchers have good intentions.
---

## Because researchers are *fools*?

--

# No
--
, not necessarily

???
Some say to prevent unintentional wrongdoing.
Researchers must be put in restrains to prevent them from cutting themselves on the sharp edge of "false positives".

---
layout: false
class: inverse, center, middle

# Traditional view

---
class: center

# Researchers confuse:

.pull-left.center[

# confirmatory

]
.pull-right.center[

# exploratory

]
---
class: center

# Preregistration separates:

.pull-left.center[

# confirmatory

]
.pull-right.center[

# exploratory

]
--
.pull-left.center[

# ≈
# preregistered

]
.pull-right.center[

# ≈
# not preregistered

]

--

# ?

---
class: inverse, center, middle

# A Formal Objective

---
class: middle

.large[
Ideally:
1. The original author has **good intentions**.
2. Their readers are **rational**.
3. The goal is to **convince** the readers.
]

---

### Adjusting for complexity in simple models

Adjusted R², Stein's Unbiased Risk Estimator, Information Criteria, etc.

$$L(\text{Model}(\text{Data}), \text{Data})  + \mathcal{C}(\text{Model})$$

--

### Alternatively:

The statistical complexity of some models is hard to estimate.

Find a situation, were $\mathcal{C}(\text{Model}) = 0$

---
## Test-set (cross validation)

$\mathcal{C}(\text{Model}) = 0$

$$
L(\text{Model}(\text{Train, Test}), \text{Test})
$$
.center[

**There is no peeking to account for.**

]

---
## $\mathcal{C}(\text{Human})$

The statistical complexity of humans is hard to estimate.

Find a situation, were $\mathcal{C}(\text{Human}) = 0$

--

.center[

# Preregistration

]

---
class: center, middle, inverse

# .blurry[Uncertainty]

---
# Statistical Relevancy

Any:

$$L(\text{Theory}, \text{Data})$$

Must benefit from:

$$P(E|H) \gg P(E|¬H)$$

---
class: center, middle, inverse

# Bayes*

*my favoritey $L(\text{Theory}, \text{Data})$

---
class: middle

.large[

$$\frac{P(H)P(E|H)}{P(E)}$$
]

--

.left[Law of total probability:]

$$P(E) = P(H)P(E|H) + P(¬H)\color{#F28705}{P(E|¬H)}$$

---
class: center, middle

## $1 - P(E|¬H)$
## =
## $P(¬E|¬H)$
## =
## Theorethical Risk

---
class: center, middle

## $\mathcal{C}(\text{Model})$ → Theorethical Risk ← $\mathcal{C}(\text{Human})$

---
class: center, middle, inverse

# .blurry[Uncertainty]

---
class: center, middle
name: theoretical-risk-plot2

```{r}
results <- readr::read_rds(here::here("data", "simulation.rds")) 
```

```{r measure-plots}
# see R/plots.R
```

```{r}
plots[["Posterior Probability"]] +
  xaringanthemer::theme_xaringan() +
    theme(legend.position = c(.2, .8),
        legend.title = element_blank(),
        text = element_text(size = 18, color = secondary_color),
        title = element_blank(),
        axis.title = element_text(size = 18),
        plot.caption = element_text(size = 15)) +
  labs(x = "Theoretical Risk",
       y = "Posterior Probability",
       caption = "P(E|H) = .8, P(H) = .1") +
  NULL
```

---
class: center, middle
layout: true
---

.large[**Preregistrations**] reduce .large[**uncertainty**] about .large[**theoretical risk**].

---

### One further idea

---

# Preregistration as Code

### version control + dynamic documents

---
class: center

.pull-left[

### Standard Preregistration

hunches

↓

preregistration

↓

data

↓

article draft
]

.pull-right[

### Preregistration as Code

simulated data

↓

article draft with mock results

↓

data

↓

article draft with real results

]

---

# Can I explore the data?

--

# Yes.

---
template: theoretical-risk-plot2

---

# Can I deviate?

--

# Sometimes.

---
template: theoretical-risk-plot2

---

# How much detail?

--

# Very detailed.

---
template: theoretical-risk-plot2

---
template: qrcode

### 2022-12-23 manuscript.Rmd
add draft of introduction

---
title: "Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration"
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
    number_sections: false
header-includes:
   - \usepackage{libertine}
fontsize: 12pt
geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
papersize: a4
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
csl: apa.csl
repro:
  files:
    - abstract.md
    - apa.csl
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
---

\newpage

# Introduction

<!--psychology has unique difficulties-->
Psychology is a difficult science [@meehlTheoreticalRisksTabular1978].
Though there might be some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
Laying open the difficulties and attempting to overcome them is not a recent trend, though it has been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
The debates about psychology's shortcomings and remedies fall roughly into two categories: statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and sociological factors [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].

<!--bla-transparency-bla-->
I propose to view both categories through the lens of transparency about the inductive process.
One side of transparency is that statistical methods make induction quantifiable; the other is that open science measures remove uncertainty about sociological factors in induction.
Any science must be able to communicate how it generates its knowledge, but transparency has an outstanding role in psychology and other empirical sciences.
To understand why transparency is crucial, we must understand that induction is integral to empirical sciences, formally using statistical procedures or informally as researchers' judgment.
However, empirical statements lose their value without transparency about the inductive process that contributed to them.
Therefore, transparency is more than a virtue that somewhat improves psychological sciences but is an indispensable property.
Without transparency, psychological sciences would be futile.

<!--induction is necessary for psychological science-->
To function as a science, psychology must be able to make statements about the world that can be compared to the true conditions.
In psychology, this is not a purely deductive endeavor.
Very few psychological theories are precise enough to derive testable statements.
While it is tempting to claim that a theory makes testable claims by, e.g., implying a mean difference between two groups, such a statement is not testable on its own.
Consider a simple t-test of mean differences.
The decision to reject the null hypothesis still depends, besides true mean difference and sample size, on the observed variance.
However, the variance is, of course, estimated from the data.
So the threshold of the deductive decision depends on a quantity that must be induced.

<!--sometimes the use of induction is well justified-->
Induction is necessary and, in the present case, innocuous for psychology as a science.
Here, it is innocuous because the introduced bias can be accounted for and vanishes with increasing sample size, but we will see later that this is not always the case.
In this specific example, we use a t-test instead of a z-test to account for the estimation of variance from data, and it is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
Induction is also necessary because it is almost unimaginable to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
If they had to, there would probably be no psychological theory that could survive an empirical test.

<!--necessity of induction-->
In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled with data.
It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions so that they fit the data well.
In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
By inducing some quantities, psychological researchers can remove them from the conjunction.
If researchers use induction for some necessary but under the theory arbitrary assumptions, their theory will not be refuted because of these assumptions.
Since empirical researcher often cannot derive every assumption from their theory, avoiding refutation because of those assumptions is a desirable property.

<!--statistical induction can be dangerous-->
By the same token, whole theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
Such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
If pushed to the limit, we arrive at a "theory" that is governed by the data.
Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
The data used for induction, can not refute these statements, so they are not tested at all.

<!--the appraisal of "induced theories" should be low-->
So what to think of such, yet untested, theory?
Researchers (and philosophers of science) differ considerably in how they think theories should be appraised, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p. 14] of a hypothesis.
But whatever measure they subscribe to, their appraisal of the empirical support of a yet untested theory would be relatively low.

<!--transparency about induction is required-->
So empirical researchers find themselves in a pinch.
On the one hand, they need induction to test their imprecise theories.
On the other hand, induction may render any test of a hypothesis ineffective.
The problem, I argue, therefore, is not induction but making induction transparent.
The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims are.
The question of this dissertation is, therefore, what must be made transparent, and how can we make it transparent?

<!--forshadow the main part-->
The first question (the "what") is theoretical in nature.
It is addressed in the thesis itself, which supplies the theoretical framework of the articles written as part of the dissertation.
Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).

<!--how it relates to the articles-->
Based on this split, the articles answer the second question (the "how").
I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these tools in practice.

# An information theorethic perspective on transparency in science

The need for transparency is closely tied to the use of induction in the empirical test of a theory.
There has been a long and vigorous debate about what it means to test a theory empirically.
One framework that is especially suited to investigate how induction plays into such empirical test is information theory.
Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude ("truth likeness") of a theory.
That is, how much information about the truth is lost when the theory is used to model reality.
Expressed mathematically we have a function $f(x)$ that gives us the probability to observe the state of the world $x$ where $f$ represents the full reality.
We are now interested, how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $x$.
Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:

$$
I(f, g) = \mathbb{E}_x\left[log(f(x))\right] - \mathbb{E}_x\left[log(g(x))\right]
$$

Most readers will recognize, that this information theoretic setup and the derivation below follow closely XXX in their derivation of the Akaike Information Criterion.
What is of interest here, is not the derivation, but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.

This setup is of course highly theoretical.
$I(f, g)$ is unknowable, since the truth is unobserved.
This fact, however, does not impede us from getting closer to the truth because we still can compare two different theories relative to each other.
Because the expectation for $f$ remains constant (left hand side) we only need to estimate the relative expected loss of information (right hand side) to make a comparative judgment.
To make a relative judgment about several competing theories it suffices to estimate:

$$
\mathbb{E}_x\left[log(g(x))\right]
$$

To allow for quantities to be induces, we must assume that our theory is parametrized, e.g., $g(x|\theta)$.
That means our theory implies a multitude of possible probability distribution that may describe reality.
Though, this assumption may be lifted, we assume each implied version of the theory equally probable a priori.
This parametrization captures the idea that some assumptions necessary in a theory to make predictions are arbitrary.
Of those arbitrary assumptions we want to find those that fit the reality with the least amount of information lost.
The best parametrization is archived by:

$$
\theta_o = \operatorname*{arg\,min}_\theta I(f, g(\cdot|\theta))
$$

The inference goal is therefore:

$$
\mathbb{E}_x\left[log(g(x|\theta_o))\right]
$$

Of course we usually do not know $\theta_o$.
That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$.
The crucial point here is to understand what happens, when we can not derive $\theta$, deductively, but must substitute it inductively with an estimate $\hat\theta(y)$.
Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_o$.
It follows that, almost surely:

$$
I(f, g(\cdot|\hat{\theta}(y)) > I(f, g(\cdot|\theta_0))
$$

Or ignoring $f$ as constant:

$$
\mathbb{E}_x[log(g(x|\hat{\theta}(y)))] > \mathbb{E}_x\left[log(g(x|\theta_o))\right] 
$$

That is to say, any induced estimate will be sub optimal.
The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased we may form an expectation over the data $y$:

$$
\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]
$$

Forming this expectation over data is a crucial step, it requires us to think beyond the data we observed to all the data we could have observed.
There are two ways to get at this expectation.
One follows in this section, another is discussed in the next section.

We might notice that we often get an unbiased estimate for $\mathbb{E}_x[log(g(x|\hat{\theta}(y)))]$, e.g., in maximum likelihood estimation.
The expectation over the data together with Taylor series expansion yields:

$$
\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_x[log(g(x|\hat\theta(y)))] - tr[J(\theta_o)I(\theta_o)^{-1}]
$$

The observed likelihood $log(g(x|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
So substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.

Fortunately, this bias $tr[J(\theta_o)I(\theta_o)^{-1}]$ can be approximated under some conditions.
Since $\theta_o$ is unknown, any such estimation process relies on properties of the inductive process that generated $\hat\theta$.
That is, it is insufficient to know the end result of the induction, we must understand the inductive process that generated those estimates.

To summaries, what this means for researchers evaluating theories.
A researcher who empirically evaluates a theory without inductive elements gets a good estimate how close they are to the truth (relative to other theories evaluated on the same data).
If we believe they did not make any mistakes and that they are truthful, we may take their assessment at face value.
We might still want transparency on how they did gather the data and how they compared it to the theories.
However, it is a simple kind of transparency, about what did happen.
If the researcher did induce some quantities, transparency about what did happen is not enough; instead a transparency about what could have happened, had the data looked different, is required.
This information theoretic perspective one perspective that warrants transparency about the process of induction or equivalently about what could have happened.
What it means to establish this kind of transparency (the "how") will be discussed at length, however, before we continue, let me motivate the need for transparency about the process (the "what") using another perspective.

# A future performance perspective on transparency

There is another line of argumentation why transparency about the process of induction is important besides closeness to truth.
Instead of verisimilitude one might be concerned with future performance, hence, how well does a theory do in predicting novel facts.
These very different motivations can be linked, but both warrant transparency on their own.
If predictions and observed facts are compared using a sufficient statistic a close link to information theory exists.
The expectation over data $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$, can be estimated by repeatedly sampling data and repeating the inductive process.
This insight connects the information theoretic setup with the process of cross validation, where the data is partitioned and the inductive process is repeated on each partition.

# Transparency about statistical models: Computational Reproducibilibiltiy

# Transparency about human researcher: Preregistration

# Discussion

## Limitations

## Future Research

Such use induction to preemptively escape refutation differs in at least one important way from altering the hypothesis after refutation.
Induction through statistical models can be formally analyzed and accounted for.
Therefore, the researchers may adjust their appraisal of the partially induced hypothesis.
Adjusting the appraisal for induction outside of statistical methods is much more uncertain.
Here the process of induction can not be separated from the data itself, since we observe such induction only on a single dataset.
Statistical models on the other hands, have well defined inductive behavior that can be analyzed over all possible datasets.

<!--you must pay the reaper-->
Accounting for how much induction was used to deductively test a statement is necessary for correct appraisal of evidential support of a hypothesis.
Researcher (and philosophers of science) differ considerably in what appraisal entails, e.g., judging the long-term performance (if they are frequentists), degrees of believe (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p. 14] of a hypothesis.
But whatever measure they subscribe to, induction can shield shield any hypothesis from refutation by data altogether.

Assume that all auxiliary assumption, regarding proper quantification of behavior and situation in valid measurements, causality, individual differences, nuisance variables, distributional, etc, hold.

While not using statistical methods would lead to the rejection of almost all psychological theories, pushed to the extreme statistical methods can shield theories from refutation altogether.
Both situations are uninformative and, therefore, rather unhelpful for the generation of knowledge.
The later situation, however, becomes less than helpful when one is under the impression that theories are well tested, when in fact there was little chance that the data could refute the theory if wrong.

# References

### 2022-12-23 manuscript.Rmd
split into subheadings

@@
 
-# An information theorethic perspective on transparency in science
+# Two perspectives that necessitate transparency
 
@@
 There has been a long and vigorous debate about what it means to test a theory empirically.
+Two frameworks lend themselves to investigate the effect of induction.
+The one framework motivates transparency under a sciences that attempts to empirically evaluate the verisimilitude ("truth likeness") of a theory.
+The other framework motivates transparency under a sciences that wants to select a theory according to its future predictive performance.
+
+## An information theorethic perspective
+
 One framework that is especially suited to investigate how induction plays into such empirical test is information theory.
-Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude ("truth likeness") of a theory.
+Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory.
 That is, how much information about the truth is lost when the theory is used to model reality.
@@
 
-Most readers will recognize, that this information theoretic setup and the derivation below follow closely XXX in their derivation of the Akaike Information Criterion.
+Most readers will recognize, that this information theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002 in their derivation of the Akaike Information Criterion in a very general form.
 What is of interest here, is not the derivation, but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
@@
 However, it is a simple kind of transparency, about what did happen.
-If the researcher did induce some quantities, transparency about what did happen is not enough; instead a transparency about what could have happened, had the data looked different, is required.
-This information theoretic perspective one perspective that warrants transparency about the process of induction or equivalently about what could have happened.
-What it means to establish this kind of transparency (the "how") will be discussed at length, however, before we continue, let me motivate the need for transparency about the process (the "what") using another perspective.
+If the researcher induces some quantities, transparency about what did happen is not enough; instead a transparency about what could have happened, had the data looked different, is required.
 
@@
 
-There is another line of argumentation why transparency about the process of induction is important besides closeness to truth.
+There is another line of argumentation about why transparency about the process of induction is important besides closeness to truth.
 Instead of verisimilitude one might be concerned with future performance, hence, how well does a theory do in predicting novel facts.
@@
 The expectation over data $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$, can be estimated by repeatedly sampling data and repeating the inductive process.
-This insight connects the information theoretic setup with the process of cross validation, where the data is partitioned and the inductive process is repeated on each partition.
+This insight connects the information theoretic setup with cross validation, where the data is partitioned and the inductive process is repeated on each partition.
 

### 2022-12-29 manuscript.Rmd
reorder headings
expand

@@
     latex_engine: "xelatex"
-    toc: false
+    toc: true
     number_sections: false
@@
 Therefore, transparency is more than a virtue that somewhat improves psychological sciences but is an indispensable property.
-Without transparency, psychological sciences would be futile.
 
@@
 
-# Two perspectives that necessitate transparency
+# What necessitates transparency
 
@@
 There has been a long and vigorous debate about what it means to test a theory empirically.
-Two frameworks lend themselves to investigate the effect of induction.
-The one framework motivates transparency under a sciences that attempts to empirically evaluate the verisimilitude ("truth likeness") of a theory.
-The other framework motivates transparency under a sciences that wants to select a theory according to its future predictive performance.
+I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
+The first framework motivates transparency when the aim of an empirical test is to evaluate the verisimilitude ("truth likeness") of a theory.
+The second framework motivates transparency under a sciences that wants to select a theory according to its expected predictive performance.
 
-## An information theorethic perspective
+## An information theoretic perspective
 
-One framework that is especially suited to investigate how induction plays into such empirical test is information theory.
 Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory.
@@
 $$
-I(f, g) = \mathbb{E}_x\left[log(f(x))\right] - \mathbb{E}_x\left[log(g(x))\right]
+\mathcal{I}(f, g) = \mathbb{E}_x\left[log(f(x))\right] - \mathbb{E}_x\left[log(g(x))\right]
 $$
 
-Most readers will recognize, that this information theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002 in their derivation of the Akaike Information Criterion in a very general form.
+Most readers will recognize, that this information theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
+7.2, in their derivation of the Akaike Information Criterion in its general form.
 What is of interest here, is not the derivation, but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
@@
 This setup is of course highly theoretical.
-$I(f, g)$ is unknowable, since the truth is unobserved.
+$\mathcal{I}(f, g)$ is unknowable, since the truth is unobserved.
 This fact, however, does not impede us from getting closer to the truth because we still can compare two different theories relative to each other.
@@
 
-To allow for quantities to be induces, we must assume that our theory is parametrized, e.g., $g(x|\theta)$.
+To allow for quantities to be induced, we must assume that our theory is parametrized, e.g., $g(x|\theta)$.
 That means our theory implies a multitude of possible probability distribution that may describe reality.
 Though, this assumption may be lifted, we assume each implied version of the theory equally probable a priori.
-This parametrization captures the idea that some assumptions necessary in a theory to make predictions are arbitrary.
+This parametrization captures the idea that some assumptions necessary in a theory to make testable statements are arbitrary.
 Of those arbitrary assumptions we want to find those that fit the reality with the least amount of information lost.
@@
 $$
-\theta_o = \operatorname*{arg\,min}_\theta I(f, g(\cdot|\theta))
+\theta_o = \operatorname*{arg\,min}_\theta \mathcal{I}(f, g(\cdot|\theta))
 $$
@@
 
-Of course we usually do not know $\theta_o$.
+Of course, we usually do not know $\theta_o$.
 That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$.
@@
 $$
-I(f, g(\cdot|\hat{\theta}(y)) > I(f, g(\cdot|\theta_0))
+\mathcal{I}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{I}(f, g(\cdot|\theta_o))
 $$
@@
 
-That is to say, any induced estimate will be sub optimal.
+That is to say, any induced estimate will be suboptimal.
 The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
@@
 
+If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
 Fortunately, this bias $tr[J(\theta_o)I(\theta_o)^{-1}]$ can be approximated under some conditions.
-Since $\theta_o$ is unknown, any such estimation process relies on properties of the inductive process that generated $\hat\theta$.
+One condition is, since $\theta_o$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
 That is, it is insufficient to know the end result of the induction, we must understand the inductive process that generated those estimates.
+That requires that the inductive process can be formally described, e.g., corrections for a large class of statistical models, most famously the class of linear models, are available.
 
-To summaries, what this means for researchers evaluating theories.
+To summarize, what does this mean for researchers evaluating theories?
 A researcher who empirically evaluates a theory without inductive elements gets a good estimate how close they are to the truth (relative to other theories evaluated on the same data).
@@
 
-# A future performance perspective on transparency
+## A future performance perspective
 
-There is another line of argumentation about why transparency about the process of induction is important besides closeness to truth.
-Instead of verisimilitude one might be concerned with future performance, hence, how well does a theory do in predicting novel facts.
-These very different motivations can be linked, but both warrant transparency on their own.
-If predictions and observed facts are compared using a sufficient statistic a close link to information theory exists.
+In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
+Instead of verisimilitude one might be concerned with future performance. That is, how well does a theory do in predicting novel facts.
+It is important to realize that the information theoretic setup above has made no appeal to expected performance on unseen data.
+Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
+If predictions and observed facts are compared using a sufficient statistic, a close link to information theory exists.
 The expectation over data $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$, can be estimated by repeatedly sampling data and repeating the inductive process.
-This insight connects the information theoretic setup with cross validation, where the data is partitioned and the inductive process is repeated on each partition.
+This insight connects the information theoretic setup with cross validation, where the data is partitioned and the inductive process is repeated on all permutations of a subset of the partitions [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+For each subset, the resulting model is then compared to the complement that was not used for induction.
+Of course, there is no strong reason to use a sufficient statistic for this comparison beyond its link to information theory.
+So researcher might differ in what they find to be a good prediction (replacing the Kullback–Leibler divergence with an arbitrary loss function) and we do not longer appeal to a concept of ground truth:
 
-# Transparency about statistical models: Computational Reproducibilibiltiy
+$$
+\mathcal{L}(x, g(x|\theta))
+$$
 
-# Transparency about human researcher: Preregistration
+Again there exists some optimal $\theta_o$:
+
+$$
+\theta_{o} = \operatorname*{arg\,min}_\theta \mathcal{L}(x, g(\cdot|\theta))
+$$
+
+Choosing $\hat\theta(y)$ based on a
+
+## A conceptual perspective
+
+Now that we have established the need for transparency about the inductive process, we can drop a few of the more technical details to get a more straightforward answer about the how.
+It bears repeating that simply laying open what has been done is not enough.
+Showing the inductive results instead of the process is insufficient to appraise the theory.
+On a conceptual level we want to compare:
+
+$$
+\mathcal{L}(\text{Theory}, \text{Reality})
+$$
+
+But to allow for induction to happen we replace theory by a model (not necessarily a statistical one), or put differently a multitude of implications about the data from the theory.
+This multitude is a result of arbitrary assumptions.
+
+$$
+\mathcal{L}(\text{Model(Reality)}, \text{Reality})
+$$
+
+The idea is that we choose from the multitude the version of our theory that best fits reality.
+However, we are forced to rely on a limited sample of reality.
+We are mislead because these two factors, induction and limited sample size, interact.
+There are two principled approaches for this problem.
+Choosing the version of our theory according to the best fit on one sample and use another sample to asses it:
+
+*Note, this is not strictly correct. No expectation over the inferred parameters is taken.*
+
+$$
+\mathcal{L}(\text{Model}(\text{Train}), \text{Test})
+$$
+
+Or use the same sample for both, but correct for the process of induction:
+
+$$
+\mathcal{L}(\text{Model}(\text{Train}), \text{Train}) + \mathcal{C}(\text{Model})
+$$
+
+Where $\mathcal{C}$ denotes complexity or equivalently the extend to which data is influencing the results.
+To deal with this bias, I propose to split the term in complexity that can be formally described and a part that can only evaluated subjectively.
+Formal, in the strict sense means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
+If a researcher employs a linear model, the complexity is calculable and even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
+However, if the researchers reconsiders the model based on the results, and fits a model with, e.g., a predictor added or removed, the complexity can not be formally judged.
+We can not ask the researcher to repeat the process on all possible data sets nor is the behavior mathematically well defined.
+What we know, however, is that the resulting linear model is more complex than a linear regression formally implies.
+
+$$
+\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
+$$
+
+That is not to say that we have no basis to judge the informal induction.
+We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
+Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data driven decision, which implies higher complexity.
+What one can say, however, is that this judgment is debatable and, therefore, subjective.
+
+# How to establish transparency
+
+Using a formally defined inductive process to preemptively escape refutation differs in at least one important way from altering the hypothesis after refutation.
+Induction through statistical models can be formally analyzed and accounted for.
+Therefore, the researchers may adjust their appraisal of the partially induced hypothesis.
+Adjusting the appraisal for induction outside of statistical methods is much more uncertain.
+Here the process of induction can not be separated from the data itself, since we observe such induction only on a single dataset.
+To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
+Otherwise, the supposedly objective test of the theory using hard data must still be judged subjectively.
+
+If this imperative is adopted, it still remains to make this process transparent to other researchers.
+Not an easy feat as we will see.
+We, therefore, start with the part that is easier made transparent.
+The complexity due to the formally defined inductive process may be made transparent by employing computational reproducibility.
+For the more informally made inductive decisions, we have a more limited toolbox that includes preregistration.
+
+## Transparency about statistical models: Computational Reproducibilibiltiy
+
+Even if the results stand at the end of a formal inductive process, this process must be made transparent to other researchers if they are to judge the empirical test.
+Such transparency can be achieved by computational reproducibility.
+Computational reproducibility is usually defined as the ability to recreate the same results from the same data set.
+This definition only suffices if the inductive process is strictly formal, that is the complexity can be calculated mathematically.
+If we must rely on repetition to assess the complexity, we must demand computational reproducibility to extent to other data sets.
+
+## Transparency about human researcher: Preregistration
 
@@
 
+### Theorethical
+
+### Practical
+
 ## Future Research
-
-Such use induction to preemptively escape refutation differs in at least one important way from altering the hypothesis after refutation.
-Induction through statistical models can be formally analyzed and accounted for.
-Therefore, the researchers may adjust their appraisal of the partially induced hypothesis.
-Adjusting the appraisal for induction outside of statistical methods is much more uncertain.
-Here the process of induction can not be separated from the data itself, since we observe such induction only on a single dataset.
-Statistical models on the other hands, have well defined inductive behavior that can be analyzed over all possible datasets.
 

### 2022-12-30 manuscript.Rmd
extent formalism for cv

@@
 
-That is to say, any induced estimate will be suboptimal.
+That is to say, any induced estimate will be sub optimal.
 The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
@@
 In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
-Instead of verisimilitude one might be concerned with future performance. That is, how well does a theory do in predicting novel facts.
+Instead of verisimilitude one might be concerned with future performance.
+That is, how well does a theory do in predicting novel facts.
+We dispense in this argument with an appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return probabilities:
+
+$$
+\mathcal{L}(y, g(y|\theta)) = \mathbb{E}_yL(y, g(y|\theta))
+$$
+
+Still the core problems of limited sample and inductive process remain.
+Again there exists only a sample $y$ over which we form the expectation, but now we must pay a little more attention to the individual elements, $i = 1, \dots,n$:
+
+<!--the unconscious statistician strikes again-->
+
+$$
+\mathbb{E}_y L(y, g(y|\theta))
+ = \sum_{i=1}^n L(y_i, g(y_i|\theta)) \frac{1}{n}
+$$
+
+Since, we allow arbitrary $L$, the trick of using the Taylor series expansion does not get us very far.
+What we can do, however, is to make use of the cross-validatory, which is indicated by $y_{-i} = y \diagdown \{y_i\}$
+
+$$
+\mathbb{E}_y L(y, g(y|\hat\theta(y)))
+ = \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))\frac{1}{n}
+$$
+
 It is important to realize that the information theoretic setup above has made no appeal to expected performance on unseen data.
@@
 
-$$
-\mathcal{L}(x, g(x|\theta))
-$$
-
-Again there exists some optimal $\theta_o$:
-
-$$
-\theta_{o} = \operatorname*{arg\,min}_\theta \mathcal{L}(x, g(\cdot|\theta))
-$$
-
-Choosing $\hat\theta(y)$ based on a
+Which we can estimate using cross-validation.
 

### 2023-01-04 manuscript.Rmd
begin section on the how

@@
 ---
-title: "Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration"
+title: "Towards Transparency and Open Science"
+subtitle: "A Principled Perspective on Computational Reproducibility and Preregistration"
 output:
@@
 I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these tools in practice.
+The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these tools.
 
@@
 Though, this assumption may be lifted, we assume each implied version of the theory equally probable a priori.
-This parametrization captures the idea that some assumptions necessary in a theory to make testable statements are arbitrary.
+This parametrization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
 Of those arbitrary assumptions we want to find those that fit the reality with the least amount of information lost.
@@
 The observed likelihood $log(g(x|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-So substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
+Note that $tr[J(\theta_o)I(\theta_o)^{-1}]$ is usually negative definitive and, therefore, we may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
+I will refer to this bias as the complexity of a model, denoted as $\mathcal{C}$, because it is closely related to how much the data is influencing the results and, hence, how detailed the model is representing the data.
+
+$$
+\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_x[log(g(x|\hat\theta(y)))] + \mathcal{C}
+$$
 
 If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
-Fortunately, this bias $tr[J(\theta_o)I(\theta_o)^{-1}]$ can be approximated under some conditions.
+Fortunately, it is possible to approximate the complexity of a model under some conditions.
 One condition is, since $\theta_o$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-That is, it is insufficient to know the end result of the induction, we must understand the inductive process that generated those estimates.
-That requires that the inductive process can be formally described, e.g., corrections for a large class of statistical models, most famously the class of linear models, are available.
-
-To summarize, what does this mean for researchers evaluating theories?
-A researcher who empirically evaluates a theory without inductive elements gets a good estimate how close they are to the truth (relative to other theories evaluated on the same data).
-If we believe they did not make any mistakes and that they are truthful, we may take their assessment at face value.
-We might still want transparency on how they did gather the data and how they compared it to the theories.
-However, it is a simple kind of transparency, about what did happen.
-If the researcher induces some quantities, transparency about what did happen is not enough; instead a transparency about what could have happened, had the data looked different, is required.
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are available.
 
@@
 That is, how well does a theory do in predicting novel facts.
-We dispense in this argument with an appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return probabilities:
+Please note that the information theoretic setup above has made no appeal to expected performance on unseen data.
+Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
+In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return probabilities:
 
@@
 
-Since, we allow arbitrary $L$, the trick of using the Taylor series expansion does not get us very far.
-What we can do, however, is to make use of the cross-validatory, which is indicated by $y_{-i} = y \diagdown \{y_i\}$
+Since, we allow arbitrary $L$, the trick of using the Taylor series expansion does not get us very far (because we simply end up with an undefined derivative).
+We can, however, repeatedly sample data and repeate the inductive process.
+That is, we make use of the cross-validatory where the data is partitioned and the inductive process is repeated on all permutations of a subset of the partitions.
+For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
 
@@
 
-It is important to realize that the information theoretic setup above has made no appeal to expected performance on unseen data.
-Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-If predictions and observed facts are compared using a sufficient statistic, a close link to information theory exists.
-The expectation over data $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$, can be estimated by repeatedly sampling data and repeating the inductive process.
-This insight connects the information theoretic setup with cross validation, where the data is partitioned and the inductive process is repeated on all permutations of a subset of the partitions [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
-For each subset, the resulting model is then compared to the complement that was not used for induction.
-Of course, there is no strong reason to use a sufficient statistic for this comparison beyond its link to information theory.
-So researcher might differ in what they find to be a good prediction (replacing the Kullback–Leibler divergence with an arbitrary loss function) and we do not longer appeal to a concept of ground truth:
+If we reinstate Kullback-Leibler divergence as a loss function, we can estimate $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$ using the cross-validatory which connects the information theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+To make the link to the first approach even more clear we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
 
-Which we can estimate using cross-validation.
+$$
+\mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta))
+$$
+
+Instead of a formal analyses to derive $\mathcal{C}$, we can simple repeat the inductive process, i.e., using cross-validation.
+That drastically expands the set of inductive processes for which we can estimate the inductive bias, since it is considerably easier to repeat the process than to derive the complexity analytically.
 
@@
 We are mislead because these two factors, induction and limited sample size, interact.
-There are two principled approaches for this problem.
-Choosing the version of our theory according to the best fit on one sample and use another sample to asses it:
-
-*Note, this is not strictly correct. No expectation over the inferred parameters is taken.*
 
 $$
-\mathcal{L}(\text{Model}(\text{Train}), \text{Test})
+\mathcal{L}(\text{Model(Reality)}, \text{Reality}) > \mathcal{L}(\text{Model(Sample)}, \text{Sample})
 $$
 
-Or use the same sample for both, but correct for the process of induction:
+$$
+\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}
+$$
 
-$$
-\mathcal{L}(\text{Model}(\text{Train}), \text{Train}) + \mathcal{C}(\text{Model})
-$$
+Transparency is necessary because induction leaves researchers overly optimistic regarding their theories fit to the data.
+The extend of this optimism is dependent on the process of induction and, hence, researcher ought to make it transparent.
+
+# How to establish transparency
+
+The above sections aimed to motivate the observation that the apparent fit of a theory to data is almost always overly optimistic.
+This observations is only of limited use, if we do not now the extend of this optimism.
+However, both setups show that the extend of the optimism ($\mathcal{C}$) is closely related to the inductive process and give us two starting points for making this bias transparent.
+The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), the second merely requires that the process is repeatable (e.g., using cross-validation).
+Both approaches require researchers to make the process of induction transparent instead of merely publishing the results.
+
+Using a formally defined inductive process to preemptively escape refutation differs in at least one important way from altering the hypothesis after refutation.
+Induction through statistical models can be formally analyzed and accounted for.
+Therefore, the researchers may adjust their appraisal of the partially induced hypothesis.
+Adjusting the appraisal for induction outside of statistical methods is much more uncertain.
+Here the process of induction can not be separated from the data itself, since we observe such induction only on a single dataset.
+To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
+Otherwise, the supposedly objective test of the theory using hard data must still be judged subjectively.
 
@@
 What one can say, however, is that this judgment is debatable and, therefore, subjective.
-
-# How to establish transparency
-
-Using a formally defined inductive process to preemptively escape refutation differs in at least one important way from altering the hypothesis after refutation.
-Induction through statistical models can be formally analyzed and accounted for.
-Therefore, the researchers may adjust their appraisal of the partially induced hypothesis.
-Adjusting the appraisal for induction outside of statistical methods is much more uncertain.
-Here the process of induction can not be separated from the data itself, since we observe such induction only on a single dataset.
-To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
-Otherwise, the supposedly objective test of the theory using hard data must still be judged subjectively.
 

### 2023-01-04 manuscript.Rmd
remove leftovers

@@
 
-Using a formally defined inductive process to preemptively escape refutation differs in at least one important way from altering the hypothesis after refutation.
-Induction through statistical models can be formally analyzed and accounted for.
-Therefore, the researchers may adjust their appraisal of the partially induced hypothesis.
-Adjusting the appraisal for induction outside of statistical methods is much more uncertain.
-Here the process of induction can not be separated from the data itself, since we observe such induction only on a single dataset.
 To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
@@
 
-Where $\mathcal{C}$ denotes complexity or equivalently the extend to which data is influencing the results.
-To deal with this bias, I propose to split the term in complexity that can be formally described and a part that can only evaluated subjectively.
-Formal, in the strict sense means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
-If a researcher employs a linear model, the complexity is calculable and even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
-However, if the researchers reconsiders the model based on the results, and fits a model with, e.g., a predictor added or removed, the complexity can not be formally judged.
-We can not ask the researcher to repeat the process on all possible data sets nor is the behavior mathematically well defined.
-What we know, however, is that the resulting linear model is more complex than a linear regression formally implies.
-
-$$
-\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
-$$
-
-That is not to say that we have no basis to judge the informal induction.
-We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
-Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data driven decision, which implies higher complexity.
-What one can say, however, is that this judgment is debatable and, therefore, subjective.
-
-If this imperative is adopted, it still remains to make this process transparent to other researchers.
-Not an easy feat as we will see.
-We, therefore, start with the part that is easier made transparent.
-The complexity due to the formally defined inductive process may be made transparent by employing computational reproducibility.
-For the more informally made inductive decisions, we have a more limited toolbox that includes preregistration.
-
 ## Transparency about statistical models: Computational Reproducibilibiltiy
-
-Even if the results stand at the end of a formal inductive process, this process must be made transparent to other researchers if they are to judge the empirical test.
-Such transparency can be achieved by computational reproducibility.
-Computational reproducibility is usually defined as the ability to recreate the same results from the same data set.
-This definition only suffices if the inductive process is strictly formal, that is the complexity can be calculated mathematically.
-If we must rely on repetition to assess the complexity, we must demand computational reproducibility to extent to other data sets.
 
@@
 
-<!--you must pay the reaper-->
-Accounting for how much induction was used to deductively test a statement is necessary for correct appraisal of evidential support of a hypothesis.
-Researcher (and philosophers of science) differ considerably in what appraisal entails, e.g., judging the long-term performance (if they are frequentists), degrees of believe (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
-14] of a hypothesis.
-But whatever measure they subscribe to, induction can shield shield any hypothesis from refutation by data altogether.
-
-Assume that all auxiliary assumption, regarding proper quantification of behavior and situation in valid measurements, causality, individual differences, nuisance variables, distributional, etc, hold.
-
-While not using statistical methods would lead to the rejection of almost all psychological theories, pushed to the extreme statistical methods can shield theories from refutation altogether.
-Both situations are uninformative and, therefore, rather unhelpful for the generation of knowledge.
-The later situation, however, becomes less than helpful when one is under the impression that theories are well tested, when in fact there was little chance that the data could refute the theory if wrong.
-
 # References

### 2023-01-04 manuscript.Rmd
grammerly

@@
 Here, it is innocuous because the introduced bias can be accounted for and vanishes with increasing sample size, but we will see later that this is not always the case.
-In this specific example, we use a t-test instead of a z-test to account for the estimation of variance from data, and it is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
+In this specific example, we use a t-test instead of a z-test to account for variance estimation from data.
+It is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
 Induction is also necessary because it is almost unimaginable to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
@@
 In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled with data.
-It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions so that they fit the data well.
+It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.
 In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
@@
 If researchers use induction for some necessary but under the theory arbitrary assumptions, their theory will not be refuted because of these assumptions.
-Since empirical researcher often cannot derive every assumption from their theory, avoiding refutation because of those assumptions is a desirable property.
+Since empirical researchers often cannot derive every assumption from their theory, avoiding refutation because of those assumptions is a desirable property.
 
@@
 Such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
-If pushed to the limit, we arrive at a "theory" that is governed by the data.
+If pushed to the limit, we arrive at a "theory" governed by the data.
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction, can not refute these statements, so they are not tested at all.
+The data used for induction can not refute these statements, so they are not tested at all.
 
@@
 The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims are.
-The question of this dissertation is, therefore, what must be made transparent, and how can we make it transparent?
+Therefore, the question of this dissertation is what must be made transparent, and how can we make it transparent?
 
@@
 The first framework motivates transparency when the aim of an empirical test is to evaluate the verisimilitude ("truth likeness") of a theory.
-The second framework motivates transparency under a sciences that wants to select a theory according to its expected predictive performance.
+The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
 
-## An information theoretic perspective
+## An information-theoretic perspective
 
@@
 That is, how much information about the truth is lost when the theory is used to model reality.
-Expressed mathematically we have a function $f(x)$ that gives us the probability to observe the state of the world $x$ where $f$ represents the full reality.
-We are now interested, how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $x$.
+Expressed mathematically, we have a function $f(x)$ that gives us the probability to observe the state of the world $x$ where $f$ represents the full reality.
+We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $x$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
@@
 
-Most readers will recognize, that this information theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
+Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
 7.2, in their derivation of the Akaike Information Criterion in its general form.
-What is of interest here, is not the derivation, but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
+What is of interest here is not the derivation but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
 
-This setup is of course highly theoretical.
+This setup is, of course, highly theoretical.
 $\mathcal{I}(f, g)$ is unknowable, since the truth is unobserved.
-This fact, however, does not impede us from getting closer to the truth because we still can compare two different theories relative to each other.
-Because the expectation for $f$ remains constant (left hand side) we only need to estimate the relative expected loss of information (right hand side) to make a comparative judgment.
+However, this fact does not impede us from getting closer to the truth because we can still compare two theories relative to each other.
+Because the expectation for $f$ remains constant (left-hand side), we only need to estimate the relative expected loss of information (right-hand side) to make a comparative judgment.
 To make a relative judgment about several competing theories it suffices to estimate:
@@
 To allow for quantities to be induced, we must assume that our theory is parametrized, e.g., $g(x|\theta)$.
-That means our theory implies a multitude of possible probability distribution that may describe reality.
-Though, this assumption may be lifted, we assume each implied version of the theory equally probable a priori.
+That means our theory implies a multitude of possible probability distributions that may describe reality.
+Though this assumption may be lifted, we assume each implied version of the theory is equally probable a priori.
 This parametrization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
-Of those arbitrary assumptions we want to find those that fit the reality with the least amount of information lost.
+Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
 The best parametrization is archived by:
@@
 
-The inference goal is therefore:
+The inference goal is, therefore:
 
@@
 That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$.
-The crucial point here is to understand what happens, when we can not derive $\theta$, deductively, but must substitute it inductively with an estimate $\hat\theta(y)$.
+The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_o$.
-It follows that, almost surely:
+It follows that almost surely:
 
@@
 
-That is to say, any induced estimate will be sub optimal.
+That is to say, any induced estimate will be sub-optimal.
 The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
 The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
-If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased we may form an expectation over the data $y$:
+If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased, we may form an expectation over the data $y$:
 
@@
 
-Forming this expectation over data is a crucial step, it requires us to think beyond the data we observed to all the data we could have observed.
+Forming this expectation over data is a crucial step; it requires us to think beyond the data we observed to all the data we could have observed.
 There are two ways to get at this expectation.
-One follows in this section, another is discussed in the next section.
+One follows in this section, and another is discussed in the next section.
 
@@
 In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
-Instead of verisimilitude one might be concerned with future performance.
+Instead of verisimilitude, one might be concerned with future performance.
 That is, how well does a theory do in predicting novel facts.
-Please note that the information theoretic setup above has made no appeal to expected performance on unseen data.
+Please note that the information-theoretic setup above has made no appeal to the expected performance on unseen data.
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
@@
 
-Still the core problems of limited sample and inductive process remain.
+Still, the core problems of limited sample and inductive process remain.
 Again there exists only a sample $y$ over which we form the expectation, but now we must pay a little more attention to the individual elements, $i = 1, \dots,n$:
@@
 
-Since, we allow arbitrary $L$, the trick of using the Taylor series expansion does not get us very far (because we simply end up with an undefined derivative).
-We can, however, repeatedly sample data and repeate the inductive process.
-That is, we make use of the cross-validatory where the data is partitioned and the inductive process is repeated on all permutations of a subset of the partitions.
+Since we allow arbitrary $L$, the trick of using the Taylor series expansion does not get us very far (because we simply end up with an undefined derivative).
+We can, however, repeatedly sample data and repeat the inductive process.
+That is, we make use of the cross-validatory where the data is partitioned, and the inductive process is repeated on all permutations of a subset of the partitions.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
@@
 
-If we reinstate Kullback-Leibler divergence as a loss function, we can estimate $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$ using the cross-validatory which connects the information theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
-To make the link to the first approach even more clear we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
+If we reinstate Kullback-Leibler divergence as a loss function, we can estimate $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$ using the cross-validatory which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
 
@@
 
-Instead of a formal analyses to derive $\mathcal{C}$, we can simple repeat the inductive process, i.e., using cross-validation.
-That drastically expands the set of inductive processes for which we can estimate the inductive bias, since it is considerably easier to repeat the process than to derive the complexity analytically.
+Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
+That drastically expands the set of inductive processes for which we can estimate the inductive bias since it is considerably easier to repeat the process than to derive the complexity analytically.
 
@@
 Showing the inductive results instead of the process is insufficient to appraise the theory.
-On a conceptual level we want to compare:
+On a conceptual level, we want to compare:
 
@@
 
-But to allow for induction to happen we replace theory by a model (not necessarily a statistical one), or put differently a multitude of implications about the data from the theory.
+But to allow for induction to happen, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
 This multitude is a result of arbitrary assumptions.
@@
 However, we are forced to rely on a limited sample of reality.
-We are mislead because these two factors, induction and limited sample size, interact.
+We are misled because these two factors, induction and limited sample size, interact.
 
@@
 
-Transparency is necessary because induction leaves researchers overly optimistic regarding their theories fit to the data.
-The extend of this optimism is dependent on the process of induction and, hence, researcher ought to make it transparent.
+Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data.
+The extent of this optimism depends on the induction process, and hence, researchers ought to make it transparent.
 
@@
 The above sections aimed to motivate the observation that the apparent fit of a theory to data is almost always overly optimistic.
-This observations is only of limited use, if we do not now the extend of this optimism.
-However, both setups show that the extend of the optimism ($\mathcal{C}$) is closely related to the inductive process and give us two starting points for making this bias transparent.
-The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), the second merely requires that the process is repeatable (e.g., using cross-validation).
+This observation is only of limited use if we do not now the extend of this optimism.
+However, both setups show that the extent of the optimism ($\mathcal{C}$) is closely related to the inductive process and give us two starting points for making this bias transparent.
+The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
 Both approaches require researchers to make the process of induction transparent instead of merely publishing the results.
@@
 
-### Theorethical
+### Theoretical
 

### 2023-01-05 manuscript.Rmd
update

@@
 Though there might be some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Laying open the difficulties and attempting to overcome them is not a recent trend, though it has been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
-The debates about psychology's shortcomings and remedies fall roughly into two categories: statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and sociological factors [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+Laying open the difficulties and attempting to overcome them is not a recent trend, though it has been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that begins to ripple through other empirical sciences.
+As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
+We can roughly divide the causes into two categories: misuse of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and sociological factors [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
 <!--bla-transparency-bla-->
-I propose to view both categories through the lens of transparency about the inductive process.
-One side of transparency is that statistical methods make induction quantifiable; the other is that open science measures remove uncertainty about sociological factors in induction.
-Any science must be able to communicate how it generates its knowledge, but transparency has an outstanding role in psychology and other empirical sciences.
-To understand why transparency is crucial, we must understand that induction is integral to empirical sciences, formally using statistical procedures or informally as researchers' judgment.
-However, empirical statements lose their value without transparency about the inductive process that contributed to them.
-Therefore, transparency is more than a virtue that somewhat improves psychological sciences but is an indispensable property.
+In my view, both categories emerge from a lack of transparency about the inductive process in the empirical test of a theory.
+Testing a theory empirically is often viewed as deductive, since the theory is making statements about the data.
+Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data leading to overconfidence in their empirical results.
+Some parts of their inductive process are well defined in the form of statistical methods, while other parts happen more informally.
+
+Statistical methods make induction quantifiable; while open science measures reduce uncertainty about sociological factors in induction.
+To understand why transparency is crucial, we must understand that these two forms of induction are integral to empirical sciences.
+Any science must be able to communicate how it generates its knowledge.
+However, transparency has an outstanding role in psychology and other empirical sciences, because empirical statements lose their value without transparency about the inductive process that contributed to them.
+Therefore, transparency is more than a virtue that somewhat improves empirical sciences but is an indispensable property.
 
@@
 That is, how much information about the truth is lost when the theory is used to model reality.
-Expressed mathematically, we have a function $f(x)$ that gives us the probability to observe the state of the world $x$ where $f$ represents the full reality.
-We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $x$.
+Expressed mathematically, we have a function $f(x)$ that gives us the liklihood to observe the state of the world $x$ where $f$ represents the ground truth.
+We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
 
+<!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
+
 $$
-\mathcal{I}(f, g) = \mathbb{E}_x\left[log(f(x))\right] - \mathbb{E}_x\left[log(g(x))\right]
+\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(g(x)^{-1}\right) - \int_{\mathcal{X}}f(x)\log\left(f(x)^{-1}\right)
+$$
+
+$$
+\mathcal{L}_{KL}(f, g) = \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]
 $$
@@
 This setup is, of course, highly theoretical.
-$\mathcal{I}(f, g)$ is unknowable, since the truth is unobserved.
+$\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
 However, this fact does not impede us from getting closer to the truth because we can still compare two theories relative to each other.
-Because the expectation for $f$ remains constant (left-hand side), we only need to estimate the relative expected loss of information (right-hand side) to make a comparative judgment.
+Because the expectation for $f$ remains constant (left-hand expectation), we only need to estimate the relative expected loss of information (right-hand expectation) to make a comparative judgment.
 To make a relative judgment about several competing theories it suffices to estimate:
@@
 $$
-\mathbb{E}_x\left[log(g(x))\right]
+\mathbb{E}_{X \sim f}\left[\log(g(x))\right]
 $$
@@
 Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
-The best parametrization is archived by:
+The best parametrization is achieved by:
 
 $$
-\theta_o = \operatorname*{arg\,min}_\theta \mathcal{I}(f, g(\cdot|\theta))
+\theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))
 $$
@@
 $$
-\mathbb{E}_x\left[log(g(x|\theta_o))\right]
+\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right]
 $$
 
-Of course, we usually do not know $\theta_o$.
-That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$.
+Of course, we usually do not know $\theta_*$.
+That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y \sim f^n$, that is it follows the empirical distribution.
 The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
-Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_o$.
+Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
 It follows that almost surely:
@@
 $$
-\mathcal{I}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{I}(f, g(\cdot|\theta_o))
+\mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))
 $$
 
-Or ignoring $f$ as constant:
+Or ignoring the constant:
 
 $$
-\mathbb{E}_x[log(g(x|\hat{\theta}(y)))] > \mathbb{E}_x\left[log(g(x|\theta_o))\right] 
+\mathbb{E}_{X \sim f}[log(g(x|\hat{\theta}(y)))] > \mathbb{E}_{X \sim f}\left[log(g(x|\theta_o))\right] 
 $$
@@
 $$
-\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))]
 $$
@@
 There are two ways to get at this expectation.
-One follows in this section, and another is discussed in the next section.
+One is the use of Taylor series expansion, which follows in this section, and another is the use of cross-validation discussed in the next section.
 
-We might notice that we often get an unbiased estimate for $\mathbb{E}_x[log(g(x|\hat{\theta}(y)))]$, e.g., in maximum likelihood estimation.
+We might notice that we often get an unbiased estimate for $\mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))]$, e.g., in maximum likelihood estimation.
 The expectation over the data together with Taylor series expansion yields:
@@
 $$
-\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_x[log(g(x|\hat\theta(y)))] - tr[J(\theta_o)I(\theta_o)^{-1}]
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]
 $$
@@
 The observed likelihood $log(g(x|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-Note that $tr[J(\theta_o)I(\theta_o)^{-1}]$ is usually negative definitive and, therefore, we may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
+Note that $tr[J(\theta_*)I(\theta_*)^{-1}]$ is usually negative definitive and, therefore, we may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
 I will refer to this bias as the complexity of a model, denoted as $\mathcal{C}$, because it is closely related to how much the data is influencing the results and, hence, how detailed the model is representing the data.
@@
 $$
-\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_x[log(g(x|\hat\theta(y)))] + \mathcal{C}
+\mathbb{E}_Y\mathbb{E}_X[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_X[log(g(x|\hat\theta(y)))] + \mathcal{C}
 $$
@@
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
-One condition is, since $\theta_o$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
+One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
 We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are available.
@@
 Still, the core problems of limited sample and inductive process remain.
-Again there exists only a sample $y$ over which we form the expectation, but now we must pay a little more attention to the individual elements, $i = 1, \dots,n$:
+Again there exists only a sample $y$ over which we form the expectation, but now we must pay a little more attention to the individual elements, $i = 1, \dots,n$ and take the expectation [@sochLawUnconsciousStatistician2020]:
 
@@
 \mathbb{E}_y L(y, g(y|\theta))
- = \sum_{i=1}^n L(y_i, g(y_i|\theta)) \frac{1}{n}
+ = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\theta))
 $$
 
-Since we allow arbitrary $L$, the trick of using the Taylor series expansion does not get us very far (because we simply end up with an undefined derivative).
-We can, however, repeatedly sample data and repeat the inductive process.
-That is, we make use of the cross-validatory where the data is partitioned, and the inductive process is repeated on all permutations of a subset of the partitions.
+Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
+That is, we make use of cross-validation where the data is partitioned, and the inductive process is repeated on all permutations of a subset of the partitions.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
@@
 
-If we reinstate Kullback-Leibler divergence as a loss function, we can estimate $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$ using the cross-validatory which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+As stated earlier, using cross-validation we can also estimate the If we reinstate Kullback-Leibler divergence as a loss function, we can estimate $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$ using the cross-validatory which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
 To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:

### 2023-01-05 manuscript.Rmd
correct integral
add some notational things

@@
 That is, how much information about the truth is lost when the theory is used to model reality.
-Expressed mathematically, we have a function $f(x)$ that gives us the liklihood to observe the state of the world $x$ where $f$ represents the ground truth.
-We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+Expressed mathematically, we have a random variable $X ~ f$, with values in $\mathcal{X}$, where the distribution $f$ represents the ground truth.
+We are now interested in how much information is lost if we use $g$, our theory, instead of $f$, the reality, to approximate $X$ over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
@@
 
-$$
-\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(g(x)^{-1}\right) - \int_{\mathcal{X}}f(x)\log\left(f(x)^{-1}\right)
-$$
-
-$$
-\mathcal{L}_{KL}(f, g) = \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]
-$$
+\begin{align}
+\mathcal{L}_{KL}(f, g) &= \int_{\mathcal{X}}f(x)\log\left(f(x)\right) - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\\
+ &= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]
+\end{align}
 

### 2023-01-06 manuscript.Rmd
refine notation further

--- 2023-01-05 manuscript.Rmd
@@
 $$
-\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(g(x)^{-1}\right) - \int_{\mathcal{X}}f(x)\log\left(f(x)^{-1}\right)
+\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x
 $$
@@
 
-To allow for quantities to be induced, we must assume that our theory is parametrized, e.g., $g(x|\theta)$.
+To allow for quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
 That means our theory implies a multitude of possible probability distributions that may describe reality.
@@
 Of course, we usually do not know $\theta_*$.
-That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y \sim f^n$, that is it follows the empirical distribution.
+That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ is independently sampled from $X\sim f$.
 The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
@@
 $$
-\mathbb{E}_{X \sim f}[log(g(x|\hat{\theta}(y)))] > \mathbb{E}_{X \sim f}\left[log(g(x|\theta_o))\right] 
+\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
 $$
@@
 The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
-If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased, we may form an expectation over the data $y$:
+If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
 
@@
 
-We might notice that we often get an unbiased estimate for $\mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))]$, e.g., in maximum likelihood estimation.
+We might notice that we often get an unbiased estimate for the observed distance on the data used to induce $\hat \theta$.
+With slight abuse of notation, $log(g(y|\theta)) \equiv \sum_{i = 1}^n log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
+The observed likelihood is often easy enough to obtain, e.g. in maximum likelihood estimation.
+
 The expectation over the data together with Taylor series expansion yields:
@@
 
+Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$ respectively.
+For more details see of this derivation see @burnhamModelSelectionMultimodel2002, Ch. 7.2.
+
 The observed likelihood $log(g(x|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
@@
 $$
-\mathbb{E}_Y\mathbb{E}_X[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_X[log(g(x|\hat\theta(y)))] + \mathcal{C}
+\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
 $$
@@
 $$
-\mathcal{L}(y, g(y|\theta)) = \mathbb{E}_yL(y, g(y|\theta))
+\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
 $$
@@
 $$
-\mathbb{E}_y L(y, g(y|\hat\theta(y)))
- = \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))\frac{1}{n}
+\mathbb{E}_{x \sim f} L(x, g(x|\hat\theta(y)))
+ = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
 $$

### 2023-01-06 manuscript.Rmd
take a look at the cross-validation section

@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return probabilities:
+In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood:
 
@@
 
-Still, the core problems of limited sample and inductive process remain.
-Again there exists only a sample $y$ over which we form the expectation, but now we must pay a little more attention to the individual elements, $i = 1, \dots,n$ and take the expectation [@sochLawUnconsciousStatistician2020]:
+Again we can define the loss we have observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020]:
 
@@
 $$
-\mathbb{E}_y L(y, g(y|\theta))
+\mathbb{E}_y L(y, g(y|\hat\theta(y)))
  = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\theta))
@@
 
+However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
+
+$$
+\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
+$$
+This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004] and resembles the expectation over data discussed in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
 Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
@@
 $$
-\mathbb{E}_{x \sim f} L(x, g(x|\hat\theta(y)))
+\mathbb{E}_{x}\mathbb{E}_{y}  L(x, g(x|\hat\theta(y)))
  = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
@@
 
-As stated earlier, using cross-validation we can also estimate the If we reinstate Kullback-Leibler divergence as a loss function, we can estimate $\mathbb{E}_y\mathbb{E}_x[log(g(x|\hat\theta(y)))]$ using the cross-validatory which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+As stated earlier, using cross-validation we can also estimate \nobreak\hfil\penalty0 \hfilneg $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
 To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:

### 2023-01-06 manuscript.Rmd
clarify split

@@
 Testing a theory empirically is often viewed as deductive, since the theory is making statements about the data.
-Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data leading to overconfidence in their empirical results.
-Some parts of their inductive process are well defined in the form of statistical methods, while other parts happen more informally.
+Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
+The inductive element in the empirical test leads to overconfidence in the empirical results, if unaccounted.
+An accurate judgment of the empirical support of a theory is only possible if the inductive process is made transparent.
 
+The above distinction arises because some parts of the inductive process are well defined in the form of statistical methods, while other parts happen more informally.
 Statistical methods make induction quantifiable; while open science measures reduce uncertainty about sociological factors in induction.
@@
 <!--induction is necessary for psychological science-->
-To function as a science, psychology must be able to make statements about the world that can be compared to the true conditions.
+To function as an empirical science, psychology must be able to make statements about the world that can be compared to the true conditions.
 In psychology, this is not a purely deductive endeavor.
 Very few psychological theories are precise enough to derive testable statements.
-While it is tempting to claim that a theory makes testable claims by, e.g., implying a mean difference between two groups, such a statement is not testable on its own.
+While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
 Consider a simple t-test of mean differences.
@@
 So what to think of such, yet untested, theory?
-Researchers (and philosophers of science) differ considerably in how they think theories should be appraised, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
+Researchers and philosophers of science differ considerably their opinion about to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
 14] of a hypothesis.
-But whatever measure they subscribe to, their appraisal of the empirical support of a yet untested theory would be relatively low.
+Whatever measure they subscribe to, an untested theory should not score very high on it.
 
@@
 
-As stated earlier, using cross-validation we can also estimate \nobreak\hfil\penalty0 \hfilneg $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+As stated earlier, using cross-validation we can also estimate \allowbreak $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
 To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:

### 2023-01-06 manuscript.Rmd
typo

@@
 So what to think of such, yet untested, theory?
-Researchers and philosophers of science differ considerably their opinion about to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
+Researchers and philosophers of science differ considerably in their opinion about how to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
 14] of a hypothesis.
-Whatever measure they subscribe to, an untested theory should not score very high on it.
+Whatever measure they subscribe to, an untested theory should not score very high.
 

### 2023-01-06 manuscript.Rmd
make transitions softer

@@
 
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
+Because both are quite technical, they are followed by a more conceptual summary of these tools.
+These sections provide the basis to understand how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.
+
 ## An information-theoretic perspective
@@
 That means our theory implies a multitude of possible probability distributions that may describe reality.
-Though this assumption may be lifted, we assume each implied version of the theory is equally probable a priori.
-This parametrization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
+This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
 Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
-The best parametrization is achieved by:
+The best parameterization is achieved by:
 
@@
 Note that $tr[J(\theta_*)I(\theta_*)^{-1}]$ is usually negative definitive and, therefore, we may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
-I will refer to this bias as the complexity of a model, denoted as $\mathcal{C}$, because it is closely related to how much the data is influencing the results and, hence, how detailed the model is representing the data.
+This overconfidence is directly related to how much induction a model entails.
+This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowDeepLearning2016; @goodfellowCapacityOverfittingUnderfitting2016].
+I therefore denote it as as $\mathcal{C}$.
 
@@
 
-As stated earlier, using cross-validation we can also estimate \allowbreak $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+As stated earlier, using cross-validation it is possible to  estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
 To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
@@
 
-Now that we have established the need for transparency about the inductive process, we can drop a few of the more technical details to get a more straightforward answer about the how.
+Now that we have established the need for transparency about the inductive process, we can drop a few of the more technical details to get a more straightforward about what we have to make transparent.
 It bears repeating that simply laying open what has been done is not enough.
@@
 But to allow for induction to happen, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
-This multitude is a result of arbitrary assumptions.
 
@@
 Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data.
-The extent of this optimism depends on the induction process, and hence, researchers ought to make it transparent.
+The extent of this optimism depends on the inductive process, not merely its results.
+Specifically, it depends on the complexity, i.e., the ability of the inductive process to adept to data.
+Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
 
@@
 
-The above sections aimed to motivate the observation that the apparent fit of a theory to data is almost always overly optimistic.
+The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic.
 This observation is only of limited use if we do not now the extend of this optimism.

### 2023-01-09 manuscript.Rmd
add transition to how

@@
 
-The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic.
+The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic, if it has inductive elements.
 This observation is only of limited use if we do not now the extend of this optimism.
@@
 
+Even a casual consideration of the above formalization should strike anyone who has ever worked with empirical data as unrealistic.
+No one can actually expect researchers to be inductive only in ways that are formally analyzable or even strictly repeatable.
+The point is to set the goal post and have a yardstick to measure how well a method aimed at improving empirical sciences does.
+
+It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
+This fact implies that for theses situations the optimism bias can not be fully quantified and that full transparency can not be archived.
 To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
 Otherwise, the supposedly objective test of the theory using hard data must still be judged subjectively.
+
+We, therefore, have two bounds that limit transparency about the inductive process.
+First, some things can not be made transparent in principle, in the sense that we can not estimate the optimism with certainty.
+Second, even formalized inductive processes only allow to estimate optimism theoretically but they have to be communicated effectively to do so in practice.
+
+In the following, I propose preregistration, as means to move induction into the formal domain, and computational reproducibility to make formal induction transparent.
 

### 2023-01-10 manuscript.Rmd
fix sentance
add most of andreas suggestions (minus references)

@@
 Psychology is a difficult science [@meehlTheoreticalRisksTabular1978].
-Though there might be some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Laying open the difficulties and attempting to overcome them is not a recent trend, though it has been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that begins to ripple through other empirical sciences.
+Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
+Laying open the difficulties and attempting to overcome them are not a recent trends, though they have been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that also begins to ripple through other empirical sciences.
 As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
-We can roughly divide the causes into two categories: misuse of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and sociological factors [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+We can roughly divide the proposed countermeasures into two categories, those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that aim to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
 <!--bla-transparency-bla-->
-In my view, both categories emerge from a lack of transparency about the inductive process in the empirical test of a theory.
+In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory.
 Testing a theory empirically is often viewed as deductive, since the theory is making statements about the data.
 Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
-The inductive element in the empirical test leads to overconfidence in the empirical results, if unaccounted.
-An accurate judgment of the empirical support of a theory is only possible if the inductive process is made transparent.
+I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results, if unaccounted.
+The extent of this bias depends on the inductive process.
+Without transparency about the inductive process, researchers are unable to judge the empirical support of a theory.
 
-The above distinction arises because some parts of the inductive process are well defined in the form of statistical methods, while other parts happen more informally.
-Statistical methods make induction quantifiable; while open science measures reduce uncertainty about sociological factors in induction.
-To understand why transparency is crucial, we must understand that these two forms of induction are integral to empirical sciences.
+The above distinction arises because some parts of the inductive process are well defined in the form of statistical methods, while researcher also engage in more informal inductive behavior.
+Statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about remaining informal sources of inductive bias.
+To understand why transparency is crucial, we must understand that formal and informal induction are integral to empirical sciences.
 Any science must be able to communicate how it generates its knowledge.
 However, transparency has an outstanding role in psychology and other empirical sciences, because empirical statements lose their value without transparency about the inductive process that contributed to them.
-Therefore, transparency is more than a virtue that somewhat improves empirical sciences but is an indispensable property.
+Therefore, transparency is more than a virtue that may improve knowledge gains in empirical sciences but is an indispensable property.
 
 <!--induction is necessary for psychological science-->
-To function as an empirical science, psychology must be able to make statements about the world that can be compared to the true conditions.
+To function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
 In psychology, this is not a purely deductive endeavor.
@@
 While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
-Consider a simple t-test of mean differences.
-The decision to reject the null hypothesis still depends, besides true mean difference and sample size, on the observed variance.
-However, the variance is, of course, estimated from the data.
+Consider the example of mean difference between a placebo and an experimental group, that we test using a simple t-test of mean differences.
+The decision to reject the null hypothesis (no mean difference between groups) still depends, besides true mean difference and sample size, on the observed variance.
+However, the variance needs to be estimated from the data.
 So the threshold of the deductive decision depends on a quantity that must be induced.
@@
 Here, it is innocuous because the introduced bias can be accounted for and vanishes with increasing sample size, but we will see later that this is not always the case.
-In this specific example, we use a t-test instead of a z-test to account for variance estimation from data.
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
+z≈`r round(qnorm(.99), 2)`).
 It is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
-Induction is also necessary because it is almost unimaginable to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would probably be no psychological theory that could survive an empirical test.
+Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
+If they had to, there would probably be no psychological theory that could be subject of an empirical test.
 
 <!--necessity of induction-->
-In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled with data.
+In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
 It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.
@@
 By the same token, whole theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
-Such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
+If strictly applied to auxiliary assumptions, such strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
 If pushed to the limit, we arrive at a "theory" governed by the data.
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction can not refute these statements, so they are not tested at all.
+The data used for induction can not refute these statements, so they they never have been subject of an empierical test.
 
@@
 14] of a hypothesis.
-Whatever measure they subscribe to, an untested theory should not score very high.
+Whatever measure they subscribe to, they would agree on a low appraisal of an untested theory.
 
@@
 On the other hand, induction may render any test of a hypothesis ineffective.
-The problem, I argue, therefore, is not induction but making induction transparent.
-The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims are.
-Therefore, the question of this dissertation is what must be made transparent, and how can we make it transparent?
+The problem, I argue, therefore, is not induction but making transparent where and to what extend induction is used in the inferential process.
+The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims as reported in the literature actually are.
+Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?
 
 <!--forshadow the main part-->
-The first question (the "what") is theoretical in nature.
-It is addressed in the thesis itself, which supplies the theoretical framework of the articles written as part of the dissertation.
+The first question (the *what*) is theoretical in nature.
+It is addressed in the synopses, which supplies the theoretical framework of the articles written as part of the dissertation.
 Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).
@@
 <!--how it relates to the articles-->
-Based on this split, the articles answer the second question (the "how").
+Based on this distinction, the articles answer the second question (the *how*).
 I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these tools.
+The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches
 
@@
 The need for transparency is closely tied to the use of induction in the empirical test of a theory.
-There has been a long and vigorous debate about what it means to test a theory empirically.
+There has been a long and vigorous debate about what it means to test a theory empirically (xxx).
 I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
-The first framework motivates transparency when the aim of an empirical test is to evaluate the verisimilitude ("truth likeness") of a theory.
+The first framework motivates transparency when the aim of an empirical test is to evaluate the verisimilitude ("closeness to truth") of a theory.
 The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
@@
 Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory.
-That is, how much information about the truth is lost when the theory is used to model reality.
-Expressed mathematically, we have a function $f(x)$ that gives us the liklihood to observe the state of the world $x$ where $f$ represents the ground truth.
+We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality.
+Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood to observe the state of the world $x$ where $f$ represents the ground truth.
 We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
@@
 
-$$
-\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x
-$$
-
-$$
-\mathcal{L}_{KL}(f, g) = \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]
-$$
+\begin{align} 
+\mathcal{L}_{KL}(f, g) &=\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x\\
+&= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]\label{eq:klexp}
+\end{align}
 
 Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
-7.2, in their derivation of the Akaike Information Criterion in its general form.
+7.2, in their derivation of the Akaike Information Criterion (XXX) in its general form.
 What is of interest here is not the derivation but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
@@
 However, this fact does not impede us from getting closer to the truth because we can still compare two theories relative to each other.
-Because the expectation for $f$ remains constant (left-hand expectation), we only need to estimate the relative expected loss of information (right-hand expectation) to make a comparative judgment.
-To make a relative judgment about several competing theories it suffices to estimate:
+Because the expectation for $f$ remains constant (left-hand expectation in Eq.
+\@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.
+\@ref(eq:klexp)) to make a comparative judgment.
+To make a relative judgment about several competing theories it suffices to estimate for any theory $g(x)$:
 
@@
 To allow for quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
-That means our theory implies a multitude of possible probability distributions that may describe reality.
+That means our theory implies a family of possible probability distributions that may describe reality.
 This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
@@
 
-The inference goal is, therefore:
+The inference goal is for comparing other theories to $g(x)$ is, therefore:
 
@@
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
-It follows that almost surely:
+It follows, almost surely, that we loose information:
 
@@
 
-We might notice that we often get an unbiased estimate for the observed distance on the data used to induce $\hat \theta$.
+We usually favor procedures that promise unbiased estimates for the observed distance on the data used to induce $\hat \theta$, given that their assumptions are met.
 With slight abuse of notation, $log(g(y|\theta)) \equiv \sum_{i = 1}^n log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
@@
 One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are available.
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known (XXX).
 
@@
 It bears repeating that simply laying open what has been done is not enough.
-Showing the inductive results instead of the process is insufficient to appraise the theory.
+Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
 On a conceptual level, we want to compare:
@@
 
-But to allow for induction to happen, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
+Where, $\mathcal{L}$ stands for loss function, i.e., how to compare predictions and reality.
+To allow for induction to happen, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
 
@@
 We, therefore, have two bounds that limit transparency about the inductive process.
-First, some things can not be made transparent in principle, in the sense that we can not estimate the optimism with certainty.
+First, some things can not be made transparent in principle because some induction happens informally, in the sense that we can not estimate the optimism with certainty.
 Second, even formalized inductive processes only allow to estimate optimism theoretically but they have to be communicated effectively to do so in practice.

### 2023-01-11 manuscript.Rmd
add pdfpages

@@
    - \usepackage{libertine}
+   - \usepackage{pdfpages}
 fontsize: 12pt
@@
     - libertine
+    - pdfpages
   apt:
@@
 
+# Published Papers
+
 # References

### 2023-01-11 manuscript.Rmd
add paper pdfs

@@
 
-# Published Papers
+# Articles
+
+## A Reproducible Data Analysis Workflow With R Markdown, Git, Make, and Docker
+
+The following article is reprinted from the following source under [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).
+
+Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow With
+R Markdown, Git, Make, and Docker. Quantitative and Computational Methods in Behavioral Sciences, 1, e3763. https://doi.org/10.5964/qcmb.3763
+
+\includepdf[pages=-]{papers/workflow.pdf}
+
+## Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once
+
+The following article is reprinted from the following source under [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).
+
+Peikert, A., van Lissa, C.
+J., & Brandmaier, A.
+M.
+(2021).
+Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once.
+*Psych*, *3*(4), 836–867.
+https://doi.org/10.3390/psych3040053
+
+\includepdf[pages=-]{papers/tutorial.pdf}
+
+## Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.
+
+The following article is prepared to be submitted to *Philosophical Psychology*.
+The preprint is reprinted here under [Creative Commons Public Domain Dedication (CC0 1.0)](https://creativecommons.org/publicdomain/zero/1.0/).
+
+\includepdf[pages=-]{papers/prereg.pdf}
 

### 2023-01-11 manuscript.Rmd
manuscript depends on papers

@@
 ---
-title: "Towards Transparency and Open Science"
+title: "\\includepdf{papers/workflow.pdf}Towards Transparency and Open Science"
 subtitle: "A Principled Perspective on Computational Reproducibility and Preregistration"
@@
     - apa.csl
+    - papers/
   tlmgr:

### 2023-01-12 manuscript.Rmd
add cover and fake abstracts in two languages

--- 2023-01-11 manuscript.Rmd
@@
     number_sections: false
-header-includes:
-   - \usepackage{libertine}
-   - \usepackage{pdfpages}
+header-includes: |
+  \usepackage{libertine}
+  \usepackage{pdfpages}
+  \usepackage{titling}
+  \pretitle{\begin{center}\LARGE
+     \begin{minipage}{\textwidth}
+     \includepdf{cover.pdf}
+     \end{minipage}
+     \newpage}
+  \posttitle{\end{center}}
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+  \setotherlanguage[]{german}
 fontsize: 12pt
-geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
+geometry: "left=3.6cm, right=3.6cm, top=4cm, bottom=4cm, bindingoffset=5mm"
 papersize: a4
 bibliography: references.bib
-abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
+abstract: "`r paste0(c('English:', readLines('abstract.md'), '\\\\par Deutsch:\\\\textgerman{', readLines('zusammenfassung.md'), '}\\\\newpage'), collapse = ' ')`"
 csl: apa.csl
@@
 
-\includepdf[pages=-]{papers/workflow.pdf}
+\includepdf[pages=-,fitpaper=true, offset=5mm 0mm,frame=false]{papers/workflow.pdf}
 
@@
 
-\includepdf[pages=-]{papers/tutorial.pdf}
+\includepdf[pages=-,fitpaper=true, offset=5mm 0mm,frame=false]{papers/tutorial.pdf}
 
@@
 
-\includepdf[pages=-]{papers/prereg.pdf}
+\includepdf[pages=-,fitpaper=true, offset=5mm 0mm,frame=false]{papers/prereg.pdf}
 

### 2023-01-12 manuscript.Rmd
adjust text spacing, number and space equations

@@
 fontsize: 12pt
+linestretch: 1.2
 geometry: "left=3.6cm, right=3.6cm, top=4cm, bottom=4cm, bindingoffset=5mm"
@@
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
-
 <!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
-
 \begin{align} 
@@
 To make a relative judgment about several competing theories it suffices to estimate for any theory $g(x)$:
-
-$$
+\begin{equation}
 \mathbb{E}_{X \sim f}\left[\log(g(x))\right]
-$$
+\end{equation}
 
@@
 The best parameterization is achieved by:
-
-$$
+\begin{equation}
 \theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))
-$$
+\end{equation}
 
 The inference goal is for comparing other theories to $g(x)$ is, therefore:
-
-$$
+\begin{equation}
 \mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right]
-$$
-
+\end{equation}
 Of course, we usually do not know $\theta_*$.
 That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ is independently sampled from $X\sim f$.
+
 The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
@@
 It follows, almost surely, that we loose information:
-
-$$
+\begin{equation}
 \mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))
-$$
-
+\end{equation}
 Or ignoring the constant:
-
-$$
+\begin{equation}
 \mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
-$$
+\end{equation}
 
@@
 If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
-
-$$
+\begin{equation}
 \mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))]
-$$
+\end{equation}
 
@@
 The expectation over the data together with Taylor series expansion yields:
-
-$$
+\begin{equation}
 \mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]
-$$
+\end{equation}
 
@@
 I therefore denote it as as $\mathcal{C}$.
-
-$$
+\begin{equation}
 \mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
-$$
+\end{equation}
 
@@
 In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood:
-
-$$
+\begin{equation}
 \mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
-$$
+\end{equation}
 
 Again we can define the loss we have observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020]:
-
-<!--the unconscious statistician strikes again-->
-
-$$
+\begin{equation}
 \mathbb{E}_y L(y, g(y|\hat\theta(y)))
  = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\theta))
-$$
+\end{equation}
 
 However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
+\begin{equation}
+\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
+\end{equation}
+This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004] and resembles the expectation over data discussed in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
 
-$$
-\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
-$$
-This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004] and resembles the expectation over data discussed in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
 Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
@@
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
-
-$$
+\begin{equation}
 \mathbb{E}_{x}\mathbb{E}_{y}  L(x, g(x|\hat\theta(y)))
  = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
-$$
+\end{equation}
 
@@
 To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
-
-$$
+\begin{equation}
 \mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta))
-$$
+\end{equation}
 
@@
 On a conceptual level, we want to compare:
-
-$$
+\begin{equation}
 \mathcal{L}(\text{Theory}, \text{Reality})
-$$
+\end{equation}
 
@@
 
-$$
+\begin{equation}
 \mathcal{L}(\text{Model(Reality)}, \text{Reality})
-$$
-
+\end{equation}
 The idea is that we choose from the multitude the version of our theory that best fits reality.
@@
 We are misled because these two factors, induction and limited sample size, interact.
-
-$$
+\begin{equation}
 \mathcal{L}(\text{Model(Reality)}, \text{Reality}) > \mathcal{L}(\text{Model(Sample)}, \text{Sample})
-$$
-
-$$
+\end{equation}
+\begin{equation}
 \mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}
-$$
+\end{equation}
 

### 2023-01-12 manuscript.Rmd
adjust spacing for abstract

@@
 bibliography: references.bib
-abstract: "`r paste0(c('English:', readLines('abstract.md'), '\\\\par Deutsch:\\\\textgerman{', readLines('zusammenfassung.md'), '}\\\\newpage'), collapse = ' ')`"
+abstract: "`r paste0(c('\\\\setstretch{1.0}English:', readLines('abstract.md'), '\\\\par Deutsch:\\\\textgerman{', readLines('zusammenfassung.md'), '}\\\\newpage'), collapse = ' ')`"
 csl: apa.csl

### 2023-01-13 manuscript.Rmd
fitpaper does not fit the included pdf

@@
 
-\includepdf[pages=-,fitpaper=true, offset=5mm 0mm,frame=false]{papers/workflow.pdf}
+\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/workflow.pdf}
 
@@
 
-\includepdf[pages=-,fitpaper=true, offset=5mm 0mm,frame=false]{papers/tutorial.pdf}
+\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/tutorial.pdf}
 
@@
 
-\includepdf[pages=-,fitpaper=true, offset=5mm 0mm,frame=false]{papers/prereg.pdf}
+\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/prereg.pdf}
 

### 2023-01-13 manuscript.Rmd
respect double sided printing * binding offset, odd/even * empty pages before and after toc

@@
   \usepackage{titling}
-  \pretitle{\begin{center}\LARGE
+  \pretitle{
      \begin{minipage}{\textwidth}
-     \includepdf{cover.pdf}
+     \includepdf[offset=5mm 0mm]{cover.pdf}
      \end{minipage}
-     \newpage}
+     \newpage
+     \clearpage%
+     \thispagestyle{empty}%
+     \addtocounter{page}{-2}%
+     \null%
+     \clearpage
+     \begin{center}\LARGE}
   \posttitle{\end{center}}
+
   \usepackage{polyglossia}
@@
 geometry: "left=3.6cm, right=3.6cm, top=4cm, bottom=4cm, bindingoffset=5mm"
+classoption: twoside
 papersize: a4
@@
 csl: apa.csl
+toc: false # it is manually placed
 repro:
@@
 \newpage
+\clearpage
+\thispagestyle{empty}
+\null
+\clearpage
+
+\tableofcontents
+
+\newpage
+\clearpage
+\thispagestyle{empty}
+\null
+\clearpage
 
@@
 
-\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/tutorial.pdf}
+\includepdf[pages=-, offset=5mm 0mm]{papers/tutorial.pdf}
 
@@
 
-\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/prereg.pdf}
+\includepdf[pages=-, offset=5mm 0mm]{papers/prereg.pdf}
 

### 2023-01-13 manuscript.Rmd
many minor fixes after reading

@@
 The extent of this bias depends on the inductive process.
-Without transparency about the inductive process, researchers are unable to judge the empirical support of a theory.
+The inductive process therefore have to be transparent to other researcher because they are otherwise unable to  judge the empirical support of a theory.
 
-The above distinction arises because some parts of the inductive process are well defined in the form of statistical methods, while researcher also engage in more informal inductive behavior.
-Statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about remaining informal sources of inductive bias.
+The above distinction, between statistical and sociological countermeasures, arises arises from two different kinds of induction.
+On the one hand, we have inductive processes that are well defined in the form of statistical methods, on the other hand researcher also engage in more informal inductive behavior outside of well defined models.
+They require different countermeasures because statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
 To understand why transparency is crucial, we must understand that formal and informal induction are integral to empirical sciences.
@@
 By the same token, whole theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
-If strictly applied to auxiliary assumptions, such strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
+If applied to auxiliary assumptions, such strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
 If pushed to the limit, we arrive at a "theory" governed by the data.
@@
 The need for transparency is closely tied to the use of induction in the empirical test of a theory.
-There has been a long and vigorous debate about what it means to test a theory empirically (xxx).
+There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002; Ch. 5 Experience as a Method].
 I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
@@
 Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
-7.2, in their derivation of the Akaike Information Criterion (XXX) in its general form.
+7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
 What is of interest here is not the derivation but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
@@
 Of course, we usually do not know $\theta_*$.
-That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ is independently sampled from $X\sim f$.
+That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ are $n$ independent samples from $X\sim f$.
 
@@
 Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$ respectively.
-For more details see of this derivation see @burnhamModelSelectionMultimodel2002, Ch. 7.2.
+For more details about this derivation see @burnhamModelSelectionMultimodel2002, Ch. 7.2.
 
 The observed likelihood $log(g(x|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-Note that $tr[J(\theta_*)I(\theta_*)^{-1}]$ is usually negative definitive and, therefore, we may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
+We may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowDeepLearning2016; @goodfellowCapacityOverfittingUnderfitting2016].
+This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016].
 I therefore denote it as as $\mathcal{C}$.
@@
 One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known (XXX).
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
 
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood:
+In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for a arbitrary loss function):
 \begin{equation}
@@
 \end{equation}
-The idea is that we choose from the multitude the version of our theory that best fits reality.
+The idea is that we choose the version of our theory that best fits reality.
 However, we are forced to rely on a limited sample of reality.
 We are misled because these two factors, induction and limited sample size, interact.
+Choosing the best version of our theory, based on a sample, is almost surely suboptimal.
+The observed loss, therefore, is an overconfident estimate of the loss in the future and closeness to truth.
 \begin{equation}

### 2023-01-16 manuscript.Rmd
single space toc
first draft of summary of reproducibility

@@
 
+\begin{spacing}{1}
 \tableofcontents
+\end{spacing}
 
@@
 The need for transparency is closely tied to the use of induction in the empirical test of a theory.
-There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002; Ch. 5 Experience as a Method].
+There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch. 5 Experience as a Method].
 I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
@@
 The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic, if it has inductive elements.
-This observation is only of limited use if we do not now the extend of this optimism.
-However, both setups show that the extent of the optimism ($\mathcal{C}$) is closely related to the inductive process and give us two starting points for making this bias transparent.
+This observation is only of limited use if we do not know the extend of this optimism.
+However, both setups show that the extent of the optimism  is closely related to the complexity of the inductive process ($\mathcal{C}$), and give us two starting points for making this bias transparent.
 The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
@@
 
+What we can do is to split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
+Formal, in the strict sense means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
+If a researcher employs a linear model, the complexity is calculable and even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
+However, if the researcher reconsiders the model based on the results, and fits a model with, e.g., a predictor added or removed, the complexity can not be formally judged because the induction is partly done by the researcher themselves.
+We can not ask the researcher to repeat the process on all possible data sets nor is the behavior mathematically well defined.
+What we know, however, is that the resulting linear model is more complex than a linear regression formally implies.
+
+$$
+\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
+$$
+
+That is not to say that we have no basis to judge the informal induction.
+We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
+Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data driven decision, which implies higher complexity.
+What one can say, however, is that this judgment is debatable and, therefore, subjective.
+
 It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
-This fact implies that for theses situations the optimism bias can not be fully quantified and that full transparency can not be archived.
+This fact implies that for theses situations the complexity and hence the optimism bias can not be fully quantified.
+The second best thing to transparency about the complexity is to communicate the uncertainty about the complexity clearly.
+However, judging the informal complexity $\mathcal{C}(\text{informal})$ must remain an subjective exercise.
+
 To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
-Otherwise, the supposedly objective test of the theory using hard data must still be judged subjectively.
+Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.
+
+To summarise, I separate complexity $\mathcal{C}$ according to the transparency that can theoretically be achieved.
+Formal inductive processes allow full transparency in the sense that $\mathcal{C}$ can be objectively quantified.
+Informal inductive processes allow only a limited transparency and must be subjectively judged.
 
@@
 In the following, I propose preregistration, as means to move induction into the formal domain, and computational reproducibility to make formal induction transparent.
+I argue that both practices are ultimately aimed at providing transparency to judge empirical support for theories, though they may have secondary benefits for science.
+
+## Transparency about statistical models: Computational Reproducibilibiltiy
+
+Assume for a moment that a researcher only engaged in formal induction while testing a theory empirically.
+If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
+In other words, having formal induction is not enough, the induction still must be made transparent.
+Such transparency can be achieved by computational reproducibility.
+Computational reproducibility is usually defined as the ability to recreate the same results from the same data set.
+However, through the lens of transparency about the inductive process, we may further refine our view on computational reproducibility.
+
+Consider a software that delivers the same results upon input of the same data, i.e., its reproducible.
+However, place two restrictions on the software.
+First, while able to compute results based on data, the software can not be understood by human researcher, e.g., because it is a binary program.
+Second, when the data set changes, the software does not work anymore.
+The first condition rules out that the software can be analyzed formally.
+The second condition prevents us from changing the data in any way and observe the results.
+We, therefore, have no means to assess the complexity of the inductive process and can not judge the empirical support.
+It follows that to provide transparency about the complexity of an inductive process the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.
+
+If we relax the second condition and have a "black box" that is nerveless repeatable on other data, we still have transparency about the inductive process, in the sense, that we can estimate the complexity of the inductive process.
+Relaxing the first condition most likely leads to relaxing the second condition as well.
+If the inductive process can be understood by humans, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
+The most minimal extension of computational reproducibility is, therefore, to require additionally that the process can be repeated on other similar data.
+
+To provide transparency, computational reproducibility must satisfy two requirements.
+First, computational reproducibility must show that the results have been the product of the proposed inductive process, i.e., the same data leads to the same results.
+Second, computational reproducibility must enable that the inductive process can be repeated on similar data.
+
+These two requirements are not easy to meet in practice.
+In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
+Removing the human from the loop enables to formally verify that the first requirement is met, and drastically reduces the burden of the second requirement.
+While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process easily, it asks a lot from the researchers who create the work.
+To address this problem, and make the application of the workflow easier I developed the R package `repro` [@R-repro].
+I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the there presented software `worcs`.
+
+To address the problem of making transparent how the data led to the results, these works divide it into four sub problems.
+First, it must be unambiguous what results are generated by which inductive process.
+This can be ensured by employing dynamic document creation.
+Second, the version of the software that is used to generate the results must be known.
+This can be ensured by documenting the version using software management.
+Third, the version of the author written code (and possibly text) must be tracked using version control systems.
+Fourth, how the data relates to software and computing infrastructure must be managed by e.g. workflow automation.
+
+Addressing these four problems through automation lowers the barrier to understand, and hence make transparent, the inductive process.
+However, it also addresses the here proposed additional requirement for computational reproducibility, that the process can be repeated on similar data, as it effectively allows to scale the process across many data sets.
+
+## Transparency about human researchers: Preregistration
+
+# Discussion
+
+## Limitations
+
+### Theorethical
+
+### Practical
+
+## Future Research
+
+<!--you must pay the reaper-->
+Accounting for how much induction was used to deductively test a statement is necessary for correct appraisal of evidential support of a hypothesis.
+Researcher (and philosophers of science) differ considerably in what appraisal entails, e.g., judging the long-term performance (if they are frequentists), degrees of believe (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
+14] of a hypothesis.
+But whatever measure they subscribe to, induction can shield shield any hypothesis from refutation by data altogether.
+
+Assume that all auxiliary assumption, regarding proper quantification of behavior and situation in valid measurements, causality, individual differences, nuisance variables, distributional, etc, hold.
+
+While not using statistical methods would lead to the rejection of almost all psychological theories, pushed to the extreme statistical methods can shield theories from refutation altogether.
+Both situations are uninformative and, therefore, rather unhelpful for the generation of knowledge.
+The later situation, however, becomes less than helpful when one is under the impression that theories are well tested, when in fact there was little chance that the data could refute the theory if wrong.
 
@@
 
-The following article is reprinted from the following source under [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).
+The following article is reprinted from the following source under the conditions of the [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).
 
@@
 The following article is prepared to be submitted to *Philosophical Psychology*.
-The preprint is reprinted here under [Creative Commons Public Domain Dedication (CC0 1.0)](https://creativecommons.org/publicdomain/zero/1.0/).
+The following preprint is reprinted from the following source under the conditions of the [Creative Commons Public Domain Dedication (CC0 1.0)](https://creativecommons.org/publicdomain/zero/1.0/).
 

### 2023-01-16 manuscript.Rmd
grammerly

@@
 
-Assume for a moment that a researcher only engaged in formal induction while testing a theory empirically.
+Assume that a researcher only engaged in formal induction while testing a theory empirically.
 If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
@@
 
-Consider a software that delivers the same results upon input of the same data, i.e., its reproducible.
+Consider software that delivers the same results upon input of the same data, i.e., it is reproducible.
 However, place two restrictions on the software.
@@
 The first condition rules out that the software can be analyzed formally.
-The second condition prevents us from changing the data in any way and observe the results.
-We, therefore, have no means to assess the complexity of the inductive process and can not judge the empirical support.
-It follows that to provide transparency about the complexity of an inductive process the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.
+The second condition prevents us from changing the data and observing the results.
+Therefore, we have no means to assess the complexity of the inductive process and can not judge the empirical support.
+It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.
 
-If we relax the second condition and have a "black box" that is nerveless repeatable on other data, we still have transparency about the inductive process, in the sense, that we can estimate the complexity of the inductive process.
+Suppose we relax the second condition and have a "black box" that is repeatable on other data.
+In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process.
 Relaxing the first condition most likely leads to relaxing the second condition as well.
-If the inductive process can be understood by humans, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
+If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
 The most minimal extension of computational reproducibility is, therefore, to require additionally that the process can be repeated on other similar data.
@@
 First, computational reproducibility must show that the results have been the product of the proposed inductive process, i.e., the same data leads to the same results.
-Second, computational reproducibility must enable that the inductive process can be repeated on similar data.
+Second, computational reproducibility must enable the inductive process to be repeated on similar data.
 
@@
 In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
-Removing the human from the loop enables to formally verify that the first requirement is met, and drastically reduces the burden of the second requirement.
-While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process easily, it asks a lot from the researchers who create the work.
-To address this problem, and make the application of the workflow easier I developed the R package `repro` [@R-repro].
+Removing the human from the loop enables researchers to verify formally that the first requirement is met and drastically reduces the burden of the second requirement.
+While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process quickly, it asks a lot from the researchers who create the work.
+To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
 I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the there presented software `worcs`.
@@
 Third, the version of the author written code (and possibly text) must be tracked using version control systems.
-Fourth, how the data relates to software and computing infrastructure must be managed by e.g. workflow automation.
+Fourth, how the data relates to software and computing infrastructure must be managed by e.g., workflow automation.
 
-Addressing these four problems through automation lowers the barrier to understand, and hence make transparent, the inductive process.
-However, it also addresses the here proposed additional requirement for computational reproducibility, that the process can be repeated on similar data, as it effectively allows to scale the process across many data sets.
+Addressing these four problems through automation lowers the barrier to understanding, making the inductive process transparent.
+However, it also addresses the proposed additional requirement for computational reproducibility, that the process can be repeated on similar data, as it effectively allows to scale the process across many data sets.
 

### 2023-01-16 manuscript.Rmd
grammerly

@@
 
-Now that we have established the need for transparency about the inductive process, we can drop a few of the more technical details to get a more straightforward about what we have to make transparent.
+Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
 It bears repeating that simply laying open what has been done is not enough.
 Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
-On a conceptual level, we want to compare:
+On a conceptual level, we want to compare the following:
 \begin{equation}
@@
 
-Where, $\mathcal{L}$ stands for loss function, i.e., how to compare predictions and reality.
-To allow for induction to happen, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
+Where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
+To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
 
@@
 We are misled because these two factors, induction and limited sample size, interact.
-Choosing the best version of our theory, based on a sample, is almost surely suboptimal.
-The observed loss, therefore, is an overconfident estimate of the loss in the future and closeness to truth.
+Choosing the best version of our theory based on a sample is almost surely suboptimal.
+Therefore, the observed loss is an overconfident estimate of the loss in the future and closeness to the truth.
 \begin{equation}
@@
 The extent of this optimism depends on the inductive process, not merely its results.
-Specifically, it depends on the complexity, i.e., the ability of the inductive process to adept to data.
+Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
 Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
@@
 
-The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic, if it has inductive elements.
-This observation is only of limited use if we do not know the extend of this optimism.
-However, both setups show that the extent of the optimism  is closely related to the complexity of the inductive process ($\mathcal{C}$), and give us two starting points for making this bias transparent.
+The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic if it has inductive elements.
+This observation is only useful if we know the extent of this optimism.
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}$) and give us two starting points for making this bias transparent.
 The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
-Both approaches require researchers to make the process of induction transparent instead of merely publishing the results.
+Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.
 
 Even a casual consideration of the above formalization should strike anyone who has ever worked with empirical data as unrealistic.
-No one can actually expect researchers to be inductive only in ways that are formally analyzable or even strictly repeatable.
-The point is to set the goal post and have a yardstick to measure how well a method aimed at improving empirical sciences does.
+No one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
+The point is to set the goalpost and have a yardstick to measure how well a method aimed at improving empirical sciences does.
 
-What we can do is to split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
-Formal, in the strict sense means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
-If a researcher employs a linear model, the complexity is calculable and even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
-However, if the researcher reconsiders the model based on the results, and fits a model with, e.g., a predictor added or removed, the complexity can not be formally judged because the induction is partly done by the researcher themselves.
-We can not ask the researcher to repeat the process on all possible data sets nor is the behavior mathematically well defined.
-What we know, however, is that the resulting linear model is more complex than a linear regression formally implies.
+However, we can split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
+Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
+If a researcher employs a linear model, the complexity is calculable.
+Even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
+However, suppose the researcher reconsiders the model based on the results and fits a model with, e.g., a predictor added or removed.
+In that case, the complexity can not be formally judged because the researcher is part of the inductive process.
+We can not ask the researcher to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
+However, we know that the resulting linear model is more complex than a linear regression formally implies.
 
@@
 
-That is not to say that we have no basis to judge the informal induction.
+That is not to say that we have no basis for judging the informal induction.
 We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
-Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data driven decision, which implies higher complexity.
+Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data-driven decision, which implies higher complexity.
 What one can say, however, is that this judgment is debatable and, therefore, subjective.
@@
 It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
-This fact implies that for theses situations the complexity and hence the optimism bias can not be fully quantified.
-The second best thing to transparency about the complexity is to communicate the uncertainty about the complexity clearly.
-However, judging the informal complexity $\mathcal{C}(\text{informal})$ must remain an subjective exercise.
+This fact implies that for these situations, the complexity and hence the optimism bias can not be fully quantified.
+The second best thing to transparency about complexity is to communicate the uncertainty about the complexity clearly.
+However, judging the informal complexity $\mathcal{C}(\text{informal})$ must remain a subjective exercise.
 
-To enable proper judgment of the whole theory, the imperative is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
+The imperative to enable proper judgment of theories is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
 Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.
@@
 Formal inductive processes allow full transparency in the sense that $\mathcal{C}$ can be objectively quantified.
-Informal inductive processes allow only a limited transparency and must be subjectively judged.
+Informal inductive processes allow only limited transparency and must be subjectively judged.
 
-We, therefore, have two bounds that limit transparency about the inductive process.
-First, some things can not be made transparent in principle because some induction happens informally, in the sense that we can not estimate the optimism with certainty.
-Second, even formalized inductive processes only allow to estimate optimism theoretically but they have to be communicated effectively to do so in practice.
+Therefore, we have two bounds that limit transparency about the inductive process.
+First, some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
+Second, even formalized inductive processes only allow estimating optimism theoretically, but they have to be communicated effectively to do so in practice.
 
-In the following, I propose preregistration, as means to move induction into the formal domain, and computational reproducibility to make formal induction transparent.
-I argue that both practices are ultimately aimed at providing transparency to judge empirical support for theories, though they may have secondary benefits for science.
+In the following, I propose preregistration as means to move induction into the formal domain and computational reproducibility to make formal induction transparent.
+I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
 

### 2023-01-18 manuscript.Rmd
update

@@
 Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Laying open the difficulties and attempting to overcome them are not a recent trends, though they have been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that also begins to ripple through other empirical sciences.
+Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that also begins to ripple through other empirical sciences.
 As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
@@
 Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would probably be no psychological theory that could be subject of an empirical test.
+If they had to, there would probably be no psychological theory that could be subjected to an empirical test.
 
@@
 Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
-Because both are quite technical, they are followed by a more conceptual summary of these tools.
+Because both frameworks are quite technical, they are followed by a more conceptual summary of these tools.
 These sections provide the basis to understand how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for a arbitrary loss function):
+In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for an arbitrary loss function):
 \begin{equation}
@@
 No one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
-The point is to set the goalpost and have a yardstick to measure how well a method aimed at improving empirical sciences does.
+The point is to set the goalpost and have a yardstick to measure how well we can judge empierical support.
+Without induction, there is no bias to correct for.
+With only formal induction, the bias can be quantified.
+The issue of how to deal with informal induction remains.
 
-However, we can split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
+As a first step, we can note that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
 Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
@@
 Such transparency can be achieved by computational reproducibility.
-Computational reproducibility is usually defined as the ability to recreate the same results from the same data set.
+Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992].
 However, through the lens of transparency about the inductive process, we may further refine our view on computational reproducibility.
@@
 
-Suppose we relax the second condition and have a "black box" that is repeatable on other data.
+Suppose we relax the second condition, then we have a "black box" that is repeatable on other data.
 In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process.
-Relaxing the first condition most likely leads to relaxing the second condition as well.
+Relaxing the first condition, intransparency to human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
@@
 
-To provide transparency, computational reproducibility must satisfy two requirements.
+We can therefore conclude that to provide transparency, computational reproducibility must satisfy two requirements.
 First, computational reproducibility must show that the results have been the product of the proposed inductive process, i.e., the same data leads to the same results.
@@
 Removing the human from the loop enables researchers to verify formally that the first requirement is met and drastically reduces the burden of the second requirement.
+Given enough computing power, automation enables to run the same analysis on many data sets.
+
 While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process quickly, it asks a lot from the researchers who create the work.
@@
 
-Addressing these four problems through automation lowers the barrier to understanding, making the inductive process transparent.
+Addressing these four problems through automation lowers the barrier for assessing the inductive process.
 However, it also addresses the proposed additional requirement for computational reproducibility, that the process can be repeated on similar data, as it effectively allows to scale the process across many data sets.
@@
 
-# Discussion
+Computational reproducibility enables transparency about formal inductive processes.
+However, informal inductive processes deprive us of full transparency.
+It is therefore prudent to replace informal induction with formal induction as much as possible.
+If a researcher makes every effort to only employ formal inductive reasoning in a reproducible manner, they still must persuade their readers of this fact.
 
-## Limitations
+The problem is that formal inductive reasoning can be part of an informal process.
+A perfectly reproducible linear regression is from the outside indistinguishable from one that was cherry picked from hundreds of possible regressions.
+However, our expectation regarding verisimilitude and future performance between the cherry picked and the simple regression should vary substantially.
 
-### Theorethical
+So what would constitute a persuasive argument that the data has not influenced the results above and beyond the complexity of the formal inductive process?
+We could simply trust the researchers.
+In fact, we would have to trust them on two accounts.
+First, we must put faith in their honesty.
+Second, we must trust their competence in deciding what to disclose how.
+This is not an easy task, because we need to know what they did, as well as what they did not do, e.g., that they did not cherry pick, and what they would have done had the data looked different.
+This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.
+To judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data looked different.
 
-### Practical
+Relying on trust for this judgment is not entirely unreasonable.
+However, while this type of trust is a requirement to judge the informal complexity, it does not distinguish a formal from an informal inductive process.
+Simply trusting that researchers know about the pitfalls of informal induction and disclose the required information is a subjective judgment.
+Relying on a subjective judgment would mean to surrender the advantage a formal inductive process provides; instead of quantification of complexity, the complexity must be judged with a considerable amount of uncertainty.
 
-## Future Research
+A clearer distinction between formal and informal can be provided by specifying the formal inductive process before any data is available.
+Such practice is called registration [@riceCurtailingUsePreregistration2019] or preregistration [@nosekPreregistrationRevolution2018]. 
+Registration is a strong argument that the data has not influenced the results more than the complexity of the formal inductive process implies.
+If the data does not yet exist, it can not influence the results.
+Note that preregistration is about the flow of information, i.e., the data used for induction, and not about the temporal order.
+Information can simply not travel backward in time, which makes registration before data acquisition so appealing.
 
-<!--you must pay the reaper-->
-Accounting for how much induction was used to deductively test a statement is necessary for correct appraisal of evidential support of a hypothesis.
-Researcher (and philosophers of science) differ considerably in what appraisal entails, e.g., judging the long-term performance (if they are frequentists), degrees of believe (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
-14] of a hypothesis.
-But whatever measure they subscribe to, induction can shield shield any hypothesis from refutation by data altogether.
-
-Assume that all auxiliary assumption, regarding proper quantification of behavior and situation in valid measurements, causality, individual differences, nuisance variables, distributional, etc, hold.
-
-While not using statistical methods would lead to the rejection of almost all psychological theories, pushed to the extreme statistical methods can shield theories from refutation altogether.
-Both situations are uninformative and, therefore, rather unhelpful for the generation of knowledge.
-The later situation, however, becomes less than helpful when one is under the impression that theories are well tested, when in fact there was little chance that the data could refute the theory if wrong.
-
-## Transparency about statistical models: Computational Reproducibilibiltiy
-
-## Transparency about human researcher: Preregistration
+Having no access to the data when specifying the analysis forces the researcher to think about the process of induction.
+They have to reason about all the possible data patterns they might encounter.
+While this makes registration so difficult, it is exactly the information we need.
+The expectation of the inductive process over all possible data is what we are trying to estimate, when we adjusting for the complexity of the inductive process.
 
@@
 
-## Limitations
+## From theorie to model
 
-### Theoretical
+## Comparing predictions to reality
 
-### Practical
+Subjectivity
 
-## Future Research
+## The issue of trust
 

### 2023-01-21 manuscript.Rmd
set sections headings as small caps
reread with different perspective

@@
   \usepackage{pdfpages}
+
   \usepackage{titling}
+  \let\paragraph\oldparagraph
+  \let\subparagraph\oldsubparagraph
+  \usepackage[sc, compact, raggedright]{titlesec}
+
   \pretitle{
@@
 Consider the example of mean difference between a placebo and an experimental group, that we test using a simple t-test of mean differences.
-The decision to reject the null hypothesis (no mean difference between groups) still depends, besides true mean difference and sample size, on the observed variance.
+The decision to reject the null hypothesis (groups have equal means) still depends, besides true mean difference and sample size, on the observed variance.
 However, the variance needs to be estimated from the data.
@@
 The need for transparency is closely tied to the use of induction in the empirical test of a theory.
-There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch. 5 Experience as a Method].
+There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5 Experience as a Method].
 I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
@@
 However, this fact does not impede us from getting closer to the truth because we can still compare two theories relative to each other.
-Because the expectation for $f$ remains constant (left-hand expectation in Eq.
-\@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.
-\@ref(eq:klexp)) to make a comparative judgment.
+Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
 To make a relative judgment about several competing theories it suffices to estimate for any theory $g(x)$:
@@
 \begin{equation}
-\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))]
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))]\label{eq:dataexp}
 \end{equation}
@@
 With slight abuse of notation, $log(g(y|\theta)) \equiv \sum_{i = 1}^n log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
-The observed likelihood is often easy enough to obtain, e.g. in maximum likelihood estimation.
+The observed likelihood is, therefore, often available, e.g. in maximum likelihood estimation.
 
@@
 \end{equation}
-This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004] and resembles the expectation over data discussed in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
+This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004].
+Note that both training ($y$) and test ($x$) data vary in this expectation.
+Therefore, it is closely related expectation over data shown in Eq.\ \@ref(eq:dataexp) in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
 
 Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
-That is, we make use of cross-validation where the data is partitioned, and the inductive process is repeated on all permutations of a subset of the partitions.
+That is, we make use of cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
@@
 \begin{equation}
-\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}
+\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}(\text{Model})
 \end{equation}
 
-Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data.
+Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data ($\mathcal{C}$).
 The extent of this optimism depends on the inductive process, not merely its results.
 Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
+The ability to adept to data is independent from the specific data that was observed.
+Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
 Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
@@
 This observation is only useful if we know the extent of this optimism.
-However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}$) and give us two starting points for making this bias transparent.
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and give us two starting points for making this bias transparent.
 The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
@@
 Even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
-However, suppose the researcher reconsiders the model based on the results and fits a model with, e.g., a predictor added or removed.
-In that case, the complexity can not be formally judged because the researcher is part of the inductive process.
+However, suppose the researcher reconsiders the model based on the results and their internal model.
+For example, they decided based on a worse than expected results and their intuition that they should add a predictor to the model and remove another.
+In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
 We can not ask the researcher to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
@@
 
-Therefore, we have two bounds that limit transparency about the inductive process.
-First, some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
-Second, even formalized inductive processes only allow estimating optimism theoretically, but they have to be communicated effectively to do so in practice.
+Therefore, we have a theoretical bound that limits transparency about the inductive process.
+Some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
+However, there is another bound on transparency that we must consider.
+That is how well is the inductive process, be it formal or informal, communicated to the intended recipient.
+To achieve transparency about the inductive process, and ultimately about the empirical support of a theory, we must increase both bounds.
 
-In the following, I propose preregistration as means to move induction into the formal domain and computational reproducibility to make formal induction transparent.
+In the following, I propose preregistration as means to move induction into the formal domain (pushing the first bound) and computational reproducibility (pushing the second bond) to make formal induction transparent.
 I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
 
-## Transparency about statistical models: Computational Reproducibilibiltiy
+## Transparency about statistical models: Computational\ Reproducibilibiltiy
 
@@
 Suppose we relax the second condition, then we have a "black box" that is repeatable on other data.
-In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process.
+In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process, i.e., using cross-validation.
 Relaxing the first condition, intransparency to human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
-The most minimal extension of computational reproducibility is, therefore, to require additionally that the process can be repeated on other similar data.
+Therefore, we must extend the traditional definition of computational reproducibility and at least also require that the process can be repeated on other similar data.
 
-We can therefore conclude that to provide transparency, computational reproducibility must satisfy two requirements.
+We can, therefore, conclude that to provide transparency, computational reproducibility must satisfy two requirements.
 First, computational reproducibility must show that the results have been the product of the proposed inductive process, i.e., the same data leads to the same results.
@@
 In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
-Removing the human from the loop enables researchers to verify formally that the first requirement is met and drastically reduces the burden of the second requirement.
-Given enough computing power, automation enables to run the same analysis on many data sets.
+Implementing computational reproducibility by automating the process from data to results, fulfills both conditions needed for transparency while reducing the effort that must be invested.
+Verifying that the results are actually produced by the inductive process and data is then a task that may be fulfilled by a computer without human intervention.
+Removing the human from the loop facilitates that the data can be easily substituted by similar data.
+Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the judgment of the complexity.
 
-While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process quickly, it asks a lot from the researchers who create the work.
+While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process easily, it asks a lot from the researchers who create the work.
 To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
@@
 
-To address the problem of making transparent how the data led to the results, these works divide it into four sub problems.
+To address the problem automating reproducibility, these works divide it into four sub problems.
 First, it must be unambiguous what results are generated by which inductive process.
@@
 It is therefore prudent to replace informal induction with formal induction as much as possible.
-If a researcher makes every effort to only employ formal inductive reasoning in a reproducible manner, they still must persuade their readers of this fact.
+If a researcher makes every effort to only employ formal inductive reasoning in a reproducible manner, they still must persuade their readers that they have not engaged in informal induction.
 
@@
 So what would constitute a persuasive argument that the data has not influenced the results above and beyond the complexity of the formal inductive process?
-We could simply trust the researchers.
-In fact, we would have to trust them on two accounts.
-First, we must put faith in their honesty.
-Second, we must trust their competence in deciding what to disclose how.
+Researchers could simply try to explain the inductive process after the fact.
 This is not an easy task, because we need to know what they did, as well as what they did not do, e.g., that they did not cherry pick, and what they would have done had the data looked different.
@@
 
-Relying on trust for this judgment is not entirely unreasonable.
-However, while this type of trust is a requirement to judge the informal complexity, it does not distinguish a formal from an informal inductive process.
-Simply trusting that researchers know about the pitfalls of informal induction and disclose the required information is a subjective judgment.
-Relying on a subjective judgment would mean to surrender the advantage a formal inductive process provides; instead of quantification of complexity, the complexity must be judged with a considerable amount of uncertainty.
+Trusting a post hoc explanation of an inductive process is not entirely unreasonable.
+However, post hoc, formal and informal induction are indistinguishable unless the account is comprehensive.
+Again, not only comprehensive for the data that was observed but also for all possible data.
+This task is very difficult to accomplish in practice.
+After all researchers would have to lay open the process that lead to the induction, i.e., their thoughts and their full mental model.
+Dissolving the distinction between formal and informal induction would mean to surrender the advantage a formal inductive process provides; instead of quantification of complexity, the complexity must be judged with a considerable amount of uncertainty.
 
-A clearer distinction between formal and informal can be provided by specifying the formal inductive process before any data is available.
+A clear distinction between formal and informal can be provided by specifying the formal inductive process before any data is available.
 Such practice is called registration [@riceCurtailingUsePreregistration2019] or preregistration [@nosekPreregistrationRevolution2018]. 
@@
 If the data does not yet exist, it can not influence the results.
-Note that preregistration is about the flow of information, i.e., the data used for induction, and not about the temporal order.
+Note that registration is about the flow of information, i.e., the data used for induction, and not about the temporal order.
 Information can simply not travel backward in time, which makes registration before data acquisition so appealing.
@@
 They have to reason about all the possible data patterns they might encounter.
-While this makes registration so difficult, it is exactly the information we need.
-The expectation of the inductive process over all possible data is what we are trying to estimate, when we adjusting for the complexity of the inductive process.
+While this makes registration so difficult, it is exactly the information we need and from which we derive the complexity of the inductive process.
+
+Note that requiring induction to be provably formal (analytically analyzable or repeatable) does not imply that induction itself is limited in its extend.
+A formal inductive process can still be designed to accommodate a wide range of data patterns, e.g., for a novel research question or a vague theory.
+There are several strategies that reflect such uncertainties.
+For example, if it is unknown which variables from a broad set of variables are involved, a variable selection mechanism can be specified or if it is unclear what functional form to expect, there exist many statistical models that allow almost arbitrary functional forms or if the theory does not suffice to define what an outlier is, this decision can made data dependent.
+
+The idea of registering the inductive process without knowledge of the data to demarque formal from informal induction suffers from two problems.
+First, how can this be accomplished in a way that is practical and fits into the well established practices of the scientific community?
+Second, what happens when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable because of other reasons?
+
+In @peikertReproducibleResearchTutorial2021 I propose a practical, yet rigorous form of preregistration, termed preregistration as code (PAC).
+In a PAC, researchers write the intended analyses as computer code, initially based on simulated data.
+They include this code in a reproducible dynamic document written in the style of a traditional academic manuscript.
+This version, with "mock" results based on simulated data serves as the registration.
+When the data is then gathered the results are updated to reflect the actual data.
+
+In XXX, I investigate the objective of preregistration and how it may be separated from the objective of confirmatory science.
+Specifically, I propose reduction of uncertainty about theoretical risk as the formal objective of preregistration.
 

### 2023-01-21 manuscript.Rmd
fix broken hyperrefs

@@
     number_sections: false
+    keep_tex:  true
 header-includes: |
@@
 
-  \usepackage{titling}
+  % below works well with hyperref, but spacing is standard (to large)
+  %\usepackage{sectsty}
+  %\sectionfont{\Large\normalfont\scshape\raggedright}
+  %\subsectionfont{\large\normalfont\scshape\raggedright}
+
+    % titlesec does not play nice with RMarkdown, so hack:
   \let\paragraph\oldparagraph
@@
   \usepackage[sc, compact, raggedright]{titlesec}
+  % titlesec does not play nice with hyperref, so hack:
+  \newcommand{\sectionbreak}{\phantomsection}
+  \newcommand{\subsectionbreak}{\phantomsection}
 
+  \usepackage{titling}
   \pretitle{

### 2023-01-22 manuscript.Rmd
increase margins

@@
 linestretch: 1.2
-geometry: "left=3.6cm, right=3.6cm, top=4cm, bottom=4cm, bindingoffset=5mm"
+# https://practicaltypography.com/page-margins.html
+# 3.81 to 5.08cm or 45-90 characters or 2-3 alphabets
+geometry: "left=4cm, right=4cm, top=3.5cm, bottom=4.5cm, bindingoffset=5mm"
 classoption: twoside

### 2023-01-22 manuscript.Rmd
decrease fontsize

--- 2023-01-21 manuscript.Rmd
@@
   \setotherlanguage[]{german}
-fontsize: 12pt
+fontsize: 11.5pt
 linestretch: 1.2
-geometry: "left=3.6cm, right=3.6cm, top=4cm, bottom=4cm, bindingoffset=5mm"
+# https://practicaltypography.com/page-margins.html
+# 3.81 to 5.08cm or 45-90 characters or 2-3 alphabets
+geometry: "left=4cm, right=4cm, top=3.5cm, bottom=4.5cm, bindingoffset=5mm"
 classoption: twoside

### 2023-01-22 manuscript.Rmd
add alternative fonts
update

@@
   \usepackage{libertine}
+  %\usepackage{ebgaramond}
+  %\usepackage{garamondlibre}
+  %\usepackage{CormorantGaramond}
+  %\usepackage{MinionPro}
   \usepackage{pdfpages}
@@
 The extent of this bias depends on the inductive process.
-The inductive process therefore have to be transparent to other researcher because they are otherwise unable to  judge the empirical support of a theory.
+The inductive process therefore has to be transparent to other researcher because they are otherwise unable to judge the empirical support of a theory.
 
-The above distinction, between statistical and sociological countermeasures, arises arises from two different kinds of induction.
+The above distinction, between statistical and sociological countermeasures, arises from two different kinds of induction.
 On the one hand, we have inductive processes that are well defined in the form of statistical methods, on the other hand researcher also engage in more informal inductive behavior outside of well defined models.
-They require different countermeasures because statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
+These different kinds of induction require different countermeasures.
+Statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
 To understand why transparency is crucial, we must understand that formal and informal induction are integral to empirical sciences.
@@
 While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
-Consider the example of mean difference between a placebo and an experimental group, that we test using a simple t-test of mean differences.
+Consider a placebo and an experimental group, where we hypothesize and test a mean difference using a simple t-test.
 The decision to reject the null hypothesis (groups have equal means) still depends, besides true mean difference and sample size, on the observed variance.
@@
 Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
-Because both frameworks are quite technical, they are followed by a more conceptual summary of these tools.
+Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
 These sections provide the basis to understand how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.
@@
 
-Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory.
-We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality.
+Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
+We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood to observe the state of the world $x$ where $f$ represents the ground truth.
@@
 \begin{equation}
-\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))]\label{eq:dataexp}
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))]\label{eq:dataexp}
 \end{equation}
@@
 
-We usually favor procedures that promise unbiased estimates for the observed distance on the data used to induce $\hat \theta$, given that their assumptions are met.
-With slight abuse of notation, $log(g(y|\theta)) \equiv \sum_{i = 1}^n log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
+We usually favor procedures to induce $\hat \theta$ that promise unbiased estimates for the observed likelihood, given that their assumptions are met.
 The observed likelihood is, therefore, often available, e.g. in maximum likelihood estimation.
+With slight abuse of notation, $\log(g(y|\theta)) \equiv \sum_{i = 1}^n \log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}\log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
 
@@
 \begin{equation}
-\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]
 \end{equation}
@@
 
-The observed likelihood $log(g(x|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
+The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
 We may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016].
+This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth, replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for an arbitrary loss function):
+In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for an arbitrary loss function):
 \begin{equation}
@@
 Some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
-However, there is another bound on transparency that we must consider.
+However, there is another bound that restricts transparency further.
 That is how well is the inductive process, be it formal or informal, communicated to the intended recipient.
@@
 
-In the following, I propose preregistration as means to move induction into the formal domain (pushing the first bound) and computational reproducibility (pushing the second bond) to make formal induction transparent.
+In the following, I propose preregistration as means to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
 I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
@@
 If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
-In other words, having formal induction is not enough, the induction still must be made transparent.
+In other words, having formal induction is not enough, the induction still must be made transparent in practice.
 Such transparency can be achieved by computational reproducibility.
@@
 In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process, i.e., using cross-validation.
-Relaxing the first condition, intransparency to human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
+Relaxing the first condition, intransparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
@@
 We can, therefore, conclude that to provide transparency, computational reproducibility must satisfy two requirements.
-First, computational reproducibility must show that the results have been the product of the proposed inductive process, i.e., the same data leads to the same results.
-Second, computational reproducibility must enable the inductive process to be repeated on similar data.
+First, computational reproducibility must ensure that the same data lead to the same results.
+Second, computational reproducibility must enable the inductive process to be repeatable on similar data.
 
@@
 
-To address the problem automating reproducibility, these works divide it into four sub problems.
+To address the problem of automating reproducibility, these works divide it into four sub problems.
 First, it must be unambiguous what results are generated by which inductive process.
@@
 Fourth, how the data relates to software and computing infrastructure must be managed by e.g., workflow automation.
-
-Addressing these four problems through automation lowers the barrier for assessing the inductive process.
-However, it also addresses the proposed additional requirement for computational reproducibility, that the process can be repeated on similar data, as it effectively allows to scale the process across many data sets.
 
@@
 This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.
-To judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data looked different.
+To judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data look different.
 
@@
 They have to reason about all the possible data patterns they might encounter.
-While this makes registration so difficult, it is exactly the information we need and from which we derive the complexity of the inductive process.
+While this makes registration so difficult, it is exactly the information we need and from which we may derive the complexity of the inductive process.
 
 Note that requiring induction to be provably formal (analytically analyzable or repeatable) does not imply that induction itself is limited in its extend.
-A formal inductive process can still be designed to accommodate a wide range of data patterns, e.g., for a novel research question or a vague theory.
+In other words, researchers still might be very unsure how the data will look and therefore rely on induction.
+However, the must be clear what they are going to induce.
+A formal inductive process can be designed to accommodate a wide range of data patterns, e.g., for a novel research question or a vague theory.
 There are several strategies that reflect such uncertainties.
@@
 First, how can this be accomplished in a way that is practical and fits into the well established practices of the scientific community?
-Second, what happens when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable because of other reasons?
+Second, how to account for informal induction, e.g., when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable because of other reasons?
 
-In @peikertReproducibleResearchTutorial2021 I propose a practical, yet rigorous form of preregistration, termed preregistration as code (PAC).
+In @peikertReproducibleResearchTutorial2021 I address the first problem and propose a practical, yet rigorous form of registration, termed preregistration as code (PAC).
 In a PAC, researchers write the intended analyses as computer code, initially based on simulated data.
@@
 This version, with "mock" results based on simulated data serves as the registration.
-When the data is then gathered the results are updated to reflect the actual data.
+When the data is gathered the results are updated to reflect the actual data.
 
-In XXX, I investigate the objective of preregistration and how it may be separated from the objective of confirmatory science.
-Specifically, I propose reduction of uncertainty about theoretical risk as the formal objective of preregistration.
+To answer the second question of how to account for informal induction, I investigate the objective of registration and how it may be separated from the objective of confirmatory science in XXX.
+Particularly, I propose to formalize the objective of registration as reduction of uncertainty about theoretical risk.
+Being able to account for uncertainty is indispensable for the question of how to deal with informal induction.
+If induction can not be repeated at will nor can it be formally analyzed, we are uncertain about the complexity of the inductive process.
+Theoretical risk is conceptually related to the here discussed concept of complexity but put into the context of Bayesian philosophy of science.
+Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory, if the evidence is more likely under the theory).
+For this class of loss functions, I show how to account for uncertainty caused by informal induction and that a reduction in uncertainty due to preregistration is universally beneficial.
 
 # Discussion
+
+This dissertation proposes transparency about the inductive process as an indispensable property of empirical sciences.
+First, I address the question of *what* has to be made transparent.
+The need for transparency is directly related to the use of induction in empirical sciences.
+Induction is often necessary to test vague theories empirically.
+However, induction introduces a bias that leads to overconfidence.
+The extent of this overconfidence is a function of the inductive process.
+Specifically, I propose that transparency must concern the extend to which data may influence the results and term this complexity.
+I show that the complexity of an inductive process can be quantified, if the process can be analyzed or repeated on similar data.
+I call such an inductive process "formal".
+However, researchers often engage in informal induction, where complexity must be judged subjectively and with some uncertainty.
+Second, I explain *how* computational reproducibility and preregistration provide transparency about complexity.
+Preregistration allows to distinguish formal from informal induction, while computational reproducibility communicates the process of formal induction making it transparent.
+
+This dissertation advances the theory and application of registration and computational reproducibility.
+Based on theoretical consideration, I refined the requirements of computational reproducibility, so that not only the same data must produce the same results but also that the inductive process may be repeated on other similar data.
+I developed a workflow for computational reproducibility that satisfies these requirements and build tooling to apply it.
+Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration and develop the reduction of uncertainty about the theoretical risk as a formal objective of registration.
 
@@
 
+## Summary
+
 # Articles
 
+In the following, I reprint the articles published as part of the dissertation as they were made accessible by the journals, as allowed by the respective licences.
+Note, that the following articles were written collaboratively with the indicated coauthors.
+The annex to § 6, para.
+2 of the Doctoral Degree Regulations of the Faculty of Life Sciences, amended on 05.03.2015, *University Gazette of Humboldt-Universität zu Berlin 12* will be submitted with initiation of the doctoral degree procedure.
+
+\newpage
+
 ## A Reproducible Data Analysis Workflow With R Markdown, Git, Make, and Docker
+
+\phantomsection
 
@@
 Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow With
-R Markdown, Git, Make, and Docker. Quantitative and Computational Methods in Behavioral Sciences, 1, e3763. https://doi.org/10.5964/qcmb.3763
+R Markdown, Git, Make, and Docker. *Quantitative and Computational Methods in Behavioral Sciences*, *1*, e3763. https://doi.org/10.5964/qcmb.3763
 
@@
 ## Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once
+
+\phantomsection
 
@@
 
+\phantomsection
+
 The following article is prepared to be submitted to *Philosophical Psychology*.

### 2023-01-22 manuscript.Rmd
remove phantomsections, not needed anymore
expand discussion

@@
   % titlesec does not play nice with hyperref, so hack:
-  \newcommand{\sectionbreak}{\phantomsection}
-  \newcommand{\subsectionbreak}{\phantomsection}
+  \newcommand{\sectionbreak}{}
+  \newcommand{\subsectionbreak}{}
 
@@
 
+My consideration of transparency centers around the use of induction to test imprecise theories.
+A theory is imprecise, in the sense that it is undetermined and must use data to make testable statements.
+I conceptualize induction to derive testable statements as choosing among a set of implications of the theory.
+Such a set of possible implications and the process of choosing among them according to a loss function is what I understand as a model.
+However, this conceptualization may be criticized.
+It is not entirely clear how to chose the process of selecting the best fitting version of the theory.
+Should the process of induction depend on the theory?
+Is the application of mathematically or computationally convenient methods, as is usual in statistical modeling, sufficient?
+Must we account for the arbitrariness of the selection process?
+
+Implicit in the conceptualization of complexity was the assumption of unbiasedness (the inductive process converges towards $\theta_*$), i.e., an ideal inductive process.
+Such assumptions implies a perfect match of the set of implications of the theories and the statistical method as well as optimality of the statistical method.
+Such optimality requirement is difficult to verify.
+Most proofs regarding optimality make restrictive assumptions, i.e., optimality depends on the data, and rely on large sample behavior.
+
+In practice, researchers are restricted by the statistical models that are available to them.
+A certain statistical method may imply a set of possible implications that only partially overlaps with the theory.
+Developing more powerful inductive algorithms is a very active field of research in the machine learning community.
+However, having more and more powerful tools for induction simply means that more and more vague theories may be tested.
+It does not necessarily lead to the development of statistical methods that fit the theories researchers are interested in.
+
+It might happen that in the end, more than one inductive process may be applicable and no criterium makes one preferable over the other, i.e., from the point of the researcher the decision is arbitrary.
+We have dealed with arbitrary assumption by choosing an inductive process and forming the expectation over all possible data.
+A possible direction for future research is to explore if forming the expectation over inductive processes is similar fruitful.
+One suggestion in such a direction is a method called multiverse analysis [@steegenIncreasingTransparencyMultiverse2016], where many different data analytic decisions are systematically explored.
+
 ## Comparing predictions to reality
 
-Subjectivity
+To complicate matters further, any statistical model is optimizing some specific loss function.
+However, the choice of loss function may be highly consequential for the appraisal of a theory.
+Not only is this choice consequential but it is easily imaginable that researchers can not agree on a metric to judge the theory.
+
+One important example of loss functions that are necessarily subjective are a priori believes.
+So even researcher who agree on a general class of metriken, e.g., posterior probability of a theory, would appraise theories differently.
+Allowing loss functions to vary between researchers also implies that the complexity of the inductive process varies.
+It follows that the correction for complexity can not be done by the authors for all their readers universally, if the loss function and complexity may depend on the researcher assessing the theory.
+This again highlights the importance of making the process transparent.
+Because if the process is transparent, the readers might change the loss function and calculate complexity accordingly.
 
@@
 
-\phantomsection
-
 The following article is reprinted from the following source under the conditions of the [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).
@@
 ## Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once
-
-\phantomsection
 
@@
 
-\phantomsection
-
 The following article is prepared to be submitted to *Philosophical Psychology*.

### 2023-01-24 manuscript.Rmd
update

@@
   bookdown::pdf_document2:
-    latex_engine: "xelatex"
+    latex_engine: "lualatex"
     toc: true
@@
 
+  %https://arthur.maheo.net/beautifying-latex-paper-with-practical-typography/
+  \SetTracking{
+    encoding={OT1,T1,T2A,LY1,OT4,QX,T5,TS1,EU1,EU2,TU}, }{1000}
+
   \usepackage{titling}
@@
   \setotherlanguage[]{german}
-fontsize: 11.5pt
+fontsize: 11pt
 linestretch: 1.2

### 2023-01-24 manuscript.Rmd

@@
   \newcommand{\subsectionbreak}{}
-
-  %https://arthur.maheo.net/beautifying-latex-paper-with-practical-typography/
-  \SetTracking{
-    encoding={OT1,T1,T2A,LY1,OT4,QX,T5,TS1,EU1,EU2,TU}, }{1000}
 
@@
 \thispagestyle{empty}
-\null
+\begin{center}
+\vspace*{\fill}
+For my mother.\vspace*{\fill}
+\end{center}
+\clearpage
+
+\newpage
+\clearpage
+\thispagestyle{empty}
+# Acknowledgments {.unlisted .unnumbered}
+
+No one appreciated yet.
+\clearpage
+\thispagestyle{empty}
+\begin{center}
+\vspace*{\fill}
+Insight must precede application.
+\vspace*{\fill}
+\end{center}
+\newpage
 \clearpage
@@
 In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory.
-Testing a theory empirically is often viewed as deductive, since the theory is making statements about the data.
+Testing a theory empirically is often viewed as deductive since the theory is making statements about the data.
 Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
-I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results, if unaccounted.
+I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results if unaccounted for.
 The extent of this bias depends on the inductive process.
-The inductive process therefore has to be transparent to other researcher because they are otherwise unable to judge the empirical support of a theory.
+The inductive process, therefore, must be transparent to other researchers because they are otherwise unable to judge the empirical support of a theory.
 
 The above distinction, between statistical and sociological countermeasures, arises from two different kinds of induction.
-On the one hand, we have inductive processes that are well defined in the form of statistical methods, on the other hand researcher also engage in more informal inductive behavior outside of well defined models.
+On the one hand, we have inductive processes that are well-defined in the form of statistical methods; on the other hand, researchers also engage in more informal inductive behavior outside of well-defined models.
 These different kinds of induction require different countermeasures.
-Statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
+Statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
 To understand why transparency is crucial, we must understand that formal and informal induction are integral to empirical sciences.
 Any science must be able to communicate how it generates its knowledge.
-However, transparency has an outstanding role in psychology and other empirical sciences, because empirical statements lose their value without transparency about the inductive process that contributed to them.
+However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive process that contributed to them.
 Therefore, transparency is more than a virtue that may improve knowledge gains in empirical sciences but is an indispensable property.
@@
 While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
+Making inferences about a mean difference requires knowledge about the variance.
+Either the variance is known, or it has to be estimated from data.
 Consider a placebo and an experimental group, where we hypothesize and test a mean difference using a simple t-test.
-The decision to reject the null hypothesis (groups have equal means) still depends, besides true mean difference and sample size, on the observed variance.
-However, the variance needs to be estimated from the data.
+The decision to reject the null hypothesis (groups have equal means) depends, besides true mean difference and sample size, on the observed variance.
+However, the variance needs to be induced from the data.
 So the threshold of the deductive decision depends on a quantity that must be induced.
@@
 Induction is necessary and, in the present case, innocuous for psychology as a science.
-Here, it is innocuous because the introduced bias can be accounted for and vanishes with increasing sample size, but we will see later that this is not always the case.
-A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
+In this case, it is innocuous because the bias from inducing the variance can be accounted for.
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
 z≈`r round(qnorm(.99), 2)`).
@@
 Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would probably be no psychological theory that could be subjected to an empirical test.
+If they had to, there would be no psychological theory that could be subjected to an empirical test.
 
@@
 In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
-By inducing some quantities, psychological researchers can remove them from the conjunction.
-If researchers use induction for some necessary but under the theory arbitrary assumptions, their theory will not be refuted because of these assumptions.
-Since empirical researchers often cannot derive every assumption from their theory, avoiding refutation because of those assumptions is a desirable property.
+Auxiliary assumptions are necessary to test a theory but do not follow from the theory itself.
+If such assumptions do not hold, they may lead to empirical falsification even though the theory holds.
+However, if an auxiliary assumption is induced, it can not be falsified by the same data that induced it.
+Inducing such assumptions, therefore, effectively removes auxiliary assumptions from the conjunction that is exposed to falsification.
+Since empirical researchers cannot always derive every assumption from their theory, avoiding refutation because of those assumptions is a desirable property.
 
@@
 By the same token, whole theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
-If applied to auxiliary assumptions, such strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
+If applied to auxiliary assumptions, such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
 If pushed to the limit, we arrive at a "theory" governed by the data.
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction can not refute these statements, so they they never have been subject of an empierical test.
+The data used for induction can not refute these statements, so they have never been subject to an empirical test.
 
@@
 On the other hand, induction may render any test of a hypothesis ineffective.
-The problem, I argue, therefore, is not induction but making transparent where and to what extend induction is used in the inferential process.
-The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims as reported in the literature actually are.
+Therefore, I argue the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
+The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
 Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?
@@
 The first question (the *what*) is theoretical in nature.
-It is addressed in the synopses, which supplies the theoretical framework of the articles written as part of the dissertation.
+It is addressed in the synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
 Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).
@@
 I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches
+The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches.
 
@@
 There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5 Experience as a Method].
-I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
-The first framework motivates transparency when the aim of an empirical test is to evaluate the verisimilitude ("closeness to truth") of a theory.
+I do not attempt to rehash the debate about what constitutes an empirical test.
+However, I aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
+The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
 The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
@@
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
-These sections provide the basis to understand how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.
+These sections provide the basis for understanding how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.
 
@@
 We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
-Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood to observe the state of the world $x$ where $f$ represents the ground truth.
+Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
 We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
@@
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
-What is of interest here is not the derivation but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
+What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
 
@@
 $\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
-However, this fact does not impede us from getting closer to the truth because we can still compare two theories relative to each other.
+However, this fact does not impede us from getting closer to the truth because we can still compare the two theories relative to each other.
 Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
-To make a relative judgment about several competing theories it suffices to estimate for any theory $g(x)$:
+To make a relative judgment about several competing theories, it suffices to estimate for any theory $g(x)$:
 \begin{equation}
@@
 
-To allow for quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
+To allow quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
 That means our theory implies a family of possible probability distributions that may describe reality.
@@
 
-The inference goal is for comparing other theories to $g(x)$ is, therefore:
+The inference goal for comparing other theories to $g(x)$ is, therefore:
 \begin{equation}
@@
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
-It follows, almost surely, that we loose information:
+It follows, almost surely, that we lose information:
 \begin{equation}
@@
 There are two ways to get at this expectation.
-One is the use of Taylor series expansion, which follows in this section, and another is the use of cross-validation discussed in the next section.
+One is the use of Taylor series expansion, which follows in this section, and another is cross-validation, discussed in the next section.
 
 We usually favor procedures to induce $\hat \theta$ that promise unbiased estimates for the observed likelihood, given that their assumptions are met.
-The observed likelihood is, therefore, often available, e.g. in maximum likelihood estimation.
+The observed likelihood is, therefore, often available, e.g., in maximum likelihood estimation.
 With slight abuse of notation, $\log(g(y|\theta)) \equiv \sum_{i = 1}^n \log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}\log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
@@
 
-Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$ respectively.
-For more details about this derivation see @burnhamModelSelectionMultimodel2002, Ch. 7.2.
+Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$, respectively.
+For more details about this derivation, see @burnhamModelSelectionMultimodel2002, Ch. 7.2.
 
 The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-We may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
+We may conclude that substituting deduced quantities with induced estimates leads to some overconfidence about how close one is to the truth.
 This overconfidence is directly related to how much induction a model entails.
@@
 Instead of verisimilitude, one might be concerned with future performance.
-That is, how well does a theory do in predicting novel facts.
-Please note that the information-theoretic setup above has made no appeal to the expected performance on unseen data.
+That is, how well does a theory do in predicting novel facts?
+Please note that the information-theoretic setup above has not appealed to the expected performance on unseen data.
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for an arbitrary loss function):
+In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
 \begin{equation}
@@
 \mathbb{E}_y L(y, g(y|\hat\theta(y)))
- = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\theta))
+ = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\hat\theta(y)))
 \end{equation}
@@
 Note that both training ($y$) and test ($x$) data vary in this expectation.
-Therefore, it is closely related expectation over data shown in Eq.\ \@ref(eq:dataexp) in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
+Therefore, it is closely related to the expectation over data shown in Eq.\ \@ref(eq:dataexp) in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].
 
 Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
-That is, we make use of cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
+That is, we use cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
@@
 
-As stated earlier, using cross-validation it is possible to  estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
 To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
 \begin{equation}
-\mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta))
+\mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta(y)))
 \end{equation}
@@
 Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
-The ability to adept to data is independent from the specific data that was observed.
-Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
+The ability to adapt to data is independent of the specific data that was observed.
+Complexity is a function of the model, i.e., $\mathcal{C}(\text{model})$, not of the data.
 Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
@@
 This observation is only useful if we know the extent of this optimism.
-However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and give us two starting points for making this bias transparent.
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{model})$) and give us two starting points for making this bias transparent.
 The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
@@
 No one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
-The point is to set the goalpost and have a yardstick to measure how well we can judge empierical support.
+The point is to set the goalpost and have a yardstick to measure how well we can judge empirical support.
 Without induction, there is no bias to correct for.
@@
 However, suppose the researcher reconsiders the model based on the results and their internal model.
-For example, they decided based on a worse than expected results and their intuition that they should add a predictor to the model and remove another.
+For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
 In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
@@
 It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
-This fact implies that for these situations, the complexity and hence the optimism bias can not be fully quantified.
-The second best thing to transparency about complexity is to communicate the uncertainty about the complexity clearly.
+This fact implies that for these situations, the complexity and, hence, the optimism bias can not be fully quantified.
+Though full transparency remains out of reach, researchers may do their best to provide a good basis for assessing complexity.
 However, judging the informal complexity $\mathcal{C}(\text{informal})$ must remain a subjective exercise.
 
-The imperative to enable proper judgment of theories is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
+The imperative to enable proper judgment of theories is simple: induce only what is necessary and what you induce should, if possible, be done formally.
 Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.
@@
 However, there is another bound that restricts transparency further.
-That is how well is the inductive process, be it formal or informal, communicated to the intended recipient.
-To achieve transparency about the inductive process, and ultimately about the empirical support of a theory, we must increase both bounds.
+That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
+To achieve transparency about the inductive process and ultimately about the empirical support of a theory, we must increase both bounds.
 
-In the following, I propose preregistration as means to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
+In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
 I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
@@
 If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
-In other words, having formal induction is not enough, the induction still must be made transparent in practice.
+In other words, more than having formal induction is needed, the induction still must be made transparent in practice.
 Such transparency can be achieved by computational reproducibility.
@@
 However, place two restrictions on the software.
-First, while able to compute results based on data, the software can not be understood by human researcher, e.g., because it is a binary program.
+First, while able to compute results based on data, the software can not be understood by human researchers, e.g., because it is a binary program.
 Second, when the data set changes, the software does not work anymore.
@@
 
-Suppose we relax the second condition, then we have a "black box" that is repeatable on other data.
-In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process, i.e., using cross-validation.
-Relaxing the first condition, intransparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
+Suppose we relax the second condition; then we have a "black box" that is repeatable on other data.
+In that case, we still have transparency about the inductive process because we can estimate the complexity of the inductive process, i.e., using cross-validation.
+Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
@@
 
-We can, therefore, conclude that to provide transparency, computational reproducibility must satisfy two requirements.
+Therefore, to provide transparency, computational reproducibility must satisfy two requirements.
 First, computational reproducibility must ensure that the same data lead to the same results.
-Second, computational reproducibility must enable the inductive process to be repeatable on similar data.
+Second, computational reproducibility must make the inductive process repeatable on similar data.
 
@@
 In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
-Implementing computational reproducibility by automating the process from data to results, fulfills both conditions needed for transparency while reducing the effort that must be invested.
-Verifying that the results are actually produced by the inductive process and data is then a task that may be fulfilled by a computer without human intervention.
+Implementing computational reproducibility by automating the process from data to results fulfills both conditions needed for transparency while reducing the effort that must be invested.
+Verifying that the results are actually produced by the inductive process and data is then a task that a computer may fulfil without human intervention.
 Removing the human from the loop facilitates that the data can be easily substituted by similar data.
@@
 
-To address the problem of automating reproducibility, these works divide it into four sub problems.
+These works address the problem of automating reproducibility by dividing it into four subproblems.
 First, it must be unambiguous what results are generated by which inductive process.
@@
 This can be ensured by documenting the version using software management.
-Third, the version of the author written code (and possibly text) must be tracked using version control systems.
-Fourth, how the data relates to software and computing infrastructure must be managed by e.g., workflow automation.
+Third, the version of the author's written code (and possibly text) must be tracked using version control systems.
+Fourth, how the data relates to software and computing infrastructure must be managed by, e.g., workflow automation.
 
@@
 Computational reproducibility enables transparency about formal inductive processes.
-However, informal inductive processes deprive us of full transparency.
-It is therefore prudent to replace informal induction with formal induction as much as possible.
-If a researcher makes every effort to only employ formal inductive reasoning in a reproducible manner, they still must persuade their readers that they have not engaged in informal induction.
+However, informal inductive processes deprive us of complete transparency.
+Therefore, it is prudent to replace informal induction with formal induction as much as possible.
+If a researcher makes every effort only to employ formal inductive reasoning in a reproducible manner, they still must persuade their readers that they have not engaged in informal induction.
 
 The problem is that formal inductive reasoning can be part of an informal process.
-A perfectly reproducible linear regression is from the outside indistinguishable from one that was cherry picked from hundreds of possible regressions.
-However, our expectation regarding verisimilitude and future performance between the cherry picked and the simple regression should vary substantially.
+A perfectly reproducible linear regression is, from the outside, indistinguishable from one that was cherry-picked from hundreds of possible regressions.
+However, our expectations regarding verisimilitude and future performance between the cherry-picked and the simple regression should vary substantially.
 
@@
 Researchers could simply try to explain the inductive process after the fact.
-This is not an easy task, because we need to know what they did, as well as what they did not do, e.g., that they did not cherry pick, and what they would have done had the data looked different.
+This task is not easy because we need to know what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
 This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.
-To judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data look different.
+For researchers to judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data look different.
 
@@
 Again, not only comprehensive for the data that was observed but also for all possible data.
-This task is very difficult to accomplish in practice.
-After all researchers would have to lay open the process that lead to the induction, i.e., their thoughts and their full mental model.
-Dissolving the distinction between formal and informal induction would mean to surrender the advantage a formal inductive process provides; instead of quantification of complexity, the complexity must be judged with a considerable amount of uncertainty.
+This task is challenging to accomplish in practice.
+After all, researchers would have to lay open the process that led to the induction, i.e., their thoughts and full mental model.
+Dissolving the distinction between formal and informal induction would mean surrendering the advantage a formal inductive process provides; instead of quantifying complexity, the complexity must be judged with considerable uncertainty.
 
@@
 If the data does not yet exist, it can not influence the results.
-Note that registration is about the flow of information, i.e., the data used for induction, and not about the temporal order.
+Note that registration is about the flow of information, i.e., the data used for induction, and not the temporal order.
 Information can simply not travel backward in time, which makes registration before data acquisition so appealing.
@@
 
-Note that requiring induction to be provably formal (analytically analyzable or repeatable) does not imply that induction itself is limited in its extend.
-In other words, researchers still might be very unsure how the data will look and therefore rely on induction.
-However, the must be clear what they are going to induce.
+Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extend.
+In other words, researchers still might be very unsure how the data will look and therefore rely on induction to fill in the blanks.
+However, they must be clear about what they are going to induce.
 A formal inductive process can be designed to accommodate a wide range of data patterns, e.g., for a novel research question or a vague theory.
-There are several strategies that reflect such uncertainties.
-For example, if it is unknown which variables from a broad set of variables are involved, a variable selection mechanism can be specified or if it is unclear what functional form to expect, there exist many statistical models that allow almost arbitrary functional forms or if the theory does not suffice to define what an outlier is, this decision can made data dependent.
+Several strategies reflect such uncertainties.
+For example, suppose it is unknown which variables from a broad set of variables are involved.
+In that case, a variable selection mechanism can be specified.
+Or suppose it is unclear what functional form to expect.
+In that case, many statistical models exist that allow almost arbitrary functional forms.
+If the theory does not suffice to define an outlier, this decision can be made data-dependent, etc.
 
-The idea of registering the inductive process without knowledge of the data to demarque formal from informal induction suffers from two problems.
-First, how can this be accomplished in a way that is practical and fits into the well established practices of the scientific community?
-Second, how to account for informal induction, e.g., when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable because of other reasons?
+The idea of registering the inductive process without knowledge of the data to demarcate formal from informal induction suffers from two problems.
+First, how can this be accomplished in a way that is practical and fits into the well-established practices of the scientific community?
+Second, how to account for informal induction, e.g., when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable for other reasons?
 
-In @peikertReproducibleResearchTutorial2021 I address the first problem and propose a practical, yet rigorous form of registration, termed preregistration as code (PAC).
+In @peikertReproducibleResearchTutorial2021, I address the first problem and propose a practical yet rigorous form of registration called preregistration as code (PAC).
 In a PAC, researchers write the intended analyses as computer code, initially based on simulated data.
 They include this code in a reproducible dynamic document written in the style of a traditional academic manuscript.
-This version, with "mock" results based on simulated data serves as the registration.
-When the data is gathered the results are updated to reflect the actual data.
+This version, with "mock" results based on simulated data, serves as the registration.
+When the data is gathered, the results are updated to reflect the actual data.
 
@@
 Being able to account for uncertainty is indispensable for the question of how to deal with informal induction.
-If induction can not be repeated at will nor can it be formally analyzed, we are uncertain about the complexity of the inductive process.
+If induction can not be repeated at will nor formally analyzed, we are uncertain about the complexity of the inductive process.
 Theoretical risk is conceptually related to the here discussed concept of complexity but put into the context of Bayesian philosophy of science.
-Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory, if the evidence is more likely under the theory).
-For this class of loss functions, I show how to account for uncertainty caused by informal induction and that a reduction in uncertainty due to preregistration is universally beneficial.
+Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence, and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory if the evidence is more likely under the theory).
+For this class of loss functions, I show how to account for uncertainty caused by informal induction and that a reduction in uncertainty due to preregistration is universally beneficial for those loss functions.
 
@@
 The extent of this overconfidence is a function of the inductive process.
-Specifically, I propose that transparency must concern the extend to which data may influence the results and term this complexity.
-I show that the complexity of an inductive process can be quantified, if the process can be analyzed or repeated on similar data.
+Specifically, I propose that transparency must concern the extent to which data may influence the results and term this complexity.
+I show that the complexity of an inductive process can be quantified if the process can be analyzed or repeated on similar data.
 I call such an inductive process "formal".
@@
 Second, I explain *how* computational reproducibility and preregistration provide transparency about complexity.
-Preregistration allows to distinguish formal from informal induction, while computational reproducibility communicates the process of formal induction making it transparent.
+Preregistration allows to distinguish between formal and informal induction, while computational reproducibility communicates the process of formal induction.
 
@@
 Based on theoretical consideration, I refined the requirements of computational reproducibility, so that not only the same data must produce the same results but also that the inductive process may be repeated on other similar data.
-I developed a workflow for computational reproducibility that satisfies these requirements and build tooling to apply it.
-Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration and develop the reduction of uncertainty about the theoretical risk as a formal objective of registration.
+I developed a workflow for computational reproducibility that satisfies these requirements and built tooling to apply it.
+Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration and developed the reduction of uncertainty about the theoretical risk as a formal objective of registration.
 
-## From theorie to model
+## From theory to model
 
 My consideration of transparency centers around the use of induction to test imprecise theories.
-A theory is imprecise, in the sense that it is undetermined and must use data to make testable statements.
+A theory is imprecise because it must be refined using data to make testable statements.
 I conceptualize induction to derive testable statements as choosing among a set of implications of the theory.
@@
 Implicit in the conceptualization of complexity was the assumption of unbiasedness (the inductive process converges towards $\theta_*$), i.e., an ideal inductive process.
-Such assumptions implies a perfect match of the set of implications of the theories and the statistical method as well as optimality of the statistical method.
+Such an assumption implies a perfect match of the set of implications of the theories and the statistical method.
+Furthermore, it assumes a certain optimality of the statistical method.
 Such optimality requirement is difficult to verify.
-Most proofs regarding optimality make restrictive assumptions, i.e., optimality depends on the data, and rely on large sample behavior.
+Most proofs regarding optimality make restrictive assumptions, i.e., optimality depends on the data and relies on large sample behavior.
 
@@
 
-It might happen that in the end, more than one inductive process may be applicable and no criterium makes one preferable over the other, i.e., from the point of the researcher the decision is arbitrary.
-We have dealed with arbitrary assumption by choosing an inductive process and forming the expectation over all possible data.
-A possible direction for future research is to explore if forming the expectation over inductive processes is similar fruitful.
+It might happen that in the end, more than one inductive process may be applicable, and no criterium makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
+We have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
+A possible direction for future research is to explore if forming the expectation over inductive processes is similarly fruitful.
 One suggestion in such a direction is a method called multiverse analysis [@steegenIncreasingTransparencyMultiverse2016], where many different data analytic decisions are systematically explored.
@@
 
-To complicate matters further, any statistical model is optimizing some specific loss function.
+To complicate matters further, any statistical model optimizes some specific loss function.
 However, the choice of loss function may be highly consequential for the appraisal of a theory.
-Not only is this choice consequential but it is easily imaginable that researchers can not agree on a metric to judge the theory.
+Not only is this choice consequential, but it is easily imaginable that researchers can not agree on a metric to judge the theory.
 
-One important example of loss functions that are necessarily subjective are a priori believes.
-So even researcher who agree on a general class of metriken, e.g., posterior probability of a theory, would appraise theories differently.
+One important example of loss functions that are necessarily subjective are prior beliefs.
+So even researchers who agree on a general class of metrics, e.g., the posterior probability of a theory, would appraise theories differently.
 Allowing loss functions to vary between researchers also implies that the complexity of the inductive process varies.
-It follows that the correction for complexity can not be done by the authors for all their readers universally, if the loss function and complexity may depend on the researcher assessing the theory.
+It follows that the correction for complexity can not be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
 This again highlights the importance of making the process transparent.
-Because if the process is transparent, the readers might change the loss function and calculate complexity accordingly.
+If the process is transparent, the readers might change the loss function and calculate complexity accordingly.
 
-## The issue of trust
+## Future Research
 
-## Summary
+This dissertation is concerned with a narrow definition of transparency.
+I have not addressed essential parts of the scientific process, among them:
+How should researchers decide which theory to test?
+How should they go about gathering data concerning the chosen theory?
+When they are ready to publish their conclusions, how should these conclusions be evaluated?
+Which conclusions warrant publication?
+Of those, which warrants widespread attention of the scientific community?
+Which must be further substantiated?
+
+I expect that the conceptual frameworks I have employed to address the question of transparency about the inductive process may be enormously fruitful to apply there as well.
+Applying statistical theory to the philosophy of science has a long tradition.
+However, I hope we might abandon the tradition of analyzing scientific practices only after they have been long established.
+Instead we should analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
+That way, theoretical insight may shape application.
 
@@
 
-In the following, I reprint the articles published as part of the dissertation as they were made accessible by the journals, as allowed by the respective licences.
-Note, that the following articles were written collaboratively with the indicated coauthors.
+In the following, I reprinted the articles published as part of the dissertation as they were made accessible by the journals, as allowed by the respective licenses.
+Note that the following articles were written collaboratively with the indicated coauthors.
 The annex to § 6, para.
-2 of the Doctoral Degree Regulations of the Faculty of Life Sciences, amended on 05.03.2015, *University Gazette of Humboldt-Universität zu Berlin 12* will be submitted with initiation of the doctoral degree procedure.
+2 of the Doctoral Degree Regulations of the Faculty of Life Sciences, amended on 05.03.2015, *University Gazette of Humboldt-Universität zu Berlin 12* will be submitted with the initiation of the doctoral degree procedure.
 

### 2023-01-24 manuscript.Rmd
move citation

@@
 Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that also begins to ripple through other empirical sciences.
+Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.

### 2023-01-27 manuscript.Rmd
fix toc bug again (partial reverse of  bb79ad90f58eec9103956a26452b86171989bb5c)
it - its sounds bit doubled to me

@@
   % titlesec does not play nice with hyperref, so hack:
-  \newcommand{\sectionbreak}{}
-  \newcommand{\subsectionbreak}{}
+  \newcommand{\sectionbreak}{\phantomsection}
+  \newcommand{\subsectionbreak}{\phantomsection}
 
@@
 Statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
-To understand why transparency is crucial, we must understand that formal and informal induction are integral to empirical sciences.
-Any science must be able to communicate how it generates its knowledge.
+To understand why transparency is crucial, it is important to understand that formal and informal induction are integral to empirical sciences.
+Any science must be able to communicate how its knowledge is generated.
 However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive process that contributed to them.

### 2023-02-01 manuscript.Rmd
I argue that the problem is not
added "that"
I argue that transparency in statistical methods is enabled
changed about to in (line 192)

@@
 On the other hand, induction may render any test of a hypothesis ineffective.
-Therefore, I argue the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
+Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
 The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
@@
 Based on this distinction, the articles answer the second question (the *how*).
-I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
+I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
 The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches.

### 2023-02-01 manuscript.Rmd
preregistration in regard of transparency
changed as means of for in regard of

@@
 I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches.
+The conceptualization of computational reproducibility and preregistration in regard of transparency is supplemented by practical guidance on how researchers may implement these approaches.
 

### 2023-02-01 manuscript.Rmd
no big fan of the phrase "to lay open" in general
for example in line 200 one could also use something like "emphasize"/"highlight"/"point out"
lay open sound to die like more of a German phrase that has been translated
changed "motivates" for "promotes"

@@
 I do not attempt to rehash the debate about what constitutes an empirical test.
-However, I aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
-The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
-The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
+However, I aim to emphasize the role of transparency in two frameworks that lend themselves to investigate induction.
+The first framework promotes transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
+The second framework promotes transparency under a science that wants to select a theory according to its expected predictive performance.
 

### 2023-02-01 manuscript.Rmd
"give us" to "suggest"

@@
 
-Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and suggest some theoretical tools to asses and control this overconfidence.
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.

### 2023-02-01 manuscript.Rmd
We are now to When
I personally don't like those narrative phrases like "we want to look at ..." or "this gives us... " and so on, but this is I think more of a personal style thingy ^^I just prefer the more general, passive style in writing

@@
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+When interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:

### 2023-02-01 manuscript.Rmd
same as last commit

@@
 Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
-We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
+The distance to the truth can be formalized in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.

### 2023-02-01 manuscript.Rmd
just added a "to"

@@
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
-What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
+What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
 

### 2023-02-01 manuscript.Rmd
together sounds somewhat personal
"and" can be kicked because one Hence

@@
 
-The expectation over the data together with Taylor series expansion yields:
+The expectation over the data complemented by the Taylor series expansion yields:
 \begin{equation}
@@
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
+This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.

### 2023-02-01 manuscript.Rmd
revert all leonies commits

--- 2023-02-01 manuscript.Rmd
@@
 To understand why transparency is crucial, it is important to understand that formal and informal induction are integral to empirical sciences.
-Any science must be able to communicate how its knowledge is generated.
+Any science must be able to communicate how it generates its knowledge.
 However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive process that contributed to them.
@@
 On the other hand, induction may render any test of a hypothesis ineffective.
-Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
+Therefore, I argue the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
 The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
@@
 Based on this distinction, the articles answer the second question (the *how*).
-I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration in regard of transparency is supplemented by practical guidance on how researchers may implement these approaches.
+I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
+The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches.
 
@@
 I do not attempt to rehash the debate about what constitutes an empirical test.
-However, I aim to emphasize the role of transparency in two frameworks that lend themselves to investigate induction.
-The first framework promotes transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
-The second framework promotes transparency under a science that wants to select a theory according to its expected predictive performance.
+However, I aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
+The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
+The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
 
-Both frameworks show how unaccounted induction leads to overconfidence in empirical results and suggest some theoretical tools to asses and control this overconfidence.
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
@@
 Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
-The distance to the truth can be formalized in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
+We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-When interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:

### 2023-02-03 manuscript.Rmd
add varios legal matters

@@
 \thispagestyle{empty}
-\null
+
+# Legal matters
+
+## Digital copy of the work {.unlisted .unnumbered}
+
+This work is made available online under the following source under the terms of the [Creative Commons Universal Public Domain Dedication | CC0](https://creativecommons.org/publicdomain/zero/1.0/):  
+<https://aaronpeikert.github.io/thesis/manuscript.pdf>
+
+## Acknowledgement {.unlisted .unnumbered}
+
+During the work on their dissertation, Aaron Peikert was a pre-doctoral fellow of the [International Max Planck Research School on Computational Methods in Psychiatry and Ageing Research (IMPRS COMP2PSYCH)](https://www.mps-ucl-centre.mpg.de/comp2psych) of
+the Max Planck Institute for Human Development, Berlin, Germany and University College London, United Kingdom.
+
+## Used resources and aids
+
+I used several digital aids to improve the writing (not the content), including grammerly.com, deepl.com, and mentor.duden.de.
+Also, to aid with the clarity of the writing, I asked Andreas Brandmaier, Carli Ochs, Leo Richter, Leonie Haggitte, and Caroline Gahrmann for feedback.
+Furthermore, I asked Leo Richter and Maximilian Ernst for help with mathematical notation and typesetting.
+I also prompted `ChatGPT` to write a song about reproducibility; while it was amusing, it was not funny enough to make an appearance in the text.
+
+## Declaration of independent work {.unlisted .unnumbered}
+
+I hereby declare that I completed the doctoral thesis independently based on the stated resources and
+aids.
+I have not applied for a doctoral degree elsewhere and do not have a corresponding doctoral degree.
+I have not submitted the doctoral thesis, or parts of it, to another academic institution and the thesis has not been accepted or rejected.
+I declare that I have acknowledged the Doctoral Degree Regulations which underlie the procedure of the
+Faculty of Life Sciences of Humboldt-Universität zu Berlin, as amended on 5th March 2015.
+Furthermore, I declare that no collaboration with commercial doctoral degree supervisors took place, and
+that the principles of Humboldt-Universität zu Berlin for ensuring good academic practice were abided by.
+
+\vspace{.5cm}
+
+\hfill{}Aaron Peikert
+
+\newpage
 \clearpage

### 2023-02-06 manuscript.Rmd
personal narative to passive
personal narrative to passive

@@
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-When interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
-Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
+When interested in how much information is lost when using $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], this results:
 <!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->

### 2023-02-06 manuscript.Rmd
unnecc
They will see that - so no technical need to highlight that
our to the

--- 2023-02-06 manuscript.Rmd
@@
 
-This setup is, of course, highly theoretical.
 $\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
@@
 
-To allow quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
-That means our theory implies a family of possible probability distributions that may describe reality.
+To allow quantities to be induced, we must assume that the theory is parameterized, e.g., $g(x|\theta)$.
+That means the theory implies a family of possible probability distributions that may describe reality.
 This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.

### 2023-02-06 manuscript.Rmd
same same
"here" is i.m.o not the best anchor for those sentences
if it is clear, to what you are referring, and that is mostly the case, I would leave it out. If you feel like there is some clarification needed about what you are referring about, then I would use 'refer','regard' or if it fits 'hence','thus','therefore' etc.

@@
 \end{equation}
-Of course, we usually do not know $\theta_*$.
+Usually $\theta_*$ is unknown.
 That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ are $n$ independent samples from $X\sim f$.
 
-The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
+The crucial point is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).

### 2023-02-06 manuscript.Rmd
personal narrative to passive, also found a can not

@@
 
-The crucial point is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
+The crucial point is to understand what happens when $\theta$ cannot be derive deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).

### 2023-02-06 manuscript.Rmd
s.o.
passive , also "think to" -> "think of"

@@
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
-It follows, almost surely, that we lose information:
+It follows, almost surely, that information is lost:
 \begin{equation}
@@
 
-Forming this expectation over data is a crucial step; it requires us to think beyond the data we observed to all the data we could have observed.
+Forming this expectation over data is a crucial step; it requires thinking beyond the observed data, of all the data we could have observed.
 There are two ways to get at this expectation.

### 2023-02-06 manuscript.Rmd
passive, also not sure if there need to be Kommata
the last part with the assumptions is maybe not necessary?

@@
 
-We usually favor procedures to induce $\hat \theta$ that promise unbiased estimates for the observed likelihood, given that their assumptions are met.
+Usually procedures inducing $\hat \theta$, that promise unbiased estimates for the observed likelihood, are favored, given that their assumptions are met.
 The observed likelihood is, therefore, often available, e.g., in maximum likelihood estimation.

### 2023-02-06 manuscript.Rmd
passive; overconfidence about -> overconfidence on
dreimal this hintereinander
passive the text part

@@
 The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-We may conclude that substituting deduced quantities with induced estimates leads to some overconfidence about how close one is to the truth.
+Which leads to the conclusion that substituting deduced quantities with induced estimates leads to some overconfidence on how close one is to the truth.
 This overconfidence is directly related to how much induction a model entails.
@@
 
-If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
+When aiming to induce quantities and correctly appraise a theory on the same data, one must know how much the appraisal for the induction has to be corrected.
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
-One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
+One condition is, since $\theta_*$ is unknown, to know the properties of the inductive process that generated $\hat\theta$.
+Then it is possible to formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
 

### 2023-02-06 manuscript.Rmd
passive

@@
 In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
-Instead of verisimilitude, one might be concerned with future performance.
+Instead of verisimilitude, there could be concerns with future performance.
 That is, how well does a theory do in predicting novel facts?
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
+In the future performance setup, it is not appealed to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
 \begin{equation}

### 2023-02-06 manuscript.Rmd
passive

@@
 
-Again we can define the loss we have observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020]:
+Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020] can be defined:
 \begin{equation}
@@
 
-However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
+However, it is not of interest how well the theory did on data that informed it but on future, yet unseen, data:
 \begin{equation}

### 2023-02-06 manuscript.Rmd
passive

@@
 
-Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
-That is, we use cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
+Instead of using the Taylor series expansion, data can be repeatedly sampled and the inductive process can be repeated.
+That is, using cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.

### 2023-02-06 manuscript.Rmd
passive
passive and maybe a comma

@@
 As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
-To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
+To make the link to the first approach even more clear, complexity may be defined in these terms [@hauensteinComputingAICBlackbox2018]:
 \begin{equation}
@@
 
-Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
-That drastically expands the set of inductive processes for which we can estimate the inductive bias since it is considerably easier to repeat the process than to derive the complexity analytically.
+Instead of a formal analysis to derive $\mathcal{C}$, it is possible to repeat the inductive process, i.e., using cross-validation.
+That drastically expands the set of inductive processes for which the inductive bias can be estimated, since it is considerably easier to repeat the process than to derive the complexity analytically.
 

### 2023-02-06 manuscript.Rmd
passive and "laying open" -> make public

@@
 
-Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
-It bears repeating that simply laying open what has been done is not enough.
+Now that the need for transparency about the inductive process has been established, a few more technical details can be droped to clarify what we have to make transparent.
+It bears repeating that making public what has been done is not enough.
 Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.

### 2023-02-06 manuscript.Rmd
passive

@@
 Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
-On a conceptual level, we want to compare the following:
+On a conceptual level, the following has to be compared:
 \begin{equation}
@@
 Where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
-To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
+To allow for induction, the theory gets replaced with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
 

### 2023-02-06 manuscript.Rmd
passive the text part

@@
 \end{equation}
-The idea is that we choose the version of our theory that best fits reality.
-However, we are forced to rely on a limited sample of reality.
-We are misled because these two factors, induction and limited sample size, interact.
-Choosing the best version of our theory based on a sample is almost surely suboptimal.
+The idea is that the version of our theory gets chosen that best fits reality.
+However, it is necessary to rely on a limited sample of reality.
+This is misleading because these two factors, induction and limited sample size, interact.
+Choosing the best version of the theory based on a sample is almost surely suboptimal.
 Therefore, the observed loss is an overconfident estimate of the loss in the future and closeness to the truth.

### 2023-02-06 manuscript.Rmd
found a can not
passive

@@
 Complexity is a function of the model, i.e., $\mathcal{C}(\text{model})$, not of the data.
-Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
+Without knowing the inductive process, researchers cannot judge the overconfidence, so the inductive process ought to be made transparent.
 
@@
 This observation is only useful if we know the extent of this optimism.
-However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{model})$) and give us two starting points for making this bias transparent.
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{model})$) and suggest two starting points for making this bias transparent.
 The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).

### 2023-02-06 manuscript.Rmd
passive and - if you don't take this then there has been a can not in line 433

@@
 
-As a first step, we can note that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
+As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
 Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
 If a researcher employs a linear model, the complexity is calculable.
-Even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
+Even if it were not, the linear model could be fitted on a large set of other data to asses its inductive behavior.
 However, suppose the researcher reconsiders the model based on the results and their internal model.
@@
 In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
-We can not ask the researcher to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
-However, we know that the resulting linear model is more complex than a linear regression formally implies.
+The researcher cannot be asked to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
+However, it is known that the resulting linear model is more complex than a linear regression formally implies.
 

### 2023-02-06 manuscript.Rmd
passive - but the sentence strikes me somewhat clonky (line 441)

@@
 
-That is not to say that we have no basis for judging the informal induction.
-We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
+That is not to say that there is no basis for judging the informal induction.
+The inductive decision can be thought of as reasonable, thus found unlikely that just about any variable would be added if the data suggests it.
 Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data-driven decision, which implies higher complexity.

### 2023-02-06 manuscript.Rmd
passive

@@
 The inductive decision can be thought of as reasonable, thus found unlikely that just about any variable would be added if the data suggests it.
-Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data-driven decision, which implies higher complexity.
-What one can say, however, is that this judgment is debatable and, therefore, subjective.
+Or the opposite, it might not seem well justified on theoretical grounds and deemed a purely data-driven decision, which implies higher complexity.
+What can be said, however, is that this judgment is debatable and, therefore, subjective.
 

### 2023-02-06 manuscript.Rmd
cannot
passive

@@
 It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
-This fact implies that for these situations, the complexity and, hence, the optimism bias can not be fully quantified.
+This fact implies that for these situations, the complexity and, hence, the optimism bias cannot be fully quantified.
 Though full transparency remains out of reach, researchers may do their best to provide a good basis for assessing complexity.
@@
 
-The imperative to enable proper judgment of theories is simple: induce only what is necessary and what you induce should, if possible, be done formally.
+The imperative to enable proper judgment of theories is simple: induce only what is necessary and what is induced should, if possible, be done formally.
 Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.

### 2023-02-06 manuscript.Rmd
Update manuscript.Rmd
cannot

@@
 
-Therefore, we have a theoretical bound that limits transparency about the inductive process.
-Some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
+Therefore, there is a theoretical bound that limits transparency about the inductive process.
+Some things cannot be made transparent in principle because some induction happens informally, so we cannot estimate the complexity with certainty.
 However, there is another bound that restricts transparency further.

### 2023-02-06 manuscript.Rmd
passive

@@
 Therefore, there is a theoretical bound that limits transparency about the inductive process.
-Some things cannot be made transparent in principle because some induction happens informally, so we cannot estimate the complexity with certainty.
+Some things cannot be made transparent in principle because some induction happens informally, so the complexity cannot be estimated with certainty.
 However, there is another bound that restricts transparency further.
 That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
-To achieve transparency about the inductive process and ultimately about the empirical support of a theory, we must increase both bounds.
+To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both bounds have to be increased.
 

### 2023-02-06 manuscript.Rmd
passive
cannot

@@
 Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992].
-However, through the lens of transparency about the inductive process, we may further refine our view on computational reproducibility.
+However, through the lens of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.
 
@@
 However, place two restrictions on the software.
-First, while able to compute results based on data, the software can not be understood by human researchers, e.g., because it is a binary program.
+First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is a binary program.
 Second, when the data set changes, the software does not work anymore.

### 2023-02-06 manuscript.Rmd
passive also there was a can not in line 481

@@
 The first condition rules out that the software can be analyzed formally.
-The second condition prevents us from changing the data and observing the results.
-Therefore, we have no means to assess the complexity of the inductive process and can not judge the empirical support.
-It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.
+The second condition prevents from changing the data and observing the results.
+Therefore, there are no means to assess the complexity of the inductive process and the empirical support cannot be judged.
+It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., it has to be ruled out one of the above conditions.
 

### 2023-02-06 manuscript.Rmd
passived the text part

@@
 
-Suppose we relax the second condition; then we have a "black box" that is repeatable on other data.
-In that case, we still have transparency about the inductive process because we can estimate the complexity of the inductive process, i.e., using cross-validation.
+Suppose the second condition is relaxed; then there is a "black box" that is repeatable on other data.
+In that case, there still is transparency about the inductive process because the complexity of the inductive process can be estimated, i.e., using cross-validation.
 Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
-Therefore, we must extend the traditional definition of computational reproducibility and at least also require that the process can be repeated on other similar data.
+Therefore, the traditional definition of computational reproducibility has to be extended and further it is required that the process can be repeated on other similar data.
 

### 2023-02-06 manuscript.Rmd
must -> has to
passive

@@
 
-Therefore, to provide transparency, computational reproducibility must satisfy two requirements.
+Therefore, to provide transparency, computational reproducibility has to satisfy two requirements.
 First, computational reproducibility must ensure that the same data lead to the same results.
@@
 Computational reproducibility enables transparency about formal inductive processes.
-However, informal inductive processes deprive us of complete transparency.
+However, informal inductive processes deprive of complete transparency.
 Therefore, it is prudent to replace informal induction with formal induction as much as possible.

### 2023-02-06 manuscript.Rmd
de personified

@@
 A perfectly reproducible linear regression is, from the outside, indistinguishable from one that was cherry-picked from hundreds of possible regressions.
-However, our expectations regarding verisimilitude and future performance between the cherry-picked and the simple regression should vary substantially.
+However, the expectations regarding verisimilitude and future performance between the cherry-picked and the simple regression should vary substantially.
 

### 2023-02-06 manuscript.Rmd
passive line 526still clunky

@@
 Researchers could simply try to explain the inductive process after the fact.
-This task is not easy because we need to know what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
+This task is not easy because it needs to be known what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
 This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.

### 2023-02-06 manuscript.Rmd
lay open -> publish?
Update manuscript.Rmd

@@
 This task is challenging to accomplish in practice.
-After all, researchers would have to lay open the process that led to the induction, i.e., their thoughts and full mental model.
+After all, researchers would have to publish the process that led to the induction, i.e., their thoughts and full mental model.
 Dissolving the distinction between formal and informal induction would mean surrendering the advantage a formal inductive process provides; instead of quantifying complexity, the complexity must be judged with considerable uncertainty.
@@
 Registration is a strong argument that the data has not influenced the results more than the complexity of the formal inductive process implies.
-If the data does not yet exist, it can not influence the results.
+If the data does not yet exist, it cannot influence the results.
 Note that registration is about the flow of information, i.e., the data used for induction, and not the temporal order.

### 2023-02-06 manuscript.Rmd
passive

@@
 They have to reason about all the possible data patterns they might encounter.
-While this makes registration so difficult, it is exactly the information we need and from which we may derive the complexity of the inductive process.
+While this makes registration so difficult, it is exactly the information we need and from which the complexity of the inductive process might be derived.
 

### 2023-02-06 manuscript.Rmd
cannot
passive
would delete here or find a better substitute

@@
 Being able to account for uncertainty is indispensable for the question of how to deal with informal induction.
-If induction can not be repeated at will nor formally analyzed, we are uncertain about the complexity of the inductive process.
-Theoretical risk is conceptually related to the here discussed concept of complexity but put into the context of Bayesian philosophy of science.
+If induction cannot be repeated at will nor formally analyzed, there is uncertainty about the complexity of the inductive process.
+Theoretical risk is conceptually related to the discussed concept of complexity but put into the context of Bayesian philosophy of science.
 Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence, and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory if the evidence is more likely under the theory).

### 2023-02-06 manuscript.Rmd
people need no invitation for criticism
passive

@@
 Such a set of possible implications and the process of choosing among them according to a loss function is what I understand as a model.
-However, this conceptualization may be criticized.
 It is not entirely clear how to chose the process of selecting the best fitting version of the theory.
@@
 Is the application of mathematically or computationally convenient methods, as is usual in statistical modeling, sufficient?
-Must we account for the arbitrariness of the selection process?
+Does the arbitrariness of the selection process needs to be accounted for?
 
@@
 It might happen that in the end, more than one inductive process may be applicable, and no criterium makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
-We have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
+It has been dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
 A possible direction for future research is to explore if forming the expectation over inductive processes is similarly fruitful.

### 2023-02-06 manuscript.Rmd
cannot

@@
 However, the choice of loss function may be highly consequential for the appraisal of a theory.
-Not only is this choice consequential, but it is easily imaginable that researchers can not agree on a metric to judge the theory.
+Not only is this choice consequential, but it is easily imaginable that researchers cannot agree on a metric to judge the theory.
 
@@
 Allowing loss functions to vary between researchers also implies that the complexity of the inductive process varies.
-It follows that the correction for complexity can not be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
+It follows that the correction for complexity cannot be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
 This again highlights the importance of making the process transparent.

### 2023-02-06 manuscript.Rmd
Update manuscript.Rmd
laying open -> highlighting

@@
 Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
+Highlighting the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
@@
 However, I hope we might abandon the tradition of analyzing scientific practices only after they have been long established.
-Instead we should analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
+Instead the aim should be to analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
 That way, theoretical insight may shape application.

### 2023-02-06 manuscript.Rmd
not -> no

@@
 Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Highlighting the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
+Highlighting the difficulties and attempting to overcome them are no recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.

### 2023-02-06 manuscript.Rmd
passive
Update manuscript.Rmd

@@
 Either the variance is known, or it has to be estimated from data.
-Consider a placebo and an experimental group, where we hypothesize and test a mean difference using a simple t-test.
+Consider a placebo and an experimental group, it is possible to hypothesize and test a mean difference using a simple t-test.
 The decision to reject the null hypothesis (groups have equal means) depends, besides true mean difference and sample size, on the observed variance.
@@
 <!--necessity of induction-->
-In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
+In other words, induction gives theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
 It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.

### 2023-02-10 manuscript.Rmd
Tone done a lot of passiv use from leonie.
This reverts commit 54fd1a178694de09a4986f501b62b1c53ce7815f.

@@
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
-What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
+What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
 
+This setup is, of course, highly theoretical.
 $\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
@@
 
-To allow quantities to be induced, we must assume that the theory is parameterized, e.g., $g(x|\theta)$.
-That means the theory implies a family of possible probability distributions that may describe reality.
+To allow quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
+That means our theory implies a family of possible probability distributions that may describe reality.
 This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
@@
 \end{equation}
-Usually $\theta_*$ is unknown.
+Of course, we usually do not know $\theta_*$.
 That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ are $n$ independent samples from $X\sim f$.
 
-The crucial point is to understand what happens when $\theta$ cannot be derive deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
+The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
-It follows, almost surely, that information is lost:
+It follows, almost surely, that we lose information:
 \begin{equation}
@@
 
-Forming this expectation over data is a crucial step; it requires thinking beyond the observed data, of all the data we could have observed.
+Forming this expectation over data is a crucial step; it requires us to think beyond the data we observed to all the data we could have observed.
 There are two ways to get at this expectation.
@@
 
-Usually procedures inducing $\hat \theta$, that promise unbiased estimates for the observed likelihood, are favored, given that their assumptions are met.
+We usually favor procedures to induce $\hat \theta$ that promise unbiased estimates for the observed likelihood, given that their assumptions are met.
 The observed likelihood is, therefore, often available, e.g., in maximum likelihood estimation.
@@
 
-The expectation over the data complemented by the Taylor series expansion yields:
+The expectation over the data together with Taylor series expansion yields:
 \begin{equation}
@@
 The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-Which leads to the conclusion that substituting deduced quantities with induced estimates leads to some overconfidence on how close one is to the truth.
+We may conclude that substituting deduced quantities with induced estimates leads to some overconfidence about how close one is to the truth.
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
+This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.
@@
 
-When aiming to induce quantities and correctly appraise a theory on the same data, one must know how much the appraisal for the induction has to be corrected.
+If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
-One condition is, since $\theta_*$ is unknown, to know the properties of the inductive process that generated $\hat\theta$.
-Then it is possible to formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
+One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
 
@@
 In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
-Instead of verisimilitude, there could be concerns with future performance.
+Instead of verisimilitude, one might be concerned with future performance.
 That is, how well does a theory do in predicting novel facts?
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, it is not appealed to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
+In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
 \begin{equation}
@@
 
-Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020] can be defined:
+Again we can define the loss we have observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020]:
 \begin{equation}
@@
 
-However, it is not of interest how well the theory did on data that informed it but on future, yet unseen, data:
+However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
 \begin{equation}
@@
 
-Instead of using the Taylor series expansion, data can be repeatedly sampled and the inductive process can be repeated.
-That is, using cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
+Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
+That is, we use cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
@@
 As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
-To make the link to the first approach even more clear, complexity may be defined in these terms [@hauensteinComputingAICBlackbox2018]:
+To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
 \begin{equation}
@@
 
-Instead of a formal analysis to derive $\mathcal{C}$, it is possible to repeat the inductive process, i.e., using cross-validation.
-That drastically expands the set of inductive processes for which the inductive bias can be estimated, since it is considerably easier to repeat the process than to derive the complexity analytically.
+Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
+That drastically expands the set of inductive processes for which we can estimate the inductive bias since it is considerably easier to repeat the process than to derive the complexity analytically.
 
@@
 
-Now that the need for transparency about the inductive process has been established, a few more technical details can be droped to clarify what we have to make transparent.
-It bears repeating that making public what has been done is not enough.
+Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
+It bears repeating that simply laying open what has been done is not enough.
 Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
-On a conceptual level, the following has to be compared:
+On a conceptual level, we want to compare the following:
 \begin{equation}
@@
 Where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
-To allow for induction, the theory gets replaced with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
+To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
 

### 2023-02-11 manuscript.Rmd
incorporate andreas feedback

@@
 Psychology is a difficult science [@meehlTheoreticalRisksTabular1978].
-Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Highlighting the difficulties and attempting to overcome them are no recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
-We can roughly divide the proposed countermeasures into two categories, those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that aim to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the claim itself.
+Highlighting the difficulties and trying to overcome them are no recent trends, though they have been invigorated by the so-called replication crisis in psychology; a crisis that has also begun to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As psychology grapples with this crisis of confidence in its empirical findings, several causes have been identified and respective remedies have been proposed.
+The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
 <!--bla-transparency-bla-->
-In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory.
-Testing a theory empirically is often viewed as deductive since the theory is making statements about the data.
-Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
-I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results if unaccounted for.
-The extent of this bias depends on the inductive process.
-The inductive process, therefore, must be transparent to other researchers because they are otherwise unable to judge the empirical support of a theory.
+In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory [@leePopperFalsificationCorroboration2021].
+Testing a theory empirically is often viewed as deductive since the theory is making statements about future observations.
+Empirical scientists, however, often simultaneously engage in induction by deriving general statements from past observation.
+I will argue that the inductive element in empirical tests leads to overconfidence in the empirical results if unaccounted for.
+The extent of this bias depends on properties of the inductive process.
+The induction, therefore, must be transparent to other researchers in order for them to be able to judge the empirical support of a theory.
 
-The above distinction, between statistical and sociological countermeasures, arises from two different kinds of induction.
-On the one hand, we have inductive processes that are well-defined in the form of statistical methods; on the other hand, researchers also engage in more informal inductive behavior outside of well-defined models.
-These different kinds of induction require different countermeasures.
-Statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
-To understand why transparency is crucial, it is important to understand that formal and informal induction are integral to empirical sciences.
+The above distinction, between statistical and sociological countermeasures, arises from two sources of inductive bias.
+On the one hand, we have inductive processes that are well-defined in the form of statistical methods, while on the other, researchers also engage in more informal inductive behavior outside of well-defined models.
+These different kinds of inductive behavior require different countermeasures.
+When rigorously applied, statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
+To understand why transparency is crucial, it is important to comprehend how integral formal and informal induction are for empirical sciences.
 Any science must be able to communicate how it generates its knowledge.
-However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive process that contributed to them.
-Therefore, transparency is more than a virtue that may improve knowledge gains in empirical sciences but is an indispensable property.
+However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive processes involved.
+Therefore, transparency is more than a virtue that may improve empirical sciences somewhat but is an indispensable property.
 
 <!--induction is necessary for psychological science-->
-To function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
-In psychology, this is not a purely deductive endeavor.
-Very few psychological theories are precise enough to derive testable statements.
-While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
+In order to function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
+In psychology, this is not a purely deductive endeavor [@hitchcockPredictionAccommodationRisk2004; @meehlAppraisingAmendingTheories1990].
+Very few psychological theories are precise enough to derive testable statements [@friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020].
+While it is tempting to claim that a theory makes deductively testable claims [@leePopperFalsificationCorroboration2021] by, for example, implying a mean difference between two groups, such a statement is not testable on its own.
 Making inferences about a mean difference requires knowledge about the variance.
-Either the variance is known, or it has to be estimated from data.
-Consider a placebo and an experimental group, it is possible to hypothesize and test a mean difference using a simple t-test.
-The decision to reject the null hypothesis (groups have equal means) depends, besides true mean difference and sample size, on the observed variance.
+Either the variance is known, than a purely deductive test is possible, or it has to be estimated from data, making induction part of the empirical test.
+Consider a placebo and an experimental group; there is possible to hypothesize and test a mean difference using a simple t-test.
+The decision to reject the null hypothesis (groups have equal means) depends on, besides true mean difference and sample size, the observed variance.
 However, the variance needs to be induced from the data.
@@
 <!--sometimes the use of induction is well justified-->
-Induction is necessary and, in the present case, innocuous for psychology as a science.
+Induction is necessary and, in the present case, harmless for psychology as a science.
 In this case, it is innocuous because the bias from inducing the variance can be accounted for.
-A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
 z≈`r round(qnorm(.99), 2)`).
-It is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
+It has been widely known for decades that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908], i.e., the inductive bias vanishes with increasing sample size.
 Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would be no psychological theory that could be subjected to an empirical test.
+If they had to, there would be currently very few psychological theory that could be subjected to an empirical test [@muthukrishnaProblemTheory2019].
 
 <!--necessity of induction-->
-In other words, induction gives theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
-It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.
+In other words, induction gives theories some leeway to be imprecise and contain "blank" spaces, filled in later based on observation.
+It allows researchers to focus on the essential statements of their theories and choose to leave some parts (such as auxiliary assumptions) to be determined inductively.
 In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
@@
 <!--transparency about induction is required-->
-So empirical researchers find themselves in a pinch.
+So empirical researchers find themselves in a predicament.
 On the one hand, they need induction to test their imprecise theories.
 On the other hand, induction may render any test of a hypothesis ineffective.
-Therefore, I argue the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
+Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
 The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
@@
 It is addressed in the synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
-Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).
+Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge in current research practice (sociological factors).
 
 <!--how it relates to the articles-->
-Based on this distinction, the articles answer the second question (the *how*).
-I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches.
+Based on this distinction, my research articles that are part of this dissertation answer the second question (the *how*).
+I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
+The conceptualization of computational reproducibility and preregistration in regard of transparency is supplemented by practical guidance on how researchers may implement these approaches.
 
@@
 I do not attempt to rehash the debate about what constitutes an empirical test.
-However, I aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
+However, I aim to examine the role of transparency in two frameworks that lend themselves to investigate the sources of inductive bias.
 The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
@@
 
-Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools on how to disclose induction so that this overconfidence can be assessed and controlled.
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
@@
 Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
-We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
+The distance to the truth can be formalized in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+The quantity of interst is how much information is lost if we use $g(x)$, our theory as description of the world..., instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
@@
 
-Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
+Many readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
-What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.
+What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
 
-This setup is, of course, highly theoretical.
-$\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
-However, this fact does not impede us from getting closer to the truth because we can still compare the two theories relative to each other.
+Note that in practice, $\mathcal{L}_{KL}(f, g)$ is unknowable, since the full truth is unobservable.
+However, this fact does not impede us from getting closer to the truth because we can still compare the competing theories relative to each other.
 Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
@@
 
-The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
+The crucial point is to understand what happens when $\theta$ cannot be derive deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
-It follows, almost surely, that we lose information:
+It follows, almost surely, that information is lost:
 \begin{equation}
@@
 The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
-The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
+The point is to make a statement about the theory, not to make a statement about the data in light of the theory.
 If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
@@
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
+This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.
@@
 
-If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
+If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for how the data influences the theory.
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
 One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's Cp [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].
 
@@
 In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
-Instead of verisimilitude, one might be concerned with future performance.
+Instead of verisimilitude, one might be concerned with future performance [@yarkoniChoosingPredictionExplanation2017].
 That is, how well does a theory do in predicting novel facts?
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
+In the future performance setup, we do not appeal to ground truth (I drop $\sim{}f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
 \begin{equation}
@@
 As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
-To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
+To make the link to the first approach even more clear, we may give an alternative definition of complexity as the expected difference between observed prediction error and expected prediction error [@hauensteinComputingAICBlackbox2018]:
 \begin{equation}
@@
 Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
-That drastically expands the set of inductive processes for which we can estimate the inductive bias since it is considerably easier to repeat the process than to derive the complexity analytically.
+The former requires the derivatives with regard to the parameters, so I call it parametric estimator of complexity.
+The latter eschews the need for that, so I call it non-parametric estimator of complexity.
+Only requiring that a process can be repeated on other data, drastically expands the set of inductive processes for which we can estimate the inductive bias.
 
@@
 Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
-It bears repeating that simply laying open what has been done is not enough.
+It bears repeating that simply revealing what has been done is not enough.
 Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
@@
 \end{equation}
+
+So the observed loss is an underestimation that has to be corrected for by considering the complexity of the model.
 \begin{equation}
@@
 The ability to adapt to data is independent of the specific data that was observed.
-Complexity is a function of the model, i.e., $\mathcal{C}(\text{model})$, not of the data.
+Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
 Without knowing the inductive process, researchers cannot judge the overconfidence, so the inductive process ought to be made transparent.
@@
 This observation is only useful if we know the extent of this optimism.
-However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{model})$) and suggest two starting points for making this bias transparent.
-The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and suggest two starting points for making this bias transparent.
+The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
 Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.
@@
 Even a casual consideration of the above formalization should strike anyone who has ever worked with empirical data as unrealistic.
-No one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
+This flexibility is often called researchers' degrees of freedom [@simmonsFalsePositivePsychologyUndisclosed2011].
+Using researchers' degrees of freedom opportunistically is an obvious problem [@yarkoniGeneralizabilityCrisis2022; @wichertsDegreesFreedomPlanning2016].
+However, no one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
 The point is to set the goalpost and have a yardstick to measure how well we can judge empirical support.
@@
 Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
-If a researcher employs a linear model, the complexity is calculable.
-Even if it were not, the linear model could be fitted on a large set of other data to asses its inductive behavior.
-However, suppose the researcher reconsiders the model based on the results and their internal model.
+If a researcher employs a linear model, the complexity is calculable parametrically.
+Even if it were not, the linear model could be fitted on a large set of other data to asses its expected inductive behavior nonparametrically.
+However, suppose the researcher reconsiders the model based on the results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
 For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
@@
 $$
-\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
+\mathcal{C}(\text{Model}) = \mathcal{C}(\text{Model}_\text{formal}) + \mathcal{C}(\text{Model}_\text{informal})
 $$
@@
 
-In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
-I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
+In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bound).
+I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.
 
@@
 Such transparency can be achieved by computational reproducibility.
-Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992].
+Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992; @peikertReproducibleDataAnalysis2021; @peikertReproducibleResearchTutorial2021].
 However, through the lens of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.
@@
 However, place two restrictions on the software.
-First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is a binary program.
+First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is an executable program, of which the source code may not be openly available.
 Second, when the data set changes, the software does not work anymore.
@@
 Removing the human from the loop facilitates that the data can be easily substituted by similar data.
-Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the judgment of the complexity.
+Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the assesment of the complexity.
 
@@
 First, it must be unambiguous what results are generated by which inductive process.
-This can be ensured by employing dynamic document creation.
+This can be ensured by employing dynamic document creation [@knuth1984literate; @xie2019; @xie2015; @Aust_papaja_Prepare_reproducible_2022].
 Second, the version of the software that is used to generate the results must be known.
-This can be ensured by documenting the version using software management.
-Third, the version of the author's written code (and possibly text) must be tracked using version control systems.
-Fourth, how the data relates to software and computing infrastructure must be managed by, e.g., workflow automation.
+This can be ensured by documenting the version using software management [@merkelDockerLightweightLinux2014; @wiebelsLeveragingContainersReproducible2021].
+Third, the version of the author's written code (and possibly text) must be tracked using version control systems [@chaconProGit2014; @vanlissaWORCSWorkflowOpen2021].
+Fourth, how the data relates to software and computing infrastructure must be managed by, e.g., workflow automation [@feldmanMakeProgramMaintaining1979; @kimImplementingGitHubActions2022; @kinsmanHowSoftwareDevelopers2021].
 
@@
 Trusting a post hoc explanation of an inductive process is not entirely unreasonable.
-However, post hoc, formal and informal induction are indistinguishable unless the account is comprehensive.
+However, post hoc, formal and informal induction are indistinguishable unless the account is exhaustive.
 Again, not only comprehensive for the data that was observed but also for all possible data.
@@
 
-Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extend.
+Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extent.
 In other words, researchers still might be very unsure how the data will look and therefore rely on induction to fill in the blanks.
@@
 In that case, a variable selection mechanism can be specified.
-Or suppose it is unclear what functional form to expect.
-In that case, many statistical models exist that allow almost arbitrary functional forms.
+Or suppose it is unclear what functional form to expect between predictor and outcome.
+In that case, many statistical models exist that allow almost arbitrary functional forms [@rissanenUniversalCodingInformation1984], e.g., smoothing splines [@cravenSmoothingNoisyData1978], random forests [@hoRandomSubspaceMethod1998], neural networks [@amariUniversalTheoremLearning1993].
 If the theory does not suffice to define an outlier, this decision can be made data-dependent, etc.
@@
 This version, with "mock" results based on simulated data, serves as the registration.
-When the data is gathered, the results are updated to reflect the actual data.
+When the data are gathered, the results are updated to reflect the actual observations.
 
@@
 The extent of this overconfidence is a function of the inductive process.
-Specifically, I propose that transparency must concern the extent to which data may influence the results and term this complexity.
+Specifically, I propose that transparency must concern the extent to which data may influence the results and which is captured in the term complexity.
 I show that the complexity of an inductive process can be quantified if the process can be analyzed or repeated on similar data.
@@
 
-My consideration of transparency centers around the use of induction to test imprecise theories.
-A theory is imprecise because it must be refined using data to make testable statements.
+My consideration of transparency centers around the necessity to allow for induction to test imprecise theories.
+I find it unconvincing that with some more effort researchers will develop better and more precise theories from thin air.
+Instead, I find it more plausible that psychological sciences will refine theories using data to make testable statements [@brandmaierTheoryguidedExplorationStructural20161205] and account for the bias the induction entails.
 I conceptualize induction to derive testable statements as choosing among a set of implications of the theory.
@@
 A certain statistical method may imply a set of possible implications that only partially overlaps with the theory.
-Developing more powerful inductive algorithms is a very active field of research in the machine learning community.
+Developing more powerful inductive algorithms is a very active field of research in the machine learning community [@brandmaierStructuralEquationModel2013].
 However, having more and more powerful tools for induction simply means that more and more vague theories may be tested.
@@
 
-It might happen that in the end, more than one inductive process may be applicable, and no criterium makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
-It has been dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
+It might happen that in the end, more than one inductive process may be applicable, and no criterion makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
+I have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
 A possible direction for future research is to explore if forming the expectation over inductive processes is similarly fruitful.
@@
 This again highlights the importance of making the process transparent.
-If the process is transparent, the readers might change the loss function and calculate complexity accordingly.
+If the process is transparent, the readers may change the loss function and calculate complexity accordingly.
 
@@
 
-This dissertation is concerned with a narrow definition of transparency.
+This dissertation proposes a narrow definition of transparency.
 I have not addressed essential parts of the scientific process, among them:
@@
 How should they go about gathering data concerning the chosen theory?
-When they are ready to publish their conclusions, how should these conclusions be evaluated?
+When are they ready to publish their conclusions?
+How should these conclusions be evaluated?
 Which conclusions warrant publication?

### 2023-02-13 manuscript.Rmd
increase width of abstract

@@
   \setotherlanguage[]{german}
+
+  \let\oldabstract\abstract
+  \let\oldendabstract\endabstract
+  \makeatletter
+  \renewenvironment{abstract}
+  {\renewenvironment{quotation}%
+                 {\list{}{\addtolength{\leftmargin}{-1cm} % change this value to add or remove length   to the the default
+                          \listparindent 1.5em%
+                          \itemindent    \listparindent%
+                          \rightmargin   \leftmargin%
+                          \parsep        \z@ \@plus\p@}%
+                  \item\relax}%
+                 {\endlist}%
+  \oldabstract}
+  {\oldendabstract}
+  \makeatother
 fontsize: 11pt

### 2023-02-13 manuscript.Rmd
small errors in front matter
hide ugly latex stuff

--- 2023-02-11 manuscript.Rmd
@@
   bookdown::pdf_document2:
+    includes:
+      in_header: "preamble.tex"
     latex_engine: "lualatex"
@@
     keep_tex:  true
-header-includes: |
-  \usepackage{libertine}
-  %\usepackage{ebgaramond}
-  %\usepackage{garamondlibre}
-  %\usepackage{CormorantGaramond}
-  %\usepackage{MinionPro}
-  \usepackage{pdfpages}
-
-  % below works well with hyperref, but spacing is standard (to large)
-  %\usepackage{sectsty}
-  %\sectionfont{\Large\normalfont\scshape\raggedright}
-  %\subsectionfont{\large\normalfont\scshape\raggedright}
-
-    % titlesec does not play nice with RMarkdown, so hack:
-  \let\paragraph\oldparagraph
-  \let\subparagraph\oldsubparagraph
-  \usepackage[sc, compact, raggedright]{titlesec}
-  % titlesec does not play nice with hyperref, so hack:
-  \newcommand{\sectionbreak}{\phantomsection}
-  \newcommand{\subsectionbreak}{\phantomsection}
-
-  \usepackage{titling}
-  \pretitle{
-     \begin{minipage}{\textwidth}
-     \includepdf[offset=5mm 0mm]{cover.pdf}
-     \end{minipage}
-     \newpage
-     \clearpage%
-     \thispagestyle{empty}%
-     \addtocounter{page}{-2}%
-     \null%
-     \clearpage
-     \begin{center}\LARGE}
-  \posttitle{\end{center}}
-
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-  \setotherlanguage[]{german}
 fontsize: 11pt
@@
 
-\newpage
-\clearpage
-\thispagestyle{empty}
-\begin{center}
-\vspace*{\fill}
-For my mother.\vspace*{\fill}
-\end{center}
-\clearpage
-
-\newpage
-\clearpage
-\thispagestyle{empty}
-# Acknowledgments {.unlisted .unnumbered}
-
-No one appreciated yet.
-\clearpage
-\thispagestyle{empty}
-\begin{center}
-\vspace*{\fill}
-Insight must precede application.
-\vspace*{\fill}
-\end{center}
-\newpage
-\clearpage
-
-\begin{spacing}{1}
-\tableofcontents
-\end{spacing}
-
-\newpage
-\clearpage
-\thispagestyle{empty}
+```{r child='frontmatter.md'}
+```
 
@@
 
-## Used resources and aids
+## Used resources and aids {.unlisted .unnumbered}
 

### 2023-02-15 manuscript.Rmd
Thanks for the reference :)

--- 2023-02-11 manuscript.Rmd
@@
 I used several digital aids to improve the writing (not the content), including grammerly.com, deepl.com, and mentor.duden.de.
-Also, to aid with the clarity of the writing, I asked Andreas Brandmaier, Carli Ochs, Leo Richter, Leonie Haggitte, and Caroline Gahrmann for feedback.
+Also, to aid with the clarity of the writing, I asked Andreas Brandmaier, Carli Ochs, Leo Richter, Leonie Hagitte, and Caroline Gahrmann for feedback.
 Furthermore, I asked Leo Richter and Maximilian Ernst for help with mathematical notation and typesetting.
@@
 Psychology is a difficult science [@meehlTheoreticalRisksTabular1978].
-Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the claim itself.
-Highlighting the difficulties and trying to overcome them are no recent trends, though they have been invigorated by the so-called replication crisis in psychology; a crisis that has also begun to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As psychology grapples with this crisis of confidence in its empirical findings, several causes have been identified and respective remedies have been proposed.
-The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
+Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
+We can roughly divide the proposed countermeasures into two categories, those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that aim to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
 <!--bla-transparency-bla-->
-In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory [@leePopperFalsificationCorroboration2021].
-Testing a theory empirically is often viewed as deductive since the theory is making statements about future observations.
-Empirical scientists, however, often simultaneously engage in induction by deriving general statements from past observation.
-I will argue that the inductive element in empirical tests leads to overconfidence in the empirical results if unaccounted for.
-The extent of this bias depends on properties of the inductive process.
-The induction, therefore, must be transparent to other researchers in order for them to be able to judge the empirical support of a theory.
+In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory.
+Testing a theory empirically is often viewed as deductive since the theory is making statements about the data.
+Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
+I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results if unaccounted for.
+The extent of this bias depends on the inductive process.
+The inductive process, therefore, must be transparent to other researchers because they are otherwise unable to judge the empirical support of a theory.
 
-The above distinction, between statistical and sociological countermeasures, arises from two sources of inductive bias.
-On the one hand, we have inductive processes that are well-defined in the form of statistical methods, while on the other, researchers also engage in more informal inductive behavior outside of well-defined models.
-These different kinds of inductive behavior require different countermeasures.
-When rigorously applied, statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
-To understand why transparency is crucial, it is important to comprehend how integral formal and informal induction are for empirical sciences.
-Any science must be able to communicate how it generates its knowledge.
-However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive processes involved.
-Therefore, transparency is more than a virtue that may improve empirical sciences somewhat but is an indispensable property.
+The above distinction, between statistical and sociological countermeasures, arises from two different kinds of induction.
+On the one hand, we have inductive processes that are well-defined in the form of statistical methods; on the other hand, researchers also engage in more informal inductive behavior outside of well-defined models.
+These different kinds of induction require different countermeasures.
+Statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
+To understand why transparency is crucial, it is important to understand that formal and informal induction are integral to empirical sciences.
+Any science must be able to communicate how its knowledge is generated.
+However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive process that contributed to them.
+Therefore, transparency is more than a virtue that may improve knowledge gains in empirical sciences but is an indispensable property.
 
 <!--induction is necessary for psychological science-->
-In order to function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
-In psychology, this is not a purely deductive endeavor [@hitchcockPredictionAccommodationRisk2004; @meehlAppraisingAmendingTheories1990].
-Very few psychological theories are precise enough to derive testable statements [@friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020].
-While it is tempting to claim that a theory makes deductively testable claims [@leePopperFalsificationCorroboration2021] by, for example, implying a mean difference between two groups, such a statement is not testable on its own.
+To function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
+In psychology, this is not a purely deductive endeavor.
+Very few psychological theories are precise enough to derive testable statements.
+While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
 Making inferences about a mean difference requires knowledge about the variance.
-Either the variance is known, than a purely deductive test is possible, or it has to be estimated from data, making induction part of the empirical test.
-Consider a placebo and an experimental group; there is possible to hypothesize and test a mean difference using a simple t-test.
-The decision to reject the null hypothesis (groups have equal means) depends on, besides true mean difference and sample size, the observed variance.
+Either the variance is known, or it has to be estimated from data.
+Consider a placebo and an experimental group, where we hypothesize and test a mean difference using a simple t-test.
+The decision to reject the null hypothesis (groups have equal means) depends, besides true mean difference and sample size, on the observed variance.
 However, the variance needs to be induced from the data.
@@
 <!--sometimes the use of induction is well justified-->
-Induction is necessary and, in the present case, harmless for psychology as a science.
+Induction is necessary and, in the present case, innocuous for psychology as a science.
 In this case, it is innocuous because the bias from inducing the variance can be accounted for.
-A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
 z≈`r round(qnorm(.99), 2)`).
-It has been widely known for decades that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908], i.e., the inductive bias vanishes with increasing sample size.
+It is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
 Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would be currently very few psychological theory that could be subjected to an empirical test [@muthukrishnaProblemTheory2019].
+If they had to, there would be no psychological theory that could be subjected to an empirical test.
 
 <!--necessity of induction-->
-In other words, induction gives theories some leeway to be imprecise and contain "blank" spaces, filled in later based on observation.
-It allows researchers to focus on the essential statements of their theories and choose to leave some parts (such as auxiliary assumptions) to be determined inductively.
+In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
+It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.
 In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
@@
 <!--transparency about induction is required-->
-So empirical researchers find themselves in a predicament.
+So empirical researchers find themselves in a pinch.
 On the one hand, they need induction to test their imprecise theories.
@@
 It is addressed in the synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
-Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge in current research practice (sociological factors).
+Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).
 
 <!--how it relates to the articles-->
-Based on this distinction, my research articles that are part of this dissertation answer the second question (the *how*).
+Based on this distinction, the articles answer the second question (the *how*).
 I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
@@
 I do not attempt to rehash the debate about what constitutes an empirical test.
-However, I aim to examine the role of transparency in two frameworks that lend themselves to investigate the sources of inductive bias.
-The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
-The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
+However, I aim to emphasize the role of transparency in two frameworks that lend themselves to investigate induction.
+The first framework promotes transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
+The second framework promotes transparency under a science that wants to select a theory according to its expected predictive performance.
 
-Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools on how to disclose induction so that this overconfidence can be assessed and controlled.
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and suggest some theoretical tools to asses and control this overconfidence.
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
@@
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-The quantity of interst is how much information is lost if we use $g(x)$, our theory as description of the world..., instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
-Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
+When interested in how much information is lost when using $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], this results:
 <!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
@@
 
-Many readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
+Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
@@
 
-Note that in practice, $\mathcal{L}_{KL}(f, g)$ is unknowable, since the full truth is unobservable.
-However, this fact does not impede us from getting closer to the truth because we can still compare the competing theories relative to each other.
+This setup is highly theoretical.
+$\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
+However, this fact does not impede us from getting closer to the truth because we can still compare the two theories relative to each other.
 Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
@@
 
-The crucial point is to understand what happens when $\theta$ cannot be derive deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
+The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
 Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
-It follows, almost surely, that information is lost:
+It follows, almost surely, that we lose information:
 \begin{equation}
@@
 The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
-The point is to make a statement about the theory, not to make a statement about the data in light of the theory.
+The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
 If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
@@
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
+This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.
@@
 
-If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for how the data influences the theory.
+If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
 One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's Cp [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].
 
@@
 In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
-Instead of verisimilitude, one might be concerned with future performance [@yarkoniChoosingPredictionExplanation2017].
+Instead of verisimilitude, one might be concerned with future performance.
 That is, how well does a theory do in predicting novel facts?
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth (I drop $\sim{}f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
+In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
 \begin{equation}
@@
 As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
-To make the link to the first approach even more clear, we may give an alternative definition of complexity as the expected difference between observed prediction error and expected prediction error [@hauensteinComputingAICBlackbox2018]:
+To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
 \begin{equation}
@@
 Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
-The former requires the derivatives with regard to the parameters, so I call it parametric estimator of complexity.
-The latter eschews the need for that, so I call it non-parametric estimator of complexity.
-Only requiring that a process can be repeated on other data, drastically expands the set of inductive processes for which we can estimate the inductive bias.
+That drastically expands the set of inductive processes for which we can estimate the inductive bias since it is considerably easier to repeat the process than to derive the complexity analytically.
 
@@
 Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
-It bears repeating that simply revealing what has been done is not enough.
+It bears repeating that simply laying open what has been done is not enough.
 Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
@@
 \end{equation}
-The idea is that the version of our theory gets chosen that best fits reality.
-However, it is necessary to rely on a limited sample of reality.
-This is misleading because these two factors, induction and limited sample size, interact.
-Choosing the best version of the theory based on a sample is almost surely suboptimal.
+The idea is that we choose the version of our theory that best fits reality.
+However, we are forced to rely on a limited sample of reality.
+We are misled because these two factors, induction and limited sample size, interact.
+Choosing the best version of our theory based on a sample is almost surely suboptimal.
 Therefore, the observed loss is an overconfident estimate of the loss in the future and closeness to the truth.
@@
 \end{equation}
-
-So the observed loss is an underestimation that has to be corrected for by considering the complexity of the model.
 \begin{equation}
@@
 The ability to adapt to data is independent of the specific data that was observed.
-Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
-Without knowing the inductive process, researchers cannot judge the overconfidence, so the inductive process ought to be made transparent.
+Complexity is a function of the model, i.e., $\mathcal{C}(\text{model})$, not of the data.
+Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
 
@@
 This observation is only useful if we know the extent of this optimism.
-However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and suggest two starting points for making this bias transparent.
-The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{model})$) and give us two starting points for making this bias transparent.
+The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
 Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.
@@
 Even a casual consideration of the above formalization should strike anyone who has ever worked with empirical data as unrealistic.
-This flexibility is often called researchers' degrees of freedom [@simmonsFalsePositivePsychologyUndisclosed2011].
-Using researchers' degrees of freedom opportunistically is an obvious problem [@yarkoniGeneralizabilityCrisis2022; @wichertsDegreesFreedomPlanning2016].
-However, no one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
+No one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
 The point is to set the goalpost and have a yardstick to measure how well we can judge empirical support.
@@
 
-As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
+As a first step, we can note that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
 Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
-If a researcher employs a linear model, the complexity is calculable parametrically.
-Even if it were not, the linear model could be fitted on a large set of other data to asses its expected inductive behavior nonparametrically.
-However, suppose the researcher reconsiders the model based on the results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
+If a researcher employs a linear model, the complexity is calculable.
+Even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
+However, suppose the researcher reconsiders the model based on the results and their internal model.
 For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
 In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
-The researcher cannot be asked to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
-However, it is known that the resulting linear model is more complex than a linear regression formally implies.
+We can not ask the researcher to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
+However, we know that the resulting linear model is more complex than a linear regression formally implies.
 
 $$
-\mathcal{C}(\text{Model}) = \mathcal{C}(\text{Model}_\text{formal}) + \mathcal{C}(\text{Model}_\text{informal})
+\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
 $$
 
-That is not to say that there is no basis for judging the informal induction.
-The inductive decision can be thought of as reasonable, thus found unlikely that just about any variable would be added if the data suggests it.
-Or the opposite, it might not seem well justified on theoretical grounds and deemed a purely data-driven decision, which implies higher complexity.
-What can be said, however, is that this judgment is debatable and, therefore, subjective.
+That is not to say that we have no basis for judging the informal induction.
+We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
+Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data-driven decision, which implies higher complexity.
+What one can say, however, is that this judgment is debatable and, therefore, subjective.
 
 It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
-This fact implies that for these situations, the complexity and, hence, the optimism bias cannot be fully quantified.
+This fact implies that for these situations, the complexity and, hence, the optimism bias can not be fully quantified.
 Though full transparency remains out of reach, researchers may do their best to provide a good basis for assessing complexity.
@@
 
-The imperative to enable proper judgment of theories is simple: induce only what is necessary and what is induced should, if possible, be done formally.
+The imperative to enable proper judgment of theories is simple: induce only what is necessary and what you induce should, if possible, be done formally.
 Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.
@@
 
-Therefore, there is a theoretical bound that limits transparency about the inductive process.
-Some things cannot be made transparent in principle because some induction happens informally, so the complexity cannot be estimated with certainty.
+Therefore, we have a theoretical bound that limits transparency about the inductive process.
+Some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
 However, there is another bound that restricts transparency further.
 That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
-To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both bounds have to be increased.
+To achieve transparency about the inductive process and ultimately about the empirical support of a theory, we must increase both bounds.
 
-In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bound).
-I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.
+In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
+I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
 
@@
 Such transparency can be achieved by computational reproducibility.
-Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992; @peikertReproducibleDataAnalysis2021; @peikertReproducibleResearchTutorial2021].
-However, through the lens of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.
+Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992].
+However, through the lens of transparency about the inductive process, we may further refine our view on computational reproducibility.
 
@@
 However, place two restrictions on the software.
-First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is an executable program, of which the source code may not be openly available.
+First, while able to compute results based on data, the software can not be understood by human researchers, e.g., because it is a binary program.
 Second, when the data set changes, the software does not work anymore.
 The first condition rules out that the software can be analyzed formally.
-The second condition prevents from changing the data and observing the results.
-Therefore, there are no means to assess the complexity of the inductive process and the empirical support cannot be judged.
-It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., it has to be ruled out one of the above conditions.
+The second condition prevents us from changing the data and observing the results.
+Therefore, we have no means to assess the complexity of the inductive process and can not judge the empirical support.
+It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.
 
-Suppose the second condition is relaxed; then there is a "black box" that is repeatable on other data.
-In that case, there still is transparency about the inductive process because the complexity of the inductive process can be estimated, i.e., using cross-validation.
+Suppose we relax the second condition; then we have a "black box" that is repeatable on other data.
+In that case, we still have transparency about the inductive process because we can estimate the complexity of the inductive process, i.e., using cross-validation.
 Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
-Therefore, the traditional definition of computational reproducibility has to be extended and further it is required that the process can be repeated on other similar data.
+Therefore, we must extend the traditional definition of computational reproducibility and at least also require that the process can be repeated on other similar data.
 
-Therefore, to provide transparency, computational reproducibility has to satisfy two requirements.
+Therefore, to provide transparency, computational reproducibility must satisfy two requirements.
 First, computational reproducibility must ensure that the same data lead to the same results.
@@
 Removing the human from the loop facilitates that the data can be easily substituted by similar data.
-Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the assesment of the complexity.
+Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the judgment of the complexity.
 
@@
 First, it must be unambiguous what results are generated by which inductive process.
-This can be ensured by employing dynamic document creation [@knuth1984literate; @xie2019; @xie2015; @Aust_papaja_Prepare_reproducible_2022].
+This can be ensured by employing dynamic document creation.
 Second, the version of the software that is used to generate the results must be known.
-This can be ensured by documenting the version using software management [@merkelDockerLightweightLinux2014; @wiebelsLeveragingContainersReproducible2021].
-Third, the version of the author's written code (and possibly text) must be tracked using version control systems [@chaconProGit2014; @vanlissaWORCSWorkflowOpen2021].
-Fourth, how the data relates to software and computing infrastructure must be managed by, e.g., workflow automation [@feldmanMakeProgramMaintaining1979; @kimImplementingGitHubActions2022; @kinsmanHowSoftwareDevelopers2021].
+This can be ensured by documenting the version using software management.
+Third, the version of the author's written code (and possibly text) must be tracked using version control systems.
+Fourth, how the data relates to software and computing infrastructure must be managed by, e.g., workflow automation.
 
@@
 Computational reproducibility enables transparency about formal inductive processes.
-However, informal inductive processes deprive of complete transparency.
+However, informal inductive processes deprive us of complete transparency.
 Therefore, it is prudent to replace informal induction with formal induction as much as possible.
@@
 A perfectly reproducible linear regression is, from the outside, indistinguishable from one that was cherry-picked from hundreds of possible regressions.
-However, the expectations regarding verisimilitude and future performance between the cherry-picked and the simple regression should vary substantially.
+However, our expectations regarding verisimilitude and future performance between the cherry-picked and the simple regression should vary substantially.
 
@@
 Researchers could simply try to explain the inductive process after the fact.
-This task is not easy because it needs to be known what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
+This task is not easy because we need to know what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
 This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.
@@
 Trusting a post hoc explanation of an inductive process is not entirely unreasonable.
-However, post hoc, formal and informal induction are indistinguishable unless the account is exhaustive.
+However, post hoc, formal and informal induction are indistinguishable unless the account is comprehensive.
 Again, not only comprehensive for the data that was observed but also for all possible data.
 This task is challenging to accomplish in practice.
-After all, researchers would have to publish the process that led to the induction, i.e., their thoughts and full mental model.
+After all, researchers would have to lay open the process that led to the induction, i.e., their thoughts and full mental model.
 Dissolving the distinction between formal and informal induction would mean surrendering the advantage a formal inductive process provides; instead of quantifying complexity, the complexity must be judged with considerable uncertainty.
@@
 Registration is a strong argument that the data has not influenced the results more than the complexity of the formal inductive process implies.
-If the data does not yet exist, it cannot influence the results.
+If the data does not yet exist, it can not influence the results.
 Note that registration is about the flow of information, i.e., the data used for induction, and not the temporal order.
@@
 They have to reason about all the possible data patterns they might encounter.
-While this makes registration so difficult, it is exactly the information we need and from which the complexity of the inductive process might be derived.
+While this makes registration so difficult, it is exactly the information we need and from which we may derive the complexity of the inductive process.
 
-Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extent.
+Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extend.
 In other words, researchers still might be very unsure how the data will look and therefore rely on induction to fill in the blanks.
@@
 In that case, a variable selection mechanism can be specified.
-Or suppose it is unclear what functional form to expect between predictor and outcome.
-In that case, many statistical models exist that allow almost arbitrary functional forms [@rissanenUniversalCodingInformation1984], e.g., smoothing splines [@cravenSmoothingNoisyData1978], random forests [@hoRandomSubspaceMethod1998], neural networks [@amariUniversalTheoremLearning1993].
+Or suppose it is unclear what functional form to expect.
+In that case, many statistical models exist that allow almost arbitrary functional forms.
 If the theory does not suffice to define an outlier, this decision can be made data-dependent, etc.
@@
 This version, with "mock" results based on simulated data, serves as the registration.
-When the data are gathered, the results are updated to reflect the actual observations.
+When the data is gathered, the results are updated to reflect the actual data.
 
@@
 Being able to account for uncertainty is indispensable for the question of how to deal with informal induction.
-If induction cannot be repeated at will nor formally analyzed, there is uncertainty about the complexity of the inductive process.
-Theoretical risk is conceptually related to the discussed concept of complexity but put into the context of Bayesian philosophy of science.
+If induction can not be repeated at will nor formally analyzed, we are uncertain about the complexity of the inductive process.
+Theoretical risk is conceptually related to the here discussed concept of complexity but put into the context of Bayesian philosophy of science.
 Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence, and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory if the evidence is more likely under the theory).
@@
 The extent of this overconfidence is a function of the inductive process.
-Specifically, I propose that transparency must concern the extent to which data may influence the results and which is captured in the term complexity.
+Specifically, I propose that transparency must concern the extent to which data may influence the results and term this complexity.
 I show that the complexity of an inductive process can be quantified if the process can be analyzed or repeated on similar data.
@@
 
-My consideration of transparency centers around the necessity to allow for induction to test imprecise theories.
-I find it unconvincing that with some more effort researchers will develop better and more precise theories from thin air.
-Instead, I find it more plausible that psychological sciences will refine theories using data to make testable statements [@brandmaierTheoryguidedExplorationStructural20161205] and account for the bias the induction entails.
+My consideration of transparency centers around the use of induction to test imprecise theories.
+A theory is imprecise because it must be refined using data to make testable statements.
 I conceptualize induction to derive testable statements as choosing among a set of implications of the theory.
 Such a set of possible implications and the process of choosing among them according to a loss function is what I understand as a model.
+However, this conceptualization may be criticized.
 It is not entirely clear how to chose the process of selecting the best fitting version of the theory.
@@
 Is the application of mathematically or computationally convenient methods, as is usual in statistical modeling, sufficient?
-Does the arbitrariness of the selection process needs to be accounted for?
+Must we account for the arbitrariness of the selection process?
 
@@
 A certain statistical method may imply a set of possible implications that only partially overlaps with the theory.
-Developing more powerful inductive algorithms is a very active field of research in the machine learning community [@brandmaierStructuralEquationModel2013].
+Developing more powerful inductive algorithms is a very active field of research in the machine learning community.
 However, having more and more powerful tools for induction simply means that more and more vague theories may be tested.
@@
 
-It might happen that in the end, more than one inductive process may be applicable, and no criterion makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
-I have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
+It might happen that in the end, more than one inductive process may be applicable, and no criterium makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
+We have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
 A possible direction for future research is to explore if forming the expectation over inductive processes is similarly fruitful.
@@
 However, the choice of loss function may be highly consequential for the appraisal of a theory.
-Not only is this choice consequential, but it is easily imaginable that researchers cannot agree on a metric to judge the theory.
+Not only is this choice consequential, but it is easily imaginable that researchers can not agree on a metric to judge the theory.
 
@@
 Allowing loss functions to vary between researchers also implies that the complexity of the inductive process varies.
-It follows that the correction for complexity cannot be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
+It follows that the correction for complexity can not be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
 This again highlights the importance of making the process transparent.
-If the process is transparent, the readers may change the loss function and calculate complexity accordingly.
+If the process is transparent, the readers might change the loss function and calculate complexity accordingly.
 
@@
 
-This dissertation proposes a narrow definition of transparency.
+This dissertation is concerned with a narrow definition of transparency.
 I have not addressed essential parts of the scientific process, among them:
@@
 How should they go about gathering data concerning the chosen theory?
-When are they ready to publish their conclusions?
-How should these conclusions be evaluated?
+When they are ready to publish their conclusions, how should these conclusions be evaluated?
 Which conclusions warrant publication?
@@
 However, I hope we might abandon the tradition of analyzing scientific practices only after they have been long established.
-Instead the aim should be to analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
+Instead we should analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
 That way, theoretical insight may shape application.

### 2023-02-15 manuscript.Rmd
personal out for adjective?

@@
 <!--necessity of induction-->
-In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
+In other words, induction gives scientific theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
 It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.

### 2023-02-15 manuscript.Rmd
can not
we arrive at -> the result is
since the personal narrative was exchanged widely I would do this here as well

@@
 If such assumptions do not hold, they may lead to empirical falsification even though the theory holds.
-However, if an auxiliary assumption is induced, it can not be falsified by the same data that induced it.
+However, if an auxiliary assumption is induced, it cannot be falsified by the same data that induced it.
 Inducing such assumptions, therefore, effectively removes auxiliary assumptions from the conjunction that is exposed to falsification.
@@
 If applied to auxiliary assumptions, such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
-If pushed to the limit, we arrive at a "theory" governed by the data.
+If pushed to the limit, the result is a "theory" governed by the data.
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.

### 2023-02-15 manuscript.Rmd
can not
and before hence is redundant; has to e > must

@@
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction can not refute these statements, so they have never been subject to an empirical test.
+The data used for induction cannot refute these statements, so they have never been subject to an empirical test.
 
@@
 Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
-The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
-Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?
+The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology hence, how well-tested the empirical claims, as reported in the literature, actually are.
+Therefore, the question of this dissertation is *what* has to be be made transparent, and *how* to best make it transparent?
 

### 2023-02-15 manuscript.Rmd
personal to generic

@@
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-When interested in how much information is lost when using $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+When interested in how much information is lost when using $g(x)$, the theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], this results:

### 2023-02-15 manuscript.Rmd
Update manuscript.Rmd

@@
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
-What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
+What is of interest is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
 

### 2023-02-15 manuscript.Rmd
passive; also facilitate > help

@@
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
-What is of interest is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
+What is of interest is not the derivation but how this conceptualization may facilitate to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
 

### 2023-02-17 manuscript.Rmd
change frontmatter, do not list legal matter, thank julia

--- 2023-02-15 manuscript.Rmd
@@
 
-\newpage
-\clearpage
-\thispagestyle{empty}
-\begin{center}
-\vspace*{\fill}
-For my mother.\vspace*{\fill}
-\end{center}
-\clearpage
+```{r child='frontmatter.md'}
+```
 
-\newpage
-\clearpage
-\thispagestyle{empty}
-# Acknowledgments {.unlisted .unnumbered}
-
-No one appreciated yet.
-\clearpage
-\thispagestyle{empty}
-\begin{center}
-\vspace*{\fill}
-Insight must precede application.
-\vspace*{\fill}
-\end{center}
-\newpage
-\clearpage
-
-\begin{spacing}{1}
-\tableofcontents
-\end{spacing}
-
-\newpage
-\clearpage
-\thispagestyle{empty}
-
-# Legal matters
+# Legal matters {.unlisted .unnumbered}
 
@@
 During the work on their dissertation, Aaron Peikert was a pre-doctoral fellow of the [International Max Planck Research School on Computational Methods in Psychiatry and Ageing Research (IMPRS COMP2PSYCH)](https://www.mps-ucl-centre.mpg.de/comp2psych) of
-the Max Planck Institute for Human Development, Berlin, Germany and University College London, United Kingdom.
+the Max Planck Institute for Human Development, Berlin, Germany, and University College London, United Kingdom.
 
-## Used resources and aids
+## Used resources and aids {.unlisted .unnumbered}
 
 I used several digital aids to improve the writing (not the content), including grammerly.com, deepl.com, and mentor.duden.de.
-Also, to aid with the clarity of the writing, I asked Andreas Brandmaier, Carli Ochs, Leo Richter, Leonie Hagitte, and Caroline Gahrmann for feedback.
+Also, to aid with the clarity of the writing, I asked Andreas Brandmaier, Carli Ochs, Leo Richter, Leonie Haggitte, and Caroline Gahrmann for feedback.
 Furthermore, I asked Leo Richter and Maximilian Ernst for help with mathematical notation and typesetting.
+Julia Delius has kindly agreed to proofread the journal articles at the time of submission and the synopsis of the thesis.
+All the errors that remain are solely my responsibility.
 I also prompted `ChatGPT` to write a song about reproducibility; while it was amusing, it was not funny enough to make an appearance in the text.
@@
 Psychology is a difficult science [@meehlTheoreticalRisksTabular1978].
-Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
-Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology that also begins to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
-We can roughly divide the proposed countermeasures into two categories, those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that aim to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the claim itself.
+Highlighting the difficulties and trying to overcome them are no recent trends, though they have been invigorated by the so-called replication crisis in psychology; a crisis that has also begun to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As psychology grapples with this crisis of confidence in its empirical findings, several causes have been identified and respective remedies have been proposed.
+The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
 <!--bla-transparency-bla-->
-In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory.
-Testing a theory empirically is often viewed as deductive since the theory is making statements about the data.
-Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
-I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results if unaccounted for.
-The extent of this bias depends on the inductive process.
-The inductive process, therefore, must be transparent to other researchers because they are otherwise unable to judge the empirical support of a theory.
+In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory [@leePopperFalsificationCorroboration2021].
+Testing a theory empirically is often viewed as deductive since the theory is making statements about future observations.
+Empirical scientists, however, often simultaneously engage in induction by deriving general statements from past observation.
+I will argue that the inductive element in empirical tests leads to overconfidence in the empirical results if unaccounted for.
+The extent of this bias depends on properties of the inductive process.
+The induction, therefore, must be transparent to other researchers in order for them to be able to judge the empirical support of a theory.
 
-The above distinction, between statistical and sociological countermeasures, arises from two different kinds of induction.
-On the one hand, we have inductive processes that are well-defined in the form of statistical methods; on the other hand, researchers also engage in more informal inductive behavior outside of well-defined models.
-These different kinds of induction require different countermeasures.
-Statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
-To understand why transparency is crucial, it is important to understand that formal and informal induction are integral to empirical sciences.
-Any science must be able to communicate how its knowledge is generated.
-However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive process that contributed to them.
-Therefore, transparency is more than a virtue that may improve knowledge gains in empirical sciences but is an indispensable property.
+The above distinction, between statistical and sociological countermeasures, arises from two sources of inductive bias.
+On the one hand, we have inductive processes that are well-defined in the form of statistical methods, while on the other, researchers also engage in more informal inductive behavior outside of well-defined models.
+These different kinds of inductive behavior require different countermeasures.
+When rigorously applied, statistical methods make the inductive bias quantifiable, while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
+To understand why transparency is crucial, it is important to comprehend how integral formal and informal induction are for empirical sciences.
+Any science must be able to communicate how it generates its knowledge.
+However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive processes involved.
+Therefore, transparency is more than a virtue that may improve empirical sciences somewhat but is an indispensable property.
 
 <!--induction is necessary for psychological science-->
-To function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
-In psychology, this is not a purely deductive endeavor.
-Very few psychological theories are precise enough to derive testable statements.
-While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
+In order to function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
+In psychology, this is not a purely deductive endeavor [@hitchcockPredictionAccommodationRisk2004; @meehlAppraisingAmendingTheories1990].
+Very few psychological theories are precise enough to derive testable statements [@friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020].
+While it is tempting to claim that a theory makes deductively testable claims [@leePopperFalsificationCorroboration2021] by, for example, implying a mean difference between two groups, such a statement is not testable on its own.
 Making inferences about a mean difference requires knowledge about the variance.
-Either the variance is known, or it has to be estimated from data.
-Consider a placebo and an experimental group, where we hypothesize and test a mean difference using a simple t-test.
-The decision to reject the null hypothesis (groups have equal means) depends, besides true mean difference and sample size, on the observed variance.
+Either the variance is known, than a purely deductive test is possible, or it has to be estimated from data, making induction part of the empirical test.
+Consider a placebo and an experimental group; there is possible to hypothesize and test a mean difference using a simple t-test.
+The decision to reject the null hypothesis (groups have equal means) depends on, besides true mean difference and sample size, the observed variance.
 However, the variance needs to be induced from the data.
@@
 <!--sometimes the use of induction is well justified-->
-Induction is necessary and, in the present case, innocuous for psychology as a science.
+Induction is necessary and, in the present case, harmless for psychology as a science.
 In this case, it is innocuous because the bias from inducing the variance can be accounted for.
-A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
 z≈`r round(qnorm(.99), 2)`).
-It is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
+It has been widely known for decades that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908], i.e., the inductive bias vanishes with increasing sample size.
 Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would be no psychological theory that could be subjected to an empirical test.
+If they had to, there would be currently very few psychological theory that could be subjected to an empirical test [@muthukrishnaProblemTheory2019].
 
 <!--necessity of induction-->
-In other words, induction gives scientific theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
-It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.
+In other words, induction gives theories some leeway to be imprecise and contain "blank" spaces, filled in later based on observation.
+It allows researchers to focus on the essential statements of their theories and choose to leave some parts (such as auxiliary assumptions) to be determined inductively.
 In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
@@
 If such assumptions do not hold, they may lead to empirical falsification even though the theory holds.
-However, if an auxiliary assumption is induced, it cannot be falsified by the same data that induced it.
+However, if an auxiliary assumption is induced, it can not be falsified by the same data that induced it.
 Inducing such assumptions, therefore, effectively removes auxiliary assumptions from the conjunction that is exposed to falsification.
@@
 If applied to auxiliary assumptions, such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
-If pushed to the limit, the result is a "theory" governed by the data.
+If pushed to the limit, we arrive at a "theory" governed by the data.
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction cannot refute these statements, so they have never been subject to an empirical test.
+The data used for induction can not refute these statements, so they have never been subject to an empirical test.
 
@@
 <!--transparency about induction is required-->
-So empirical researchers find themselves in a pinch.
+So empirical researchers find themselves in a predicament.
 On the one hand, they need induction to test their imprecise theories.
@@
 Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
-The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology hence, how well-tested the empirical claims, as reported in the literature, actually are.
-Therefore, the question of this dissertation is *what* has to be be made transparent, and *how* to best make it transparent?
+The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
+Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?
 
@@
 It is addressed in the synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
-Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).
+Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge in current research practice (sociological factors).
 
 <!--how it relates to the articles-->
-Based on this distinction, the articles answer the second question (the *how*).
+Based on this distinction, my research articles that are part of this dissertation answer the second question (the *how*).
 I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
@@
 I do not attempt to rehash the debate about what constitutes an empirical test.
-However, I aim to emphasize the role of transparency in two frameworks that lend themselves to investigate induction.
-The first framework promotes transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
-The second framework promotes transparency under a science that wants to select a theory according to its expected predictive performance.
+However, I aim to examine the role of transparency in two frameworks that lend themselves to investigate the sources of inductive bias.
+The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
+The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.
 
-Both frameworks show how unaccounted induction leads to overconfidence in empirical results and suggest some theoretical tools to asses and control this overconfidence.
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools on how to disclose induction so that this overconfidence can be assessed and controlled.
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
@@
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-When interested in how much information is lost when using $g(x)$, the theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
-Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], this results:
+The quantity of interst is how much information is lost if we use $g(x)$, our theory as description of the world..., instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
 <!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
@@
 
-Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
+Many readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
@@
 
-This setup is highly theoretical.
-$\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
-However, this fact does not impede us from getting closer to the truth because we can still compare the two theories relative to each other.
+Note that in practice, $\mathcal{L}_{KL}(f, g)$ is unknowable, since the full truth is unobservable.
+However, this fact does not impede us from getting closer to the truth because we can still compare the competing theories relative to each other.
 Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
@@
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
+This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.
@@
 The ability to adapt to data is independent of the specific data that was observed.
-Complexity is a function of the model, i.e., $\mathcal{C}(\text{model})$, not of the data.
-Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.
+Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
+Without knowing the inductive process, researchers cannot judge the overconfidence, so the inductive process ought to be made transparent.
 
@@
 This observation is only useful if we know the extent of this optimism.
-However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{model})$) and give us two starting points for making this bias transparent.
-The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
+However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and suggest two starting points for making this bias transparent.
+The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
 Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.
@@
 
-As a first step, we can note that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
+As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
 Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
-If a researcher employs a linear model, the complexity is calculable.
-Even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
-However, suppose the researcher reconsiders the model based on the results and their internal model.
+If a researcher employs a linear model, the complexity is calculable parametrically.
+Even if it were not, the linear model could be fitted on a large set of other data to asses its expected inductive behavior nonparametrically.
+However, suppose the researcher reconsiders the model based on the results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
 For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
 In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
-We can not ask the researcher to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
-However, we know that the resulting linear model is more complex than a linear regression formally implies.
+The researcher cannot be asked to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
+However, it is known that the resulting linear model is more complex than a linear regression formally implies.
 
 $$
-\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
+\mathcal{C}(\text{Model}) = \mathcal{C}(\text{Model}_\text{formal}) + \mathcal{C}(\text{Model}_\text{informal})
 $$
 
-That is not to say that we have no basis for judging the informal induction.
-We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
-Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data-driven decision, which implies higher complexity.
-What one can say, however, is that this judgment is debatable and, therefore, subjective.
+That is not to say that there is no basis for judging the informal induction.
+The inductive decision can be thought of as reasonable, thus found unlikely that just about any variable would be added if the data suggests it.
+Or the opposite, it might not seem well justified on theoretical grounds and deemed a purely data-driven decision, which implies higher complexity.
+What can be said, however, is that this judgment is debatable and, therefore, subjective.
 
 It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
-This fact implies that for these situations, the complexity and, hence, the optimism bias can not be fully quantified.
+This fact implies that for these situations, the complexity and, hence, the optimism bias cannot be fully quantified.
 Though full transparency remains out of reach, researchers may do their best to provide a good basis for assessing complexity.
@@
 
-Therefore, we have a theoretical bound that limits transparency about the inductive process.
-Some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
+Therefore, there is a theoretical bound that limits transparency about the inductive process.
+Some things cannot be made transparent in principle because some induction happens informally, so the complexity cannot be estimated with certainty.
 However, there is another bound that restricts transparency further.
 That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
-To achieve transparency about the inductive process and ultimately about the empirical support of a theory, we must increase both bounds.
+To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both bounds have to be increased.
 
-In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
-I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.
+In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bound).
+I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.
 
@@
 Such transparency can be achieved by computational reproducibility.
-Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992].
-However, through the lens of transparency about the inductive process, we may further refine our view on computational reproducibility.
+Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992; @peikertReproducibleDataAnalysis2021; @peikertReproducibleResearchTutorial2021].
+However, through the lens of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.
 
@@
 However, place two restrictions on the software.
-First, while able to compute results based on data, the software can not be understood by human researchers, e.g., because it is a binary program.
+First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is an executable program, of which the source code may not be openly available.
 Second, when the data set changes, the software does not work anymore.
 The first condition rules out that the software can be analyzed formally.
-The second condition prevents us from changing the data and observing the results.
-Therefore, we have no means to assess the complexity of the inductive process and can not judge the empirical support.
-It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.
+The second condition prevents from changing the data and observing the results.
+Therefore, there are no means to assess the complexity of the inductive process and the empirical support cannot be judged.
+It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., it has to be ruled out one of the above conditions.
 
-Suppose we relax the second condition; then we have a "black box" that is repeatable on other data.
-In that case, we still have transparency about the inductive process because we can estimate the complexity of the inductive process, i.e., using cross-validation.
+Suppose the second condition is relaxed; then there is a "black box" that is repeatable on other data.
+In that case, there still is transparency about the inductive process because the complexity of the inductive process can be estimated, i.e., using cross-validation.
 Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
 If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
-Therefore, we must extend the traditional definition of computational reproducibility and at least also require that the process can be repeated on other similar data.
+Therefore, the traditional definition of computational reproducibility has to be extended and further it is required that the process can be repeated on other similar data.
 
-Therefore, to provide transparency, computational reproducibility must satisfy two requirements.
+Therefore, to provide transparency, computational reproducibility has to satisfy two requirements.
 First, computational reproducibility must ensure that the same data lead to the same results.
@@
 Trusting a post hoc explanation of an inductive process is not entirely unreasonable.
-However, post hoc, formal and informal induction are indistinguishable unless the account is comprehensive.
+However, post hoc, formal and informal induction are indistinguishable unless the account is exhaustive.
 Again, not only comprehensive for the data that was observed but also for all possible data.
 This task is challenging to accomplish in practice.
-After all, researchers would have to lay open the process that led to the induction, i.e., their thoughts and full mental model.
+After all, researchers would have to publish the process that led to the induction, i.e., their thoughts and full mental model.
 Dissolving the distinction between formal and informal induction would mean surrendering the advantage a formal inductive process provides; instead of quantifying complexity, the complexity must be judged with considerable uncertainty.
@@
 They have to reason about all the possible data patterns they might encounter.
-While this makes registration so difficult, it is exactly the information we need and from which we may derive the complexity of the inductive process.
+While this makes registration so difficult, it is exactly the information we need and from which the complexity of the inductive process might be derived.
 
-Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extend.
+Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extent.
 In other words, researchers still might be very unsure how the data will look and therefore rely on induction to fill in the blanks.
@@
 
-My consideration of transparency centers around the use of induction to test imprecise theories.
-A theory is imprecise because it must be refined using data to make testable statements.
+My consideration of transparency centers around the necessity to allow for induction to test imprecise theories.
+I find it unconvincing that with some more effort researchers will develop better and more precise theories from thin air.
+Instead, I find it more plausible that psychological sciences will refine theories using data to make testable statements [@brandmaierTheoryguidedExplorationStructural20161205] and account for the bias the induction entails.
 I conceptualize induction to derive testable statements as choosing among a set of implications of the theory.
 Such a set of possible implications and the process of choosing among them according to a loss function is what I understand as a model.
-However, this conceptualization may be criticized.
 It is not entirely clear how to chose the process of selecting the best fitting version of the theory.
@@
 
-It might happen that in the end, more than one inductive process may be applicable, and no criterium makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
-We have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
+It might happen that in the end, more than one inductive process may be applicable, and no criterion makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
+I have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
 A possible direction for future research is to explore if forming the expectation over inductive processes is similarly fruitful.
@@
 Allowing loss functions to vary between researchers also implies that the complexity of the inductive process varies.
-It follows that the correction for complexity can not be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
+It follows that the correction for complexity cannot be done by the authors for all their readers universally if the loss function and complexity may depend on the researcher assessing the theory.
 This again highlights the importance of making the process transparent.
-If the process is transparent, the readers might change the loss function and calculate complexity accordingly.
+If the process is transparent, the readers may change the loss function and calculate complexity accordingly.
 

### 2023-02-17 manuscript.Rmd
first round of changes from Julia

@@
 bibliography: references.bib
-abstract: "`r paste0(c('\\\\setstretch{1.0}English:', readLines('abstract.md'), '\\\\par Deutsch:\\\\textgerman{', readLines('zusammenfassung.md'), '}\\\\newpage'), collapse = ' ')`"
+abstract: "`r paste0(c('\\\\setstretch{1.0}\\\\textsc{Abstract:}', readLines('abstract.md'), '\\\\vspace{1mm}\\\\par \\\\textsc{Zusammenfassung:}\\\\textgerman{', readLines('zusammenfassung.md'), '}\\\\newpage'), collapse = ' ')`"
 csl: apa.csl
@@
 Highlighting the difficulties and trying to overcome them are no recent trends, though they have been invigorated by the so-called replication crisis in psychology; a crisis that has also begun to ripple through other empirical sciences [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As psychology grapples with this crisis of confidence in its empirical findings, several causes have been identified and respective remedies have been proposed.
-The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+As psychology grapples with this crisis of confidence in its empirical findings, several causes have been identified and respective remedies have been suggested.
+The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g.; @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
@@
 Testing a theory empirically is often viewed as deductive since the theory is making statements about future observations.
-Empirical scientists, however, often simultaneously engage in induction by deriving general statements from past observation.
+Empirical scientists, however, often simultaneously engage in induction by deriving general statements from past observations.
 I will argue that the inductive element in empirical tests leads to overconfidence in the empirical results if unaccounted for.
@@
 
-The above distinction, between statistical and sociological countermeasures, arises from two sources of inductive bias.
+The above distinction between statistical and sociological countermeasures arises from two sources of inductive bias.
 On the one hand, we have inductive processes that are well-defined in the form of statistical methods, while on the other, researchers also engage in more informal inductive behavior outside of well-defined models.
@@
 However, transparency has an outstanding role in psychology and other empirical sciences because empirical statements lose their value without transparency about the inductive processes involved.
-Therefore, transparency is more than a virtue that may improve empirical sciences somewhat but is an indispensable property.
+Therefore, transparency is more than a virtue that may improve empirical sciences somewhat, rather it is an indispensable property.
 
@@
 Making inferences about a mean difference requires knowledge about the variance.
-Either the variance is known, than a purely deductive test is possible, or it has to be estimated from data, making induction part of the empirical test.
-Consider a placebo and an experimental group; there is possible to hypothesize and test a mean difference using a simple t-test.
-The decision to reject the null hypothesis (groups have equal means) depends on, besides true mean difference and sample size, the observed variance.
+Either the variance is known, then a purely deductive test is possible, or it has to be estimated from data, making induction part of the empirical test.
+Consider a placebo and an experimental group; there it is possible to hypothesize and test a mean difference using a simple t-test.
+The decision to reject the null hypothesis (groups have equal means) depends on the observed variance, besides true mean difference and sample size.
 However, the variance needs to be induced from the data.
@@
 In this case, it is innocuous because the bias from inducing the variance can be accounted for.
-A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs.
-z≈`r round(qnorm(.99), 2)`).
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is $t≈`r round(qt(.99, 10), 2)`$ vs.
+$z≈`r round(qnorm(.99), 2)`)$.
 It has been widely known for decades that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908], i.e., the inductive bias vanishes with increasing sample size.
-Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
-If they had to, there would be currently very few psychological theory that could be subjected to an empirical test [@muthukrishnaProblemTheory2019].
+Induction is also necessary because it is virtually impossible to ask psychologists to specify every detail, such as the variance, a priori from their theory.
+If they had to, there would currently be very little psychological theory that could be subjected to an empirical test [@muthukrishnaProblemTheory2019].
 
@@
 <!--statistical induction can be dangerous-->
-By the same token, whole theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
-If applied to auxiliary assumptions, such a strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
+By the same token, entire theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
+If applied to auxiliary assumptions, such a strategy of changing a theory post hoc in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
 If pushed to the limit, we arrive at a "theory" governed by the data.
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction can not refute these statements, so they have never been subject to an empirical test.
+The data used for induction can not refute these statements, so they have never been subjected to an empirical test.
 
 <!--the appraisal of "induced theories" should be low-->
-So what to think of such, yet untested, theory?
+So what is to be thought
 Researchers and philosophers of science differ considerably in their opinion about how to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
@@
 
-# What necessitates transparency
+# What makes transparency nessesary?
 
@@
 
-# How to establish transparency
+# How to establish transparency?
 
@@
 Even a casual consideration of the above formalization should strike anyone who has ever worked with empirical data as unrealistic.
-This flexibility is often called researchers' degrees of freedom [@simmonsFalsePositivePsychologyUndisclosed2011].
+This flexibility is often called researchers' degrees of freedom [@simmonsFalsepositivePsychologyUndisclosed2011].
 Using researchers' degrees of freedom opportunistically is an obvious problem [@yarkoniGeneralizabilityCrisis2022; @wichertsDegreesFreedomPlanning2016].
@@
 
-## Transparency about statistical models: Computational\ Reproducibilibiltiy
+## Transparency about statistical models: Computational\ Reproducibility
 
@@
 
+# References
+
+<div id="refs"></div>
+
+\clearpage\newpage
+
 # Articles
@@
 
-## A Reproducible Data Analysis Workflow With R Markdown, Git, Make, and Docker
+## A reproducible data analysis workflow With R\ Markdown, Git, Make, and Docker
 
@@
 
-## Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once
+## Reproducible research in R: A tutorial on how to do the same thing more than once
 
@@
 \includepdf[pages=-, offset=5mm 0mm]{papers/prereg.pdf}
-
-# References

### 2023-02-17 manuscript.Rmd
update dockerfile

@@
 repro:
+  packages:
+    - gert
+    - aaronpeikert/repro@5075336
   files:
+    - zusammenfassung.md
     - abstract.md
+    - frontmatter.md
+    - cover.pdf
+    - preamble.tex
     - apa.csl
@@
     - pdfpages
+    - lualatex-math
+    - lualatex-base
+    - titling
   apt:

### 2023-02-17 manuscript.Rmd
add version
typo

--- 2023-02-17 manuscript.Rmd
@@
     - pdfpages
+    - lualatex-math
+    - lualatexbase
+    - titling
   apt:
@@
 
+You are reading version ``r repro::current_hash(backend = "gert")``.
 This work is made available online under the following source under the terms of the [Creative Commons Universal Public Domain Dedication | CC0](https://creativecommons.org/publicdomain/zero/1.0/):  

### 2023-02-17 manuscript.Rmd
typo

--- 2023-02-17 manuscript.Rmd
@@
     - pdfpages
+    - lualatex-math
+    - luatexbase
+    - titling
   apt:

### 2023-02-17 manuscript.Rmd
add easteregg

@@
 All the errors that remain are solely my responsibility.
-I also prompted `ChatGPT` to write a song about reproducibility; while it was amusing, it was not funny enough to make an appearance in the text.
+[I also prompted `ChatGPT` to write a song about reproducibility; while it was amusing, it was not funny enough to make an appearance in the text](https://github.com/aaronpeikert/repro-talk/blob/master/gpt-songs.md).
 

### 2023-02-17 manuscript.Rmd
second batch of julia's suggestions

@@
 <!--the appraisal of "induced theories" should be low-->
-So what is to be thought
+So what is to be thought of such as yet untested theory?
 Researchers and philosophers of science differ considerably in their opinion about how to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p.
@@
 <!--transparency about induction is required-->
-So empirical researchers find themselves in a predicament.
+Empirical researchers thus find themselves in a predicament.
 On the one hand, they need induction to test their imprecise theories.
@@
 The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
-Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?
+Therefore, the questions of this dissertation are *what* must be made transparent, and *how* to best make it transparent?
 
@@
 The first question (the *what*) is theoretical in nature.
-It is addressed in the synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
+It is addressed in this synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
 Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge in current research practice (sociological factors).
@@
 I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
-The conceptualization of computational reproducibility and preregistration in regard of transparency is supplemented by practical guidance on how researchers may implement these approaches.
+The conceptualization of computational reproducibility and preregistration from a transparency perspective is supplemented by practical guidance on how researchers can implement these approaches.
 
@@
 The need for transparency is closely tied to the use of induction in the empirical test of a theory.
-There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5 Experience as a Method].
+There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5, Experience as a Method].
 I do not attempt to rehash the debate about what constitutes an empirical test.
@@
 
-Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools on how to disclose induction so that this overconfidence can be assessed and controlled.
+Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools to disclose induction so that this overconfidence can be assessed and controlled.
 Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
@@
 Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
-The quantity of interst is how much information is lost if we use $g(x)$, our theory as description of the world..., instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
+The quantity of interest is how much information is lost if we use $g(x)$, our theory as description of the world, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
 Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
@@
 
-Many readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch.
-7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
+Many readers will recognize that this information-theoretic setup and the derivation below closely follow @burnhamModelSelectionMultimodel2002, Chapter 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
 What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.
@@
 To allow quantities to be induced, we must assume that the theory is parameterized, e.g., $g(x|\theta)$.
-That means the theory implies a family of possible probability distributions that may describe reality.
+That means that the theory implies a family of possible probability distributions that may describe reality.
 This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
@@
 
-The crucial point is to understand what happens when $\theta$ cannot be derive deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
-Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
+The crucial point is to understand what happens when $\theta$ cannot be derived deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
+Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continuous).
 It follows, almost surely, that information is lost:
@@
 
-That is to say, any induced estimate will be sub-optimal.
+That is to say, any induced estimate will be suboptimal.
 The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
@@
 Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$, respectively.
-For more details about this derivation, see @burnhamModelSelectionMultimodel2002, Ch. 7.2.
+For more details about this derivation, see @burnhamModelSelectionMultimodel2002, Chapter 7.2.
 
 The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
-We may conclude that substituting deduced quantities with induced estimates leads to some overconfidence about how close one is to the truth.
+We may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
+This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, in how much detail the model may represent the data [@goodfellowCapacityOverfittingUnderfitting2016, Chapter.
+5.2; @mikkelsonComplexityVerisimilitudeRealism2001].
 I therefore denote it as as $\mathcal{C}$.
@@
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
-One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's Cp [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].
+Since $\theta_*$ is unknown, one condition is, that we know the properties of the inductive process that generated $\hat\theta$.
+We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's $C_p$ [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].
 
-## A future performance perspective
+## A future-performance perspective
 
@@
 Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
-In the future performance setup, we do not appeal to ground truth (I drop $\sim{}f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
+In the future-performance setup, we do not appeal to ground truth (I drop $\sim{}f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
 \begin{equation}
@@
 
-Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020] can be defined:
+Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020, Chapter 1.5.8] can be defined:
 \begin{equation}
@@
 
-However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
+However, what we are interested in is not how well the theory did on data that informed it, but on future, yet unseen, data:
 \begin{equation}
@@
 Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
-That is, we use cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
+That is, we use cross-validation where the data are partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
 For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
@@
 
-As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
+As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; this connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
 To make the link to the first approach even more clear, we may give an alternative definition of complexity as the expected difference between observed prediction error and expected prediction error [@hauensteinComputingAICBlackbox2018]:
@@
 
-Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
+Now that we have established the need for transparency about the inductive process, we can glance over of the more technical details to clarify what we have to make transparent.
 It bears repeating that simply revealing what has been done is not enough.
@@
 
-Where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
+where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
 To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
@@
 
-So the observed loss is an underestimation that has to be corrected for by considering the complexity of the model.
+The observed loss is thus an underestimation that has to be corrected for by considering the complexity of the model.
 \begin{equation}
@@
 Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
-The ability to adapt to data is independent of the specific data that was observed.
+The ability to adapt to data is independent of the specific data that were observed.
 Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
@@
 The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
-Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.
+Both approaches require researchers to make the inductive process transparent rather than merely publishing the results.
 
@@
 As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
-Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
+Formal discription, in the strict sense, means that the complexity can be mathematically derived, and in a looser sense, that the process can be repeated at will.
 If a researcher employs a linear model, the complexity is calculable parametrically.
-Even if it were not, the linear model could be fitted on a large set of other data to asses its expected inductive behavior nonparametrically.
-However, suppose the researcher reconsiders the model based on the results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
+Even if it were not calculable, the linear model could be fitted on a large set of other data to assess its expected inductive behavior nonparametrically.
+However, suppose the researcher reconsiders the model based on their results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
 For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
 In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
-The researcher cannot be asked to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
+The researcher cannot be asked to repeat the process on all possible data sets, nor is their behavior mathematically well-defined.
 However, it is known that the resulting linear model is more complex than a linear regression formally implies.
@@
 
-The imperative to enable proper judgment of theories is simple: induce only what is necessary and what is induced should, if possible, be done formally.
+The imperative to enable proper judgment of theories is simple: induce only what is necessary, and what is induced should, if possible, be done formally.
 Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.
 
-To summarise, I separate complexity $\mathcal{C}$ according to the transparency that can theoretically be achieved.
+To summarize, I separate complexity $\mathcal{C}$ according to the transparency that can theoretically be achieved.
 Formal inductive processes allow full transparency in the sense that $\mathcal{C}$ can be objectively quantified.
@@
 Therefore, there is a theoretical bound that limits transparency about the inductive process.
-Some things cannot be made transparent in principle because some induction happens informally, so the complexity cannot be estimated with certainty.
+Some things cannot be made transparent in principle because some induction happens informally, and the complexity cannot be estimated with certainty.
 However, there is another bound that restricts transparency further.
 That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
-To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both bounds have to be increased.
+To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both boundaries have to be increased.
 
-In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bound).
+In the following, I propose preregistration as a means to move induction into the formal domain (pushing the first boundary) and computational reproducibility to make formal induction transparent (pushing the second boundary).
 I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.
@@
 If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
-In other words, more than having formal induction is needed, the induction still must be made transparent in practice.
+In other words, formal induction is needed, but it still must be made transparent in practice.
 Such transparency can be achieved by computational reproducibility.
 Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992; @peikertReproducibleDataAnalysis2021; @peikertReproducibleResearchTutorial2021].
-However, through the lens of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.
+However, from the perspective of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.
 
-Consider software that delivers the same results upon input of the same data, i.e., it is reproducible.
+Consider software that delivers the same results upon input of the same data, i.e., it enables reproducible analyses.
 However, place two restrictions on the software.
@@
 The first condition rules out that the software can be analyzed formally.
-The second condition prevents from changing the data and observing the results.
+The second condition prevents changes of the data and observation of the results.
 Therefore, there are no means to assess the complexity of the inductive process and the empirical support cannot be judged.
-It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., it has to be ruled out one of the above conditions.
+It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., at least one of the above conditions needs to be ruled out..
 
@@
 In that case, there still is transparency about the inductive process because the complexity of the inductive process can be estimated, i.e., using cross-validation.
-Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
-If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
-Therefore, the traditional definition of computational reproducibility has to be extended and further it is required that the process can be repeated on other similar data.
+Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxation of the second condition, repeatability on other data, as well.
+If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on similar data.
+Therefore, the traditional definition of computational reproducibility has to be extended.
+Further, it is required that the process can be repeated on other similar data.
 
@@
 Implementing computational reproducibility by automating the process from data to results fulfills both conditions needed for transparency while reducing the effort that must be invested.
-Verifying that the results are actually produced by the inductive process and data is then a task that a computer may fulfil without human intervention.
+Verifying that the results are actually produced by the inductive process and data is then a task that a computer may fulfill without human intervention.
 Removing the human from the loop facilitates that the data can be easily substituted by similar data.
-Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the assesment of the complexity.
+Given enough computing power, automatic reproducibility allows one to scale the inductive process to many data sets and therefore enables the assessment of the complexity $\mathcal{C}$.
 
-While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process easily, it asks a lot from the researchers who create the work.
+While this workflow provides transparency and facilitates judgement of the inductive process by recipients of the work (readers, collaborators, editors, etc.), it asks a lot of the researchers who creating it.
 To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
-I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the there presented software `worcs`.
+I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the software `worcs` presented there.
 
@@
 Third, the version of the author's written code (and possibly text) must be tracked using version control systems [@chaconProGit2014; @vanlissaWORCSWorkflowOpen2021].
-Fourth, how the data relates to software and computing infrastructure must be managed by, e.g., workflow automation [@feldmanMakeProgramMaintaining1979; @kimImplementingGitHubActions2022; @kinsmanHowSoftwareDevelopers2021].
+Fourth, how the data relates to software and computing infrastructure must be managed by workflow automation, for example [@feldmanMakeProgramMaintaining1979; @kimImplementingGitHubActions2022; @kinsmanHowSoftwareDevelopers2021].
 
@@
 Computational reproducibility enables transparency about formal inductive processes.
-However, informal inductive processes deprive of complete transparency.
-Therefore, it is prudent to replace informal induction with formal induction as much as possible.
-If a researcher makes every effort only to employ formal inductive reasoning in a reproducible manner, they still must persuade their readers that they have not engaged in informal induction.
+However, informal inductive processes prevent complete transparency.
+Therefore, it is prudent to replace informal induction by formal induction wherever possible.
+Nonetheless, even if a researcher has made every effort only to employ formal inductive reasoning in a reproducible manner, they must still persuade their readers that they have not engaged in informal induction.
 
@@
 A perfectly reproducible linear regression is, from the outside, indistinguishable from one that was cherry-picked from hundreds of possible regressions.
-However, the expectations regarding verisimilitude and future performance between the cherry-picked and the simple regression should vary substantially.
+However, the expectations regarding verisimilitude and future performance should vary substantially between the cherry-picked and the simple regression.
 
@@
 Researchers could simply try to explain the inductive process after the fact.
-This task is not easy because it needs to be known what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
-This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.
-For researchers to judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data look different.
+This task is not easy because one needs to know what they did, as well as what they did not do, e.g., that they did not cherry-pick, and what they would have done had the data looked different.
+This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge its complexity.
+For researchers to judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data looked different.
 
@@
 However, post hoc, formal and informal induction are indistinguishable unless the account is exhaustive.
-Again, not only comprehensive for the data that was observed but also for all possible data.
+Again, the report should not only be comprehensive for the data that were observed but also for all possible data.
 This task is challenging to accomplish in practice.
@@
 
-A clear distinction between formal and informal can be provided by specifying the formal inductive process before any data is available.
-Such practice is called registration [@riceCurtailingUsePreregistration2019] or preregistration [@nosekPreregistrationRevolution2018]. 
-Registration is a strong argument that the data has not influenced the results more than the complexity of the formal inductive process implies.
-If the data does not yet exist, it cannot influence the results.
+A clear distinction between formal and informal induction can be provided by specifying the formal inductive process before any data is available.
+Such a practice is called registration [@riceCurtailingUsePreregistration2019] or preregistration [@nosekPreregistrationRevolution2018]. 
+Registration provides a strong argument that the data have not influenced the results more than the complexity of the formal inductive process implies.
+If the data do not yet exist, they cannot influence the results ($\mathcal{C} = 0)$.
 Note that registration is about the flow of information, i.e., the data used for induction, and not the temporal order.
@@
 Having no access to the data when specifying the analysis forces the researcher to think about the process of induction.
-They have to reason about all the possible data patterns they might encounter.
+They have to reason about all possible data patterns they might encounter.
 While this makes registration so difficult, it is exactly the information we need and from which the complexity of the inductive process might be derived.
@@
 Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extent.
-In other words, researchers still might be very unsure how the data will look and therefore rely on induction to fill in the blanks.
+In other words, researchers may still be very unsure how the data will look and therefore rely on induction to fill in the blanks.
 However, they must be clear about what they are going to induce.
@@
 Several strategies reflect such uncertainties.
-For example, suppose it is unknown which variables from a broad set of variables are involved.
+For example, suppose it is unknown which variables from a broad set of variables are involved in a process.
 In that case, a variable selection mechanism can be specified.
@@
 In that case, many statistical models exist that allow almost arbitrary functional forms [@rissanenUniversalCodingInformation1984], e.g., smoothing splines [@cravenSmoothingNoisyData1978], random forests [@hoRandomSubspaceMethod1998], neural networks [@amariUniversalTheoremLearning1993].
-If the theory does not suffice to define an outlier, this decision can be made data-dependent, etc.
+If the theory does not suffice to define an outlier, this decision can be made depending on the data, etc.
 
@@
 First, how can this be accomplished in a way that is practical and fits into the well-established practices of the scientific community?
-Second, how to account for informal induction, e.g., when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable for other reasons?
+Second, how does one account for informal induction, e.g., when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable for other reasons?
 
@@
 This version, with "mock" results based on simulated data, serves as the registration.
-When the data are gathered, the results are updated to reflect the actual observations.
+When the data have been collected, the results are updated to reflect the actual observations.
 
-To answer the second question of how to account for informal induction, I investigate the objective of registration and how it may be separated from the objective of confirmatory science in XXX.
-Particularly, I propose to formalize the objective of registration as reduction of uncertainty about theoretical risk.
+To answer the second question of how to account for informal induction, I investigate the objective of registration and how it may be separated from the objective of confirmatory science in @peikertWhyDoesPreregistration2023.
+In particular, I propose to formalize the objective of registration as a reduction of uncertainty about theoretical risk.
 Being able to account for uncertainty is indispensable for the question of how to deal with informal induction.
@@
 Theoretical risk is conceptually related to the discussed concept of complexity but put into the context of Bayesian philosophy of science.
-Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence, and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory if the evidence is more likely under the theory).
+Specifically, instead of the arbitrary loss functions considered here, the loss functions in @peikertWhyDoesPreregistration2023 are restricted to those that apply to binary evidence, and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory if the evidence is more likely under the theory).
 For this class of loss functions, I show how to account for uncertainty caused by informal induction and that a reduction in uncertainty due to preregistration is universally beneficial for those loss functions.
@@
 The extent of this overconfidence is a function of the inductive process.
-Specifically, I propose that transparency must concern the extent to which data may influence the results and which is captured in the term complexity.
+Specifically, I propose that transparency must concern the extent to which data may influence the results and which is captured by the term complexity.
 I show that the complexity of an inductive process can be quantified if the process can be analyzed or repeated on similar data.
-I call such an inductive process "formal".
+I call such an inductive process "formal."
 However, researchers often engage in informal induction, where complexity must be judged subjectively and with some uncertainty.
 Second, I explain *how* computational reproducibility and preregistration provide transparency about complexity.
-Preregistration allows to distinguish between formal and informal induction, while computational reproducibility communicates the process of formal induction.
+Preregistration allows a distinction between formal and informal induction, while computational reproducibility communicates the process of formal induction.
 
 This dissertation advances the theory and application of registration and computational reproducibility.
-Based on theoretical consideration, I refined the requirements of computational reproducibility, so that not only the same data must produce the same results but also that the inductive process may be repeated on other similar data.
-I developed a workflow for computational reproducibility that satisfies these requirements and built tooling to apply it.
+Based on theoretical considerations, I refined the requirements of computational reproducibility, so that not only the same data must produce the same results but also that the inductive process may be repeated on other similar data.
+I developed a workflow for computational reproducibility that satisfies these requirements and built tools to apply it.
 Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration and developed the reduction of uncertainty about the theoretical risk as a formal objective of registration.
@@
 My consideration of transparency centers around the necessity to allow for induction to test imprecise theories.
-I find it unconvincing that with some more effort researchers will develop better and more precise theories from thin air.
+I find it unconvincing that with extra effort, researchers will develop remarkably better and more precise theories from thin air.
 Instead, I find it more plausible that psychological sciences will refine theories using data to make testable statements [@brandmaierTheoryguidedExplorationStructural20161205] and account for the bias the induction entails.
@@
 Such a set of possible implications and the process of choosing among them according to a loss function is what I understand as a model.
-It is not entirely clear how to chose the process of selecting the best fitting version of the theory.
+It is not entirely clear how to select the process of selecting the best fitting version of the theory.
 Should the process of induction depend on the theory?
@@
 
-Implicit in the conceptualization of complexity was the assumption of unbiasedness (the inductive process converges towards $\theta_*$), i.e., an ideal inductive process.
+In the conceptualization of complexity, the assumption of unbiasedness was implicit (the inductive process converges towards $\theta_*$), i.e., representing an ideal inductive process.
 Such an assumption implies a perfect match of the set of implications of the theories and the statistical method.
 Furthermore, it assumes a certain optimality of the statistical method.
-Such optimality requirement is difficult to verify.
+Such an optimality requirement is difficult to verify.
 Most proofs regarding optimality make restrictive assumptions, i.e., optimality depends on the data and relies on large sample behavior.
@@
 In practice, researchers are restricted by the statistical models that are available to them.
-A certain statistical method may imply a set of possible implications that only partially overlaps with the theory.
+A certain statistical method may imply a set of possible implications that only partially overlap with the theory.
 Developing more powerful inductive algorithms is a very active field of research in the machine learning community [@brandmaierStructuralEquationModel2013].
-However, having more and more powerful tools for induction simply means that more and more vague theories may be tested.
+However, having increasingly powerful tools for induction simply means that ever more vague theories may be tested.
 It does not necessarily lead to the development of statistical methods that fit the theories researchers are interested in.
 
-It might happen that in the end, more than one inductive process may be applicable, and no criterion makes one preferable over the other, i.e., from the point of the researcher, the decision is arbitrary.
+In the end, more than one inductive process may perhaps be applicable, and no criterion makes one preferable over the other, i.e., from the perspective of the researcher, the decision is arbitrary.
 I have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
-A possible direction for future research is to explore if forming the expectation over inductive processes is similarly fruitful.
+A possible direction for future research is to explore wether forming the expectation over inductive processes is similarly fruitful.
 One suggestion in such a direction is a method called multiverse analysis [@steegenIncreasingTransparencyMultiverse2016], where many different data analytic decisions are systematically explored.
@@
 One important example of loss functions that are necessarily subjective are prior beliefs.
-So even researchers who agree on a general class of metrics, e.g., the posterior probability of a theory, would appraise theories differently.
+Thus, even researchers who agreed on a general class of metrics, e.g., the posterior probability of a theory, would appraise theories differently.
 Allowing loss functions to vary between researchers also implies that the complexity of the inductive process varies.
@@
 How should researchers decide which theory to test?
-How should they go about gathering data concerning the chosen theory?
+How should they go about collecting data concerning the chosen theory?
 When are they ready to publish their conclusions?
@@
 Which conclusions warrant publication?
-Of those, which warrants widespread attention of the scientific community?
+Of those, which warrant widespread attention of the scientific community?
 Which must be further substantiated?
 
-I expect that the conceptual frameworks I have employed to address the question of transparency about the inductive process may be enormously fruitful to apply there as well.
+I expect that the conceptual frameworks I have employed to address the question of transparency about the inductive process may also be enormously fruitful in that context.
 Applying statistical theory to the philosophy of science has a long tradition.
-However, I hope we might abandon the tradition of analyzing scientific practices only after they have been long established.
+However, I hope we might abandon the tradition of only analyzing scientific practices long after they have been established.
 Instead the aim should be to analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
@@
 
-In the following, I reprinted the articles published as part of the dissertation as they were made accessible by the journals, as allowed by the respective licenses.
+In the following, I have reprinted the articles published as part of the dissertation as they were made accessible by the journals, as allowed by the respective licenses.
 Note that the following articles were written collaboratively with the indicated coauthors.

### 2023-02-17 manuscript.Rmd
cite myself more

@@
 Based on theoretical considerations, I refined the requirements of computational reproducibility, so that not only the same data must produce the same results but also that the inductive process may be repeated on other similar data.
-I developed a workflow for computational reproducibility that satisfies these requirements and built tools to apply it.
-Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration and developed the reduction of uncertainty about the theoretical risk as a formal objective of registration.
+I developed a workflow for computational reproducibility that satisfies these requirements [@peikertReproducibleResearchTutorial2021; @peikertReproducibleDataAnalysis2021] and built tools to apply it [@vanlissaWORCSWorkflowOpen2021; @R-repro].
+Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration [@peikertReproducibleResearchTutorial2021] and developed the reduction of uncertainty about the theoretical risk as a formal objective of registration [@peikertWhyDoesPreregistration2023].
 

### 2023-02-19 manuscript.Rmd
fix citations

@@
 This overconfidence is directly related to how much induction a model entails.
-This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, in how much detail the model may represent the data [@goodfellowCapacityOverfittingUnderfitting2016, Chapter.
+This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, in how much detail the model may represent the data [@goodfellowDeepLearning2016, Chapter.
 5.2; @mikkelsonComplexityVerisimilitudeRealism2001].

### 2023-02-19 manuscript.Rmd
increase bindingoffset

@@
 # 3.81 to 5.08cm or 45-90 characters or 2-3 alphabets
-geometry: "left=4cm, right=4cm, top=3.5cm, bottom=4.5cm, bindingoffset=5mm"
+geometry: "left=4cm, right=4cm, top=3.5cm, bottom=4.5cm, bindingoffset=7mm"
 classoption: twoside
@@
 
-\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/workflow.pdf}
+\includepdf[pages=-, offset=7mm 0mm,frame=false]{papers/workflow.pdf}
 
@@
 
-\includepdf[pages=-, offset=5mm 0mm]{papers/tutorial.pdf}
+\includepdf[pages=-, offset=7mm 0mm]{papers/tutorial.pdf}
 
@@
 
-\includepdf[pages=-, offset=5mm 0mm]{papers/prereg.pdf}
+\includepdf[pages=-, offset=7mm 0mm]{papers/prereg.pdf}

### 2023-02-19 manuscript.Rmd

--- 2023-02-19 manuscript.Rmd
@@
 Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
-The data used for induction can not refute these statements, so they have never been subjected to an empirical test.
+The data used for induction cannot refute these statements, so they have never been subjected to an empirical test.
 

### 2023-02-19 manuscript.Rmd
spotted one more can not
replace url with zenodo doi

@@
 
-You are reading version ``r repro::current_hash(backend = "gert")``.
-This work is made available online under the following source under the terms of the [Creative Commons Universal Public Domain Dedication | CC0](https://creativecommons.org/publicdomain/zero/1.0/):  
-<https://aaronpeikert.github.io/thesis/manuscript.pdf>
+This work (version ``r repro::current_hash(backend = "gert")``) is made available online under the following source under the terms of the [Creative Commons Universal Public Domain Dedication | CC0](https://creativecommons.org/publicdomain/zero/1.0/):
+<https://doi.org/10.5281/zenodo.7654989>
 
@@
 For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
-In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
+In that case, the complexity cannot be formally judged because the researcher is an intractable part of the inductive process.
 The researcher cannot be asked to repeat the process on all possible data sets, nor is their behavior mathematically well-defined.

### 2023-02-20 manuscript.Rmd
set margins for frontmatter/body differently

@@
 # 3.81 to 5.08cm or 45-90 characters or 2-3 alphabets
-geometry: "left=4cm, right=4cm, top=3.5cm, bottom=4.5cm, bindingoffset=7mm"
+# asymetric geometry is set after/in frontmatter.md
+geometry: "left=4cm, right=4cm, top=3cm, bottom=5cm, bindingoffset=7mm"
 classoption: twoside

### 2023-02-20 manuscript.Rmd
let articles begin on right page
ammend license note

@@
 
-This work (version ``r repro::current_hash(backend = "gert")``) is made available online under the following source under the terms of the [Creative Commons Universal Public Domain Dedication | CC0](https://creativecommons.org/publicdomain/zero/1.0/):
+This work (version ``r repro::current_hash(backend = "gert")``) is made available online under the following source under the terms of the [Creative Commons Universal Public Domain Dedication | CC0](https://creativecommons.org/publicdomain/zero/1.0/), unless noted otherwise:
 <https://doi.org/10.5281/zenodo.7654989>
@@
 
-\clearpage\newpage
+\pagenumbering{gobble}
+\cleardoublepage\newpage
 
@@
 
+\cleardoublepage
 \newpage
@@
 
+\cleardoublepage
+
 \includepdf[pages=-, offset=7mm 0mm,frame=false]{papers/workflow.pdf}
+
+\cleardoublepage
 
@@
 
+\cleardoublepage
+
 \includepdf[pages=-, offset=7mm 0mm]{papers/tutorial.pdf}
+
+\cleardoublepage
 
@@
 
+\cleardoublepage
+
 \includepdf[pages=-, offset=7mm 0mm]{papers/prereg.pdf}

### 2023-02-21 manuscript.Rmd
add lilly's corrections

@@
 As psychology grapples with this crisis of confidence in its empirical findings, several causes have been identified and respective remedies have been suggested.
-The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g.; @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
+The proposed countermeasures can be broadly categorized into those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that are designed to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].
 
@@
 In this case, it is innocuous because the bias from inducing the variance can be accounted for.
-A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10, is $t≈`r round(qt(.99, 10), 2)`$ vs.
+A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the threshold of significance is higher for a t-test (the critical value for a one-sided test with α-level set to 1% and degrees of freedom set to 10 is $t≈`r round(qt(.99, 10), 2)`$ vs.
 $z≈`r round(qnorm(.99), 2)`)$.
@@
 
-# What makes transparency nessesary?
+# What makes transparency necessary?
 
@@
 \mathcal{L}_{KL}(f, g) &=\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x\\
-&= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]\label{eq:klexp}
+&= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]\label{eq:klexp}\,.
 \end{align}
@@
 Note that in practice, $\mathcal{L}_{KL}(f, g)$ is unknowable, since the full truth is unobservable.
-However, this fact does not impede us from getting closer to the truth because we can still compare the competing theories relative to each other.
+However, this fact does not impede us from getting closer to the truth, because we can still compare the competing theories relative to each other.
 Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
@@
 \begin{equation}
-\mathbb{E}_{X \sim f}\left[\log(g(x))\right]
+\mathbb{E}_{X \sim f}\left[\log(g(x))\right]\,.
 \end{equation}
@@
 \begin{equation}
-\theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))
+\theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))\,.
 \end{equation}
@@
 \begin{equation}
-\mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))
+\mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))\,,
 \end{equation}
-Or ignoring the constant:
+or
 \begin{equation}
-\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
+\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))]\,,
 \end{equation}
+ignoring the constant.
 
@@
 The observed likelihood is, therefore, often available, e.g., in maximum likelihood estimation.
-With slight abuse of notation, $\log(g(y|\theta)) \equiv \sum_{i = 1}^n \log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}\log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
+With slight abuse of notation, let $\log(g(y|\theta)) \equiv \sum_{i = 1}^n \log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}\log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
 
-The expectation over the data together with Taylor series expansion yields:
+The expectation over the data together with Taylor series expansion yields
 \begin{equation}
-\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]\,,
 \end{equation}
-
-Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$, respectively.
+where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$, respectively.
 For more details about this derivation, see @burnhamModelSelectionMultimodel2002, Chapter 7.2.
@@
 5.2; @mikkelsonComplexityVerisimilitudeRealism2001].
-I therefore denote it as as $\mathcal{C}$.
+I therefore denote it as as $\mathcal{C}$, i.e.,
 \begin{equation}
-\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
+\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}\,.
 \end{equation}
@@
 Fortunately, it is possible to approximate the complexity of a model under some conditions.
-Since $\theta_*$ is unknown, one condition is, that we know the properties of the inductive process that generated $\hat\theta$.
-We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's $C_p$ [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].
+Since $\theta_*$ is unknown, one condition is that we know the properties of the inductive process that generated $\hat\theta$.
+We can then formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$.
+Corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's $C_p$ [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].
 
@@
 \begin{equation}
-\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
+\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))\,.
 \end{equation}
 
-Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020, Chapter 1.5.8] can be defined:
+Again, the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020, Chapter 1.5.8] can be defined:
 \begin{equation}
 \mathbb{E}_y L(y, g(y|\hat\theta(y)))
- = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\hat\theta(y)))
+ = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\hat\theta(y)))\,.
 \end{equation}
-
 However, what we are interested in is not how well the theory did on data that informed it, but on future, yet unseen, data:
 \begin{equation}
-\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
+\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))\,.
 \end{equation}
-This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004].
+This expectation over what is often called training and test data is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004].
 Note that both training ($y$) and test ($x$) data vary in this expectation.
@@
 
-Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
+Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., use cross-validation.
 The former requires the derivatives with regard to the parameters, so I call it parametric estimator of complexity.
 The latter eschews the need for that, so I call it non-parametric estimator of complexity.
-Only requiring that a process can be repeated on other data, drastically expands the set of inductive processes for which we can estimate the inductive bias.
+Only requiring that a process can be repeated on other data drastically expands the set of inductive processes for which we can estimate the inductive bias.
 
@@
 \begin{equation}
-\mathcal{L}(\text{Theory}, \text{Reality})
+\mathcal{L}(\text{Theory}, \text{Reality})\,,
+\end{equation}
+where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
+To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory:
+\begin{equation}
+\mathcal{L}(\text{Model(Reality)}, \text{Reality})\,.
 \end{equation}
 
-where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
-To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.
-
-\begin{equation}
-\mathcal{L}(\text{Model(Reality)}, \text{Reality})
-\end{equation}
 The idea is that the version of our theory gets chosen that best fits reality.
@@
 Choosing the best version of the theory based on a sample is almost surely suboptimal.
-Therefore, the observed loss is an overconfident estimate of the loss in the future and closeness to the truth.
+Therefore, the observed loss is an overconfident estimate of the loss in the future and closeness to the truth:
 \begin{equation}
-\mathcal{L}(\text{Model(Reality)}, \text{Reality}) > \mathcal{L}(\text{Model(Sample)}, \text{Sample})
+\mathcal{L}(\text{Model(Reality)}, \text{Reality}) > \mathcal{L}(\text{Model(Sample)}, \text{Sample})\,.
 \end{equation}
 
-The observed loss is thus an underestimation that has to be corrected for by considering the complexity of the model.
+The observed loss is thus an underestimation that has to be corrected for by considering the complexity of the model:
 \begin{equation}
-\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}(\text{Model})
+\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}(\text{Model})\,.
 \end{equation}
@@
 However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and suggest two starting points for making this bias transparent.
-The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
+The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria).
+The second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
 Both approaches require researchers to make the inductive process transparent rather than merely publishing the results.
@@
 
-As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
-Formal discription, in the strict sense, means that the complexity can be mathematically derived, and in a looser sense, that the process can be repeated at will.
+As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a second part that can only be evaluated subjectively.
+Formal description, in the strict sense, means that the complexity can be mathematically derived.
+In a looser sense, that the process can be repeated at will.
 If a researcher employs a linear model, the complexity is calculable parametrically.
@@
 The researcher cannot be asked to repeat the process on all possible data sets, nor is their behavior mathematically well-defined.
-However, it is known that the resulting linear model is more complex than a linear regression formally implies.
-
+However, it is clear that the resulting linear model is more complex than a linear regression formally implies:
 $$
-\mathcal{C}(\text{Model}) = \mathcal{C}(\text{Model}_\text{formal}) + \mathcal{C}(\text{Model}_\text{informal})
+\mathcal{C}(\text{Model}) = \mathcal{C}(\text{Model}_\text{formal}) + \mathcal{C}(\text{Model}_\text{informal})\,.
 $$
@@
 In the following, I propose preregistration as a means to move induction into the formal domain (pushing the first boundary) and computational reproducibility to make formal induction transparent (pushing the second boundary).
-I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.
+I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, although they may have secondary benefits for science.
 
-## Transparency about statistical models: Computational\ Reproducibility
+## Transparency about statistical models: Computational\ reproducibility
 
@@
 
-While this workflow provides transparency and facilitates judgement of the inductive process by recipients of the work (readers, collaborators, editors, etc.), it asks a lot of the researchers who creating it.
+While this workflow provides transparency and facilitates judgement of the inductive process by recipients of the work (readers, collaborators, editors, etc.), it asks a lot of the researchers who create it.
 To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
@@
 Third, the version of the author's written code (and possibly text) must be tracked using version control systems [@chaconProGit2014; @vanlissaWORCSWorkflowOpen2021].
-Fourth, how the data relates to software and computing infrastructure must be managed by workflow automation, for example [@feldmanMakeProgramMaintaining1979; @kimImplementingGitHubActions2022; @kinsmanHowSoftwareDevelopers2021].
+Fourth, how the data relates to software and computing infrastructure must be managed by workflow automation [e.g., @feldmanMakeProgramMaintaining1979; @kimImplementingGitHubActions2022; @kinsmanHowSoftwareDevelopers2021].
 
@@
 Therefore, it is prudent to replace informal induction by formal induction wherever possible.
-Nonetheless, even if a researcher has made every effort only to employ formal inductive reasoning in a reproducible manner, they must still persuade their readers that they have not engaged in informal induction.
+Nonetheless, even if a researcher has made every effort to only employ formal inductive reasoning in a reproducible manner, they must still persuade their readers that they have not engaged in informal induction.
 
@@
 This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge its complexity.
-For researchers to judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data looked different.
+For researchers to judge the complexity, the process must be transparent to them, i.e., how would the results change if the data looked different.
 
@@
 They have to reason about all possible data patterns they might encounter.
-While this makes registration so difficult, it is exactly the information we need and from which the complexity of the inductive process might be derived.
+While this makes registration so difficult, it is exactly the information we need and from which the complexity of the inductive process may be derived.
 
-Note that requiring induction to be provably formal (analytically analyzable or repeatable), does not imply that induction itself is limited in its extent.
+Note that requiring induction to be provably formal (i.e., analytically analyzable or repeatable), does not imply that induction itself is limited in its extent.
 In other words, researchers may still be very unsure how the data will look and therefore rely on induction to fill in the blanks.
@@
 Or suppose it is unclear what functional form to expect between predictor and outcome.
-In that case, many statistical models exist that allow almost arbitrary functional forms [@rissanenUniversalCodingInformation1984], e.g., smoothing splines [@cravenSmoothingNoisyData1978], random forests [@hoRandomSubspaceMethod1998], neural networks [@amariUniversalTheoremLearning1993].
+In that case, many statistical models exist that allow almost arbitrary functional forms [@rissanenUniversalCodingInformation1984], e.g., smoothing splines [@cravenSmoothingNoisyData1978], random forests [@hoRandomSubspaceMethod1998], or neural networks [@amariUniversalTheoremLearning1993].
 If the theory does not suffice to define an outlier, this decision can be made depending on the data, etc.
@@
 Is the application of mathematically or computationally convenient methods, as is usual in statistical modeling, sufficient?
-Does the arbitrariness of the selection process needs to be accounted for?
+Does the arbitrariness of the selection process need to be accounted for?
 
@@
 I have dealt with arbitrary assumptions by choosing an inductive process and forming the expectation over all possible data.
-A possible direction for future research is to explore wether forming the expectation over inductive processes is similarly fruitful.
+A possible direction for future research is to explore whether forming the expectation over inductive processes is similarly fruitful.
 One suggestion in such a direction is a method called multiverse analysis [@steegenIncreasingTransparencyMultiverse2016], where many different data analytic decisions are systematically explored.
@@
 
-## Future Research
+## Future research
 
@@
 However, I hope we might abandon the tradition of only analyzing scientific practices long after they have been established.
-Instead the aim should be to analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
+Instead, the aim should be to analyze new practices of science, such as computational reproducibility or preregistration, as they emerge.
 That way, theoretical insight may shape application.
@@
 
-## Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.
+## Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization
 
@@
 
+Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization*. PsyArXiv. https://doi.org/10.31234/osf.io/cs8wb
+
 \cleardoublepage

### 2023-04-04 presentation.Rmd
update date
update link
add first version of talk slides

@@
   - "Aaron Peikert"
-date: '`r Sys.Date()`'
+date: 'Defense | 22.05.2023'
 output:
   xaringan::moon_reader:
-    seal: false
-    css: xaringan-themer.css
+    css:
+    - xaringan-themer.css
+    - "https://cdnjs.cloudflare.com/ajax/libs/animate.css/4.1.1/animate.min.css"
     nature:
-      slideNumberFormat: "%current%"
-      highlightStyle: github
+      slideNumberFormat: |
+        <div class="progress-bar-container">
+          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
+          </div>
+        </div>
+      highlightStyle: solarized-dark
       highlightLines: true
       ratio: 16:9
-      countIncrementalSlides: true
+      countIncrementalSlides: false
+    seal: true
 repro:
@@
 class: center, middle
-name: qrcode
+
+<style>
+.remark-slide-number {
+  position: inherit;
+}
+
+.remark-slide-number .progress-bar-container {
+  position: absolute;
+  bottom: 0;
+  height: 2px;
+  display: block;
+  left: 0;
+  right: 0;
+}
+
+.remark-slide-number .progress-bar {
+  height: 100%;
+  background-color: #b58900;
+}
+
+@keyframes khighlight {
+  from {color: inherit;}
+  to {color: var(--red);}
+}
+
+@keyframes ksuppress {
+  from {color: inherit;}
+  to {color: var(--grey);}
+}
+
+@keyframes kdehighlight {
+  from {color: var(--red);}
+  to {color: var(--textgrey);;}
+}
+
+/* The element to apply the animation to */
+.highlight {
+  animation-name: khighlight;
+  animation-duration: 1s;
+  animation-fill-mode: forwards;
+}
+
+.suppress {
+  animation-name: ksuppress;
+  animation-duration: 1s;
+  animation-fill-mode: forwards;
+}
+
+.big-borders {
+  margin-left: 150px;
+  margin-right: 150px;
+}
+
+ul {
+  list-style: none;
+}
+
+li::before {
+  content: "⚬ ";
+  color: var(--red);
+}
+
+.transparent {
+  opacity: 0
+}
+
+</style>
 
@@
 ```{r xaringan-themer, include=FALSE, warning=FALSE}
-primary_color <- "#024959"
-secondary_color <- "#F28705"
-style_duo(
-  primary_color = primary_color,
-  secondary_color = secondary_color,
-  header_font_google = google_font("Didact Gothic", "400", "400i"),
-  text_font_google   = google_font("IBM Plex Sans", "400", "400i"),
-  text_font_size = "1.5rem",
-  extra_fonts = list(google_font("Parisienne", "400", "400i"))
+colors = list(
+  red = "#dc322f",
+  white = "#FFFFFF",
+  black = "#002b36",
+  grey = "#93a1a1",
+  textgrey = "#657b83",
+  yellow = "#b58900"
+)
+
+style_solarized_light(
+  text_font_family = "LibertinusSansRegular",
+  text_font_url = "https://fontlibrary.org//face/libertinus-sans",
+  header_font_family = "LibertinusSerifDisplayRegular",
+  header_font_url = "https://fontlibrary.org//face/libertinus-serif",
+  text_font_size = "1.2rem",
+  text_color = "#657b83",
+  text_bold_color = "#b58900",
+  text_slide_number_color = "#657b83",
+  colors = unlist(colors)
 )
@@
                            ".large" = list(`font-size` = "130%"),
+                           ".huge" = list(`font-size` = "170%"),
+                           ".Huge" = list(`font-size` = "250%"),
+                           ".HUGE" = list(`font-size` = "600%"),
                            ".small" = list(`font-size` = "80%"),
@@
                                              `-webkit-filter` = "blur(2px)"),
-                           ".curly" = list("font-family" = "Parisienne"),
-                           ".white" = list(color = "#FFFFFF")
-                    )
-                )
+                           ".curly" = list("font-family" = "Parisienne")))
 ```
 
-```{r, echo=FALSE}
-link <- "github.com/aaronpeikert/bayes-prereg"
-```
+# `r rmarkdown::metadata$title`
 
-```{r, echo=FALSE, out.width = "30%"}
-fs::dir_create("images")
-generate_svg(qr_code(link), here::here("images/", "qr_slides.svg"), foreground = secondary_color, background = primary_color, show = FALSE)
-knitr::include_graphics("images/qr_slides.svg")
-```
+## `r rmarkdown::metadata$subtitle`
 
-`r link`
+### `r rmarkdown::metadata$author`
 
-.pull-right[![CC0](https://mirrors.creativecommons.org/presskit/buttons/88x31/svg/cc-zero.svg)]
-
----
-class: center, middle
-
-# Transparency in Science
-
-## Principled Approaches to Computational Reproducibility and Preregistration
-
-### Aaron Peikert
-
-### LCBC Berlin  | 22.11.2022
+### `r rmarkdown::metadata$date`
 
@@
 ---
+class: middle
 
@@
   geom_col() +
-  geom_text(position = position_stack(vjust = 0.5)) +
+  geom_text(position = position_stack(vjust = 0.5), color = "#fdf6e3") +
   coord_flip() +
@@
   scale_x_discrete(expand=c(0,0), limits = NULL) +
-  scale_fill_manual(values = c("YES" = "#E3350E", "NO" = "#F28705", "NA" = "lightgrey")) +
+  scale_fill_manual(values = c("YES" = colors$red, "NO" = colors$black, "NA" = colors$grey)) +
   scale_size_continuous(range = c(4, 20)) +
@@
 
-.small.right[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]
+.small.right.animate__animated.animate__fadeIn.animate__slower[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]
 
---
-
-.small.right[***They call it "reproducibility" and fail to publish the raw data.**]
-
----
-class: inverse, center, middle
-
-# Transparency in Science*
-
-*Statisticians' Edition
+.small.right.animate__animated.animate__fadeIn.animate__slower[***They call it "reproducibility" and fail to publish the raw data.**]
 
@@
 
-# World → Theory
+#.textgrey[Psychology is a difficult science]
+
+.small.right.animate__animated.animate__fadeIn.animate__slower.animate__delay-2s[Meehl, P.
+E.
+(1978).
+Theoretical risks and tabular asterisks: Sir Karl, Sir Ronald, and the slow progress of soft psychology.
+*Journal of Consulting and Clinical Psychology*, 46(4), 806–834. https://doi.org/10.1037/0022-006X.46.4.806]
 
@@
 
-# World ← Theory
+.huge[Meehl, P. E. (1978). Theoretical risks and tabular asterisks:  
+Sir Karl, Sir Ronald, and the slow progress of soft psychology.  
+*Journal of Consulting and Clinical Psychology*, 46(4), 806–834.]
 
@@
 
-# World ↔ Theory
-
---
-
-.curly.white.large[It's a match!]
+.huge.supress[Meehl, P. E. (1978). Theoretical risks and tabular asterisks:  
+Sir Karl, Sir Ronald, and the .highlight[slow progress] of .highlight[soft psychology].  
+*Journal of Consulting and Clinical Psychology*, 46(4), 806–834.]
 
 ---
-class: center, middle, inverse
+class:  middle, inverse, center
 
-# Principled Approach
+# The Progress of Soft Theories
 
@@
 
-# $L(\text{Theory}, \text{Data})$
+# Theories
+
+.huge.animate__animated.animate__fadeIn.animate__slower[= allow to deduce predictions
+]
 
@@
 
-# ѴαЯi€𝜏y
+#.textgrey[Hard Theories]
 
-### $L(\text{Theory}, \text{Data})$
+.huge[= make .highlight[precise] predictions for given circumstances
+]
 
@@
 
-$$L_1(\text{Theory}, \text{Data})$$
-$$\vdots$$
-$$L_\infty(\text{Theory}, \text{Data})$$
+#.textgrey[Soft Theories]
 
----
-class: center, middle, inverse
-
-# Statistical Models
-
---
-
-= automated induction
-
----
-class: center, middle
-
-# Theory
-
----
-class: center, middle
-
-# T█e█ry
+.huge[= make .highlight[many] predictions for given circumstances
+]
 
@@
 
-$$Model(\text{Theory})$$
----
-class: middle
+.huge[
 
-$$Model(\text{Theory}, \text{Data})$$
+We allow predictions to be:
 
----
-
-### How statisticians judge a model:
-
-Sampling the world and compare:
-
-$$
-L(\text{Model}(\text{Data}), \text{Data}) + \mathcal{C}(\text{Model})
-$$
-e.g.: Adjusted R², Stein's Unbiased Risk Estimator, Information Criteria, etc.
-
-.center[
-
-**We account for peeking at the data**
-
-]
-
----
-
-### Researcher have a harder job:
-
-Sampling the world and compare:
-
-$$
-L(\text{Model}(\text{Data}), \text{Data}) + \mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})
-$$
-.center[
-
-**We must account for the models and humans ability to make sense of any data.**
+- close to truth
+- probabilistic
+- parameterized
 
@@
 
-.left[What happens if]
+#.textgrey[Verisimilitude]
 
-.center[
+(=closeness to truth)
 
-$\mathcal{C}(\text{Model})$ or $\mathcal{C}(\text{Human})$
+.huge[
+
+$\mathcal{L}(\text{Theory}, \text{Data})$
 
@@
 
-.right[is unknown?]
+---
+class: center, middle
 
----
-class: center, middle, inverse
+#.textgrey[Randomness]
 
-# .blurry[Uncertainty]
+(=probabilistic)
 
----
+.huge[
 
-The goal:
-
-.center[
-
-$L(\text{Model}(\text{Data}), \text{Data}) + \mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$
+$\mathcal{L}(\text{Theory}, \text{Data})$ is just an estimate.
 
@@
 
+---
+class: center, middle
+
+#.textgrey[Induction]
+
+(=parameterized)
+
+.huge[
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+]
+
+---
+class: middle, center, inverse
+
+# Induction
+
+.transparent[(=parameterized)]
+
+.huge.transparent[
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+]
+
+---
+class: center, inverse
+
+.huge[
+
+$\text{Model}(\text{Theory}, \text{Data})$
+
+]
+
+Find the version of a theory, that is best supported by data.
+
 --
 
-The enemy:
+or
+
+Of any prediction that a theory can make, choose the one that minimizes $\mathcal{L}(\cdot, \text{Data})$.
+
+--
+
+or
+
+```julia
+for prediction in theory
+  current = loss(prediction, data)
+  if current < best
+    best = current
+  end
+  return best
+end
+```
+
+---
+class: center, middle
+
+.big-borders[
+
+##.left.red[Induction]
+
+###.textgrey[enables]
+
+##.right.red[soft theories]
+
+]
+
+---
+class: center, middle
+
+.Huge[One]   .red[little]   .Huge[problem.]
+
+---
+class: center, middle
+
+###.textgrey[Verisimilitude × Randomness × Induction]
+
+---
+class: center, middle
+
+# Bias
+
+---
+class: center, middle
+
+Psychology is difficult .transparent[→ soft theories → require induction → biased expectation → replication crisis]
+
+---
+class: center, middle
+
+Psychology is difficult .animate__animated.animate__fadeIn.red[→ soft theories] .transparent[→ require induction → biased expectation → replication crisis]
+
+---
+class: center, middle
+
+Psychology is difficult .dehighlight[→ soft theories] .animate__animated.animate__fadeIn.red[→ require induction] .transparent[→ biased expectation → replication crisis]
+
+---
+class: center, middle
+
+Psychology is difficult → soft theories .dehighlight[→ require induction] .animate__animated.animate__fadeIn.red[→ biased expectation] .transparent[→ replication crisis]
+
+---
+class: center, middle
+
+Psychology is difficult → soft theories → require induction .dehighlight[→ biased expectation] .animate__animated.animate__fadeIn.red[→ replication crisis]
+
+---
+class: center, middle
+
+#.textgrey[
+
+Why is .highlight[bias] a problem?
+
+]
+
+---
+class: center, middle
+
+#.blurry[Uncertainty]
+
+---
+class: center, middle
+
+If the inductive bias is uncertain evaluating,
+
+.huge[
+
+$\mathcal{L}(\text{Theory}, \text{Data})$
+
+]
+
+becomes uncertain and unconvincing.
+
+---
+class: center, middle
+
+$$\mathcal{L}(\text{Theory}, \text{Data})$$
+
+$$=$$
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+$+$
+
+.highlight[
+
+$\mathcal{C}(\text{Model})$
+
+]
+
+.small.right.transparent[
+
+This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.
+
+]
+
+---
+class: center, middle
+
+$$\mathcal{L}(\text{Theory}, \text{Data})$$
+
+$$=$$
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+$+$
+
+.red[
+
+$\mathcal{C}(\text{Model})$
+
+]
+
+.small.right[
+
+This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.
+
+]
+
+---
+class: center, middle
+
+.large.transparent[Under some circumstances,]
+
+.huge.red[
+
+$\mathcal{C}(\text{Model})$
+
+]
+
+.large[is the ability of a model to accommodate random data.]
+
+---
+class: center, middle
+
+.large[Under some circumstances,]
+
+.huge.red[
+
+$\mathcal{C}(\text{Model})$
+
+]
+
+.large[can be approximated.]
+
+---
+class: center, middle
+
+.large[ .highlight[Reproducibility] is a necessary* condition to estimate:]
+
+.huge[
+$\mathcal{C}(\text{Model})$]
+
+.transparent.small.right.grey[*In an ideal world.]
+
+---
+class: center, middle
+
+.large[
+
+Some things are
+
+.huge[.highlight[non]-reproducible]
+
+on a fundamental level.
+
+]
+
+---
+class: center, middle
+
+.huge.red[
+
+$\mathcal{C}(\text{Human})$
+
+]
+
+---
+class: middle, center
 
@@
 
-.blurry2[$$\mathcal{C}(\text{Model})$$]
+###.center[Reproducibility]
+
+Allows estimating:
+
+.large[
+
+$$\mathcal{C}(\text{Statistical Model})$$
+
+]
 
@@
 
-.blurry2[$$\mathcal{C}(\text{Human})$$]
+###.center[Preregistration]
+
+Does reduce:
+
+.large[
+
+$$\mathcal{C}(\text{Human})$$
 
@@
 
---
+]
+---
+class: center, middle
 
-The solution:
+.large[.highlight[Reproducibility] is necessary to estimate:]
 
-.pull-left.center[
+.large[
 
-### Computational Reproducibility
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model})$ .transparent[ $\require{cancel} + \cancel{\mathcal{C}(\text{Human})}$
+]
 
@@
 
-.pull-right.center[
-
-### Preregistration
-
-]
-
----
-
-.pull-left.center[
-
-### Computational Reproducibility
-
-Conceptually trivial
-
-]
-
-.pull-right.center[
-
-### Preregistration
-
-Conceptually difficult
-
-]
-
---
-
-.pull-left.center[
-
-technologically difficult
-
-low social hurdles
-
-]
-
-.pull-right.center[
-
-technologically trivial
-
-high social hurdles
-
-]
-
----
-
-.pull-left.center[
-
-### Computational Reproducibility
-
-Conceptually trivial 😀
-
-technologically difficult 😒
-
-low social hurdles 😀
-
-]
-
-.pull-right.center[
-
-### Preregistration
-
-Conceptually difficult 😒
-
-technologically trivial 😀
-
-high social hurdles 😒
-
-]
-
---
-
-Reducing uncertainty about:
-
-.center[
-
-$\mathcal{C}(\text{Model})$ &emsp; vs &emsp; $\mathcal{C}(\text{Human})$
-
-]
-
----
-class: center, middle, inverse
-
-# Computational Reproducibility
+.small.right.transparent[*In an ideal world.]
 
@@
 
-“Insanity is doing the same thing over and over again and expecting different results.”
+.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]
 
-.right[—Albert Einstein (disputed)]
+.large[
 
---
+$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$
 
-As it turns out, doing the **same** thing is pretty complicated.
+]
+
+.small.right.grey[*In an ideal world.]
+
+---
+class: center, middle
+
+.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]
+
+.large[
+
+$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \cancel{\mathcal{C}(\text{Human})}$
+
+]
+
+.small.right.grey[*In an ideal world.]
+---
+background-image: url("images/contribution1.svg")
+background-position: center
+background-size: contain
+---
+background-image: url("images/contribution2.svg")
+background-position: center
+background-size: contain
+---
+background-image: url("images/contribution3.svg")
+background-position: center
+background-size: contain
+---
+background-image: url("images/contribution4.svg")
+background-position: center
+background-size: contain
+---
+class: middle
+
+**.red[1\. Paper:]**  
+Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow. *Quantitative and Computational Methods in Behavioral Sciences*, *1*. https://doi.org/10.5964/qcmb.3763
+
+**2\. Paper:**  
+Peikert, A., van Lissa, C.
+J., & Brandmaier, A.
+M.
+(2021).
+Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once.
+*Psych*, *3*(4), 836–867.
+https://doi.org/10.3390/psych3040053
+
+**3\. Paper:**  
+Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.* PsyArXiv. .textgrey[https://doi.org/10.31234/osf.io/cs8wb]
+
+**Synopsis:**  
+Peikert, A.
+(2023).
+*Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration* [Doctoral dissertation, Humboldt-Universität zu Berlin].
+Zenodo.
+https://www.doi.org/10.5281/zenodo.7654989
+
+---
+class: inverse, middle, center
+
+# Automate Reproducibility
+
+---
+class:  middle, center
+
+# Four Problems
+
+---
+class:  middle, center
+
+# Four Concepts
+
+---
+class:  middle, center
+
+# Four Tools
 
@@
 
-1. Multiple inconsistent versions of code and data
---
-
-2. Copy-and-paste errors
---
-
-3. Ambiguous order of code execution
---
-
-4. Broken dependencies
-
----
-class: center, middle
-
-| Problem       | Solution            | Implementation    |
-|---------------|---------------------|-------------------|
-| versions?     | version control     | Git               |
-| copy&paste?   | dynamic documents   | R Markdown        |
-| order?        | workflow automation | Make / GH Actions |
-| dependencies? | software management | Docker            |
-
----
-class: center, middle
-background-color: white
-
-```{r, out.width='80%'}
-knitr::include_graphics("https://raw.githubusercontent.com/aaronpeikert/repro-tutorial/main/images/nutshell.svg")
-```
-
----
-class: center, middle
-
-.small[<https://github.com/aaronpeikert/repro-talk/releases/download/marbach/repro-talk.pdf>]
-
----
-class: center, middle, inverse
-
-# Preregistration
-
----
-
-## Is there a replication crisis?*
-
-.center[
-
-```{r, out.width = '70%'}
-crisis_plot /
-  (crisis_plot +
-     aes(label = str_c(prop * 100, "%")) + 
-     scale_size_continuous(range = c(3, 12))) + 
-  plot_layout(heights = c(.7, .3)) + 
-  plot_annotation(theme = theme_void() + theme(
-    panel.background = element_rect(fill = "transparent", colour = NA),
-    plot.background = element_rect(fill = "transparent", colour = NA)
-))
-```
+.pull-left.large[
+.textgrey[1\. Copy&Paste Mistakes]  
+.transparent[2\. Conflicting Versions]  
+.transparent[3\. Missing Software]  
+.transparent[4\. Unclear Procedure]  
 ]
 
-.small.right[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]
-
----
-class: center, middle
-layout: true
-
----
-
-## Because researchers are *fraudsters*?
-
---
-
-# No
-
-???
-Some say to prevent intentional wrongdoing. 
-However, preregistrations are easily circumvented with bad intentions.
-We must assume researchers have good intentions.
----
-
-## Because researchers are *fools*?
-
---
-
-# No
---
-, not necessarily
-
-???
-Some say to prevent unintentional wrongdoing.
-Researchers must be put in restrains to prevent them from cutting themselves on the sharp edge of "false positives".
-
----
-layout: false
-class: inverse, center, middle
-
-# Traditional view
-
----
-class: center
-
-# Researchers confuse:
-
-.pull-left.center[
-
-# confirmatory
-
-]
-.pull-right.center[
-
-# exploratory
-
+.pull-right.large[
+.transparent[Dynamic Documents]  
+.transparent[Version Control]  
+.transparent[Software Management]  
+.transparent[Workflow Orchestration]  
 ]
 ---
-class: center
+class: middle
 
-# Preregistration separates:
-
-.pull-left.center[
-
-# confirmatory
-
-]
-.pull-right.center[
-
-# exploratory
-
-]
---
-.pull-left.center[
-
-# ≈
-# preregistered
-
-]
-.pull-right.center[
-
-# ≈
-# not preregistered
-
+.pull-left.large[
+.textgrey[1\. Copy&Paste Mistakes]  
+.transparent[2\. Conflicting Versions]  
+.transparent[3\. Missing Software]  
+.transparent[4\. Unclear Procedure]  
 ]
 
---
+.pull-right.large[
+.highlight[Dynamic Documents]  
+.transparent[Version Control]  
+.transparent[Software Management]  
+.transparent[Workflow Orchestration]  
+]
+---
+class: middle
 
-# ?
+.pull-left.large[
+.suppress[1\. Copy&Paste Mistakes]  
+.textgrey[2\. Conflicting Versions]  
+.transparent[3\. Missing Software]  
+.transparent[4\. Unclear Procedure]  
+]
 
+.pull-right.large[
+.suppress[Dynamic Documents]  
+.transparent[Version Control]  
+.transparent[Software Management]  
+.transparent[Workflow Orchestration]  
+]
 ---
-class: inverse, center, middle
+class: middle
 
-# A Formal Objective
+.pull-left.large[
+.grey[1\. Copy&Paste Mistakes]  
+.textgrey[2\. Conflicting Versions]  
+.transparent[3\. Missing Software]  
+.transparent[4\. Unclear Procedure]  
+]
+
+.pull-right.large[
+.grey[Dynamic Documents]  
+.highlight[Version Control]  
+.transparent[Software Management]  
+.transparent[Workflow Orchestration]  
+]
+---
+class: middle
+
+.pull-left.large[
+.grey[1\. Copy&Paste Mistakes]  
+.suppress[2\. Conflicting Versions]  
+.textgrey[3\. Missing Software]  
+.transparent[4\. Unclear Procedure]  
+]
+
+.pull-right.large[
+.grey[Dynamic Documents]  
+.suppress[Version Control]  
+.transparent[Software Management]  
+.transparent[Workflow Orchestration]  
+]
+---
+class: middle
+
+.pull-left.large[
+.grey[1\. Copy&Paste Mistakes]  
+.grey[2\. Conflicting Versions]  
+.textgrey[3\. Missing Software]  
+.transparent[4\. Unclear Procedure]  
+]
+
+.pull-right.large[
+.grey[Dynamic Documents]  
+.grey[Version Control]  
+.highlight[Software Management]  
+.transparent[Workflow Orchestration]  
+]
+---
+class: middle
+
+.pull-left.large[
+.grey[1\. Copy&Paste Mistakes]  
+.grey[2\. Conflicting Versions]  
+.suppress[3\. Missing Software]  
+.textgrey[4\. Unclear Procedure]  
+]
+
+.pull-right.large[
+.grey[Dynamic Documents]  
+.grey[Version Control]  
+.suppress[Software Management]  
+.transparent[Workflow Orchestration]  
+]
+---
+class: middle
+
+.pull-left.large[
+.grey[1\. Copy&Paste Mistakes]  
+.grey[2\. Conflicting Versions]  
+.grey[3\. Missing Software]  
+.textgrey[4\. Unclear Procedure]  
+]
+
+.pull-right.large[
+.grey[Dynamic Documents]  
+.grey[Version Control]  
+.grey[Software Management]  
+.highlight[Workflow Orchestration]  
+]
 
@@
 
-.large[
-Ideally:
-1. The original author has **good intentions**.
-2. Their readers are **rational**.
-3. The goal is to **convince** the readers.
-]
+**1\. Paper:**  
+Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow. *Quantitative and Computational Methods in Behavioral Sciences*, *1*. https://doi.org/10.5964/qcmb.3763
+
+**.red[2\. Paper:]**  
+Peikert, A., van Lissa, C.
+J., & Brandmaier, A.
+M.
+(2021).
+Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once.
+*Psych*, *3*(4), 836–867.
+https://doi.org/10.3390/psych3040053
+
+**3\. Paper:**  
+Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.* PsyArXiv. .textgrey[https://doi.org/10.31234/osf.io/cs8wb]
+
+**Synopsis:**  
+Peikert, A.
+(2023).
+*Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration* [Doctoral dissertation, Humboldt-Universität zu Berlin].
+Zenodo.
+https://www.doi.org/10.5281/zenodo.7654989
 
 ---
+class: inverse, middle, center
 
-### Adjusting for complexity in simple models
-
-Adjusted R², Stein's Unbiased Risk Estimator, Information Criteria, etc.
-
-$$L(\text{Model}(\text{Data}), \text{Data})  + \mathcal{C}(\text{Model})$$
-
---
-
-### Alternatively:
-
-The statistical complexity of some models is hard to estimate.
-
-Find a situation, were $\mathcal{C}(\text{Model}) = 0$
-
----
-## Test-set (cross validation)
-
-$\mathcal{C}(\text{Model}) = 0$
-
-$$
-L(\text{Model}(\text{Train, Test}), \text{Test})
-$$
-.center[
-
-**There is no peeking to account for.**
-
-]
-
----
-## $\mathcal{C}(\text{Human})$
-
-The statistical complexity of humans is hard to estimate.
-
-Find a situation, were $\mathcal{C}(\text{Human}) = 0$
-
---
-
-.center[
-
-# Preregistration
-
-]
-
----
-class: center, middle, inverse
-
-# .blurry[Uncertainty]
-
----
-# Statistical Relevancy
-
-Any:
-
-$$L(\text{Theory}, \text{Data})$$
-
-Must benefit from:
-
-$$P(E|H) \gg P(E|¬H)$$
-
----
-class: center, middle, inverse
-
-# Bayes*
-
-*my favoritey $L(\text{Theory}, \text{Data})$
-
----
-class: middle
-
-.large[
-
-$$\frac{P(H)P(E|H)}{P(E)}$$
-]
-
---
-
-.left[Law of total probability:]
-
-$$P(E) = P(H)P(E|H) + P(¬H)\color{#F28705}{P(E|¬H)}$$
+# Reproducibility × Preregistration
 
@@
 class: center, middle
-
-## $1 - P(E|¬H)$
-## =
-## $P(¬E|¬H)$
-## =
-## Theorethical Risk
-
----
-class: center, middle
-
-## $\mathcal{C}(\text{Model})$ → Theorethical Risk ← $\mathcal{C}(\text{Human})$
-
----
-class: center, middle, inverse
-
-# .blurry[Uncertainty]
-
----
-class: center, middle
-name: theoretical-risk-plot2
-
-```{r}
-results <- readr::read_rds(here::here("data", "simulation.rds")) 
-```
-
-```{r measure-plots}
-# see R/plots.R
-```
-
-```{r}
-plots[["Posterior Probability"]] +
-  xaringanthemer::theme_xaringan() +
-    theme(legend.position = c(.2, .8),
-        legend.title = element_blank(),
-        text = element_text(size = 18, color = secondary_color),
-        title = element_blank(),
-        axis.title = element_text(size = 18),
-        plot.caption = element_text(size = 15)) +
-  labs(x = "Theoretical Risk",
-       y = "Posterior Probability",
-       caption = "P(E|H) = .8, P(H) = .1") +
-  NULL
-```
-
----
-class: center, middle
-layout: true
----
-
-.large[**Preregistrations**] reduce .large[**uncertainty**] about .large[**theoretical risk**].
-
----
-
-### One further idea
-
----
-
-# Preregistration as Code
-
-### version control + dynamic documents
-
----
-class: center
 
@@
 article draft
+
 ]
+
+--
 
@@
 ]
+---
+class: middle
+
+**1\. Paper:**  
+Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow. *Quantitative and Computational Methods in Behavioral Sciences*, *1*. https://doi.org/10.5964/qcmb.3763
+
+**2\. Paper:**  
+Peikert, A., van Lissa, C.
+J., & Brandmaier, A.
+M.
+(2021).
+Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once.
+*Psych*, *3*(4), 836–867.
+https://doi.org/10.3390/psych3040053
+
+**.red[3\. Paper:]**  
+Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.* PsyArXiv. .textgrey[https://doi.org/10.31234/osf.io/cs8wb]
+
+**Synopsis:**  
+Peikert, A.
+(2023).
+*Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration* [Doctoral dissertation, Humboldt-Universität zu Berlin].
+Zenodo.
+https://www.doi.org/10.5281/zenodo.7654989
 
 ---
+class: inverse, middle, center
 
-# Can I explore the data?
-
---
-
-# Yes.
+# The Goal of Preregistration
 
 ---
-template: theoretical-risk-plot2
+class: middle, center
+
+#.blurry2[Uncertainty]
 
 ---
-
-# Can I deviate?
-
---
-
-# Sometimes.
-
+background-image: url("images/contribution.svg")
+background-position: center
+background-size: contain
 ---
-template: theoretical-risk-plot2
-
----
-
-# How much detail?
-
---
-
-# Very detailed.
-
----
-template: theoretical-risk-plot2
-
----
-template: qrcode

### 2023-05-15 presentation.Rmd
add more slides
add points about ML roughly equal to preregistration

@@
 
+.large[
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+]
+---
+class: center, middle
+
 Psychology is difficult .transparent[→ soft theories → require induction → biased expectation → replication crisis]
@@
 
-$\mathcal{L}(\text{Theory}, \text{Data})$
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
 
@@
 
-.large[ .highlight[Reproducibility] is a necessary* condition to estimate:]
+.large[ .highlight[Reproducibility] is a necessary condition to estimate:]
 
@@
 
-.transparent.small.right.grey[*In an ideal world.]
+.transparent[*In an ideal world.]
 
@@
 
+.transparent.blurry2.large[about]
+
+.transparent.huge.blurry2[
+
+$\mathcal{C}(\text{Human})$
+
+]
+
+---
+class: middle, center
+
+#.suppress.blurry2[Uncertainty]
+
+.blurry2.large[about]
+
+.highlight.huge.blurry2[
+
+$\mathcal{C}(\text{Human})$
+
+]
+
+---
+class: center, middle
+
+.large[
+In .highlight[Machine Learning], $\mathcal{C}(\text{Model})$ is often unknown.
+]
+
+---
+class: center, middle
+
+#.textgrey[Cross Validaton]
+
+.right.large.transparent[**but resetting researchers memory is unethical]
+
+.right.large.transparent[**impossible**]
+
+---
+class: center, middle
+
+#.textgrey[Cross Validaton]
+
+.right.large[**but resetting researchers memory is ~~unethical~~**]
+
+.right.large[**impossible**]
+
+---
+class: center, middle
+
+#.textgrey[Test-set]
+
+$\mathcal{C}(\text{Model}) = 0$
+
+---
+class: center, middle
+
+# Preregistration
+
+$\mathcal{C}(\text{Human}) = 0$
+
+---
+class: center, middle
+
+# Unconvential
+
+---
+class: center, middle
+
+# Explatory Preregistrations?
+
+$\mathcal{C} = \mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$
+
+Preregistration: $\mathcal{C}(\text{Human}) = 0$
+
+$\mathcal{C}(\text{Model})$ still might be high.
+
+---
+class: center, middle
+
+```{r}
+results <- readr::read_rds(here::here("data", "simulation.rds")) 
+```
+
+```{r measure-plots}
+# see R/plots.R
+```
+
+```{r}
+p <- plots[["Posterior Probability"]] +
+  xaringanthemer::theme_xaringan() +
+    theme(legend.position = c(.2, .8),
+        legend.title = element_blank(),
+        text = element_text(size = 18, color = colors$red),
+        title = element_blank(),
+        axis.title = element_text(size = 18),
+        plot.caption = element_text(size = 15)) +
+  labs(x = "Exploratory ↔ Confirmatory",
+       y = "Posterior Probability",
+       caption = "P(E|H) = .8, P(H) = .1") +
+  NULL
+p
+```
+
+---
+class: center, middle
+
+```{r}
+p +
+  scale_x_continuous(limits = c(.8, 1)) +
+  NULL
+```
+
 ---

### 2023-05-15 presentation.Rmd
start with a high level summary

@@
 ---
-class: inverse, center, middle
+background-image: url("images/contribution.svg")
+background-position: center
+background-size: contain
+
+---
+class: center, middle, inverse
+
+.large.red[Human]   .Huge.yellow[Bias]   .large.red[Machine]
+
+---
+class: center, middle
+
+.large.red[Preregistration]   .Huge.yellow[Induction]   .large.red[Reproducibility]
+
+---
+class: center, middle
 
@@
 background-size: contain
----

### 2023-05-20 presentation.Rmd
reorder slides

@@
 ---
-background-image: url("images/contribution.svg")
+background-image: url("images/contribution_path.svg")
 background-position: center
@@
 
+$\mathcal{L}(\text{Theory}, \text{Reality})$
+
+]
+
+.transparent[is just an estimate.]
+
+---
+class: center, middle
+
+#.textgrey[Randomness]
+
+(=probabilistic)
+
+.huge[
+
 $\mathcal{L}(\text{Theory}, \text{Data})$
@@
 
----
-class: center, middle
-
-#.textgrey[Randomness]
-
-(=probabilistic)
+is just an estimate.
+
+---
+class: center, middle
+
+#.textgrey[Induction]
+
+(=parameterized)
 
@@
 
-$\mathcal{L}(\text{Theory}, \text{Data})$ is just an estimate.
-
-]
-
----
-class: center, middle
-
-#.textgrey[Induction]
-
-(=parameterized)
-
-.huge[
-
 $\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
@@
 
+.transparent[is just an estimate.]
+
 ---
@@
 ]
+
+is just an estimate.
 
@@
 
+##.right.red[precision]
+
+]
+
+---
+class: center, middle
+
+.big-borders[
+
+##.left.red[Induction]
+
+###.textgrey[enables]
+
+##.right.red[predictions]
+
+]
+
+---
+class: center, middle
+
+.big-borders[
+
+##.left.red[Induction]
+
+###.textgrey[enables]
+
 ##.right.red[soft theories]
@@
 ---
-class: center, middle
-
-# Bias
+class: center, middle, inverse
+
+#.red[Bias]
 
@@
 
+$$\mathcal{L}(\text{Theory}, \text{Data})$$
+
+$$=$$
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+$+$
+
+.highlight[
+
+$\mathcal{C}(\text{Model})$
+
+]
+
+.small.right.transparent[
+
+This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.
+
+]
+
+---
+class: center, middle
+
+$$\mathcal{L}(\text{Theory}, \text{Data})$$
+
+$$=$$
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
+
+$+$
+
+.red[
+
+$\mathcal{C}(\text{Model})$
+
+]
+
+.small.right[
+
+This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.
+
+]
+
+---
+class: center, middle, inverse
+
+.large[.highlight[Reproducibility] is necessary to estimate:]
+
+.large[
+
+$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model})$ .transparent[ $\require{cancel} + \cancel{\mathcal{C}(\text{Human})}$
+]
+
+]
+
+.small.right.transparent[*In an ideal world.]
+
+---
+class: center, middle
+
+#.textgrey[
+
+Why is .highlight[bias] a problem?
+
+]
+
+---
+class: center, middle
+
 #.blurry[Uncertainty]
@@
 
-$$\mathcal{L}(\text{Theory}, \text{Data})$$
-
-$$=$$
-
-$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
-
-$+$
-
-.highlight[
-
-$\mathcal{C}(\text{Model})$
-
-]
-
-.small.right.transparent[
-
-This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.
-
-]
-
----
-class: center, middle
-
-$$\mathcal{L}(\text{Theory}, \text{Data})$$
-
-$$=$$
-
-$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$
-
-$+$
-
-.red[
-
-$\mathcal{C}(\text{Model})$
-
-]
-
-.small.right[
-
-This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.
-
-]
-
----
-class: center, middle
-
-.large.transparent[Under some circumstances,]
+.large[
+
+Some things are
+
+.huge[.highlight[non]-reproducible]
+
+on a fundamental level.
+
+]
+
+---
+class: center, middle
 
@@
 
-$\mathcal{C}(\text{Model})$
-
-]
-
-.large[is the ability of a model to accommodate random data.]
-
----
-class: center, middle
-
-.large[Under some circumstances,]
-
-.huge.red[
-
-$\mathcal{C}(\text{Model})$
-
-]
-
-.large[can be approximated.]
-
----
-class: center, middle
-
-.large[ .highlight[Reproducibility] is a necessary condition to estimate:]
-
-.huge[
-$\mathcal{C}(\text{Model})$]
-
-.transparent[*In an ideal world.]
-
----
-class: center, middle
+$\mathcal{C}(\text{Human})$
+
+]
+
+---
+class: center, middle
+
+.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]
 
@@
 
-Some things are
-
-.huge[.highlight[non]-reproducible]
-
-on a fundamental level.
-
-]
-
----
-class: center, middle
-
-.huge.red[
-
-$\mathcal{C}(\text{Human})$
-
-]
-
----
-class: middle, center
+$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$
+
+]
+
+.small.right.grey[*In an ideal world.]
+
+---
+class: center, middle
+
+.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]
+
+.large[
+
+$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \cancel{\mathcal{C}(\text{Human})}$
+
+]
+
+.small.right.grey[*In an ideal world.]
+
+---
+class: middle, center, inverse
 
@@
 
-$$\mathcal{C}(\text{Statistical Model})$$
+$$\mathcal{C}(\text{Model})$$
 
@@
 ]
----
-class: center, middle
-
-.large[.highlight[Reproducibility] is necessary to estimate:]
-
-.large[
-
-$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model})$ .transparent[ $\require{cancel} + \cancel{\mathcal{C}(\text{Human})}$
-]
-
-]
-
-.small.right.transparent[*In an ideal world.]
-
----
-class: center, middle
-
-.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]
-
-.large[
-
-$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$
-
-]
-
-.small.right.grey[*In an ideal world.]
-
----
-class: center, middle
-
-.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]
-
-.large[
-
-$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \cancel{\mathcal{C}(\text{Human})}$
-
-]
-
-.small.right.grey[*In an ideal world.]
----
-background-image: url("images/contribution1.svg")
+
+---
+background-image: url("images/contribution1_path.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution2.svg")
+background-image: url("images/contribution2_path.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution3.svg")
+background-image: url("images/contribution3_path.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution4.svg")
+background-image: url("images/contribution4_path.svg")
 background-position: center

### 2023-05-20 presentation.Rmd
dont use path versions yet

@@
 ---
-background-image: url("images/contribution_path.svg")
+background-image: url("images/contribution.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution1_path.svg")
+background-image: url("images/contribution1.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution2_path.svg")
+background-image: url("images/contribution2.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution3_path.svg")
+background-image: url("images/contribution3.svg")
 background-position: center
@@
 ---
-background-image: url("images/contribution4_path.svg")
+background-image: url("images/contribution4.svg")
 background-position: center

### 2023-07-21 manuscript.Rmd
a few more punctations in formula
add author

@@
 subtitle: "A Principled Perspective on Computational Reproducibility and Preregistration"
+author: "Aaron Peikert"
 output:
@@
 \begin{equation}
-\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right]
+\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right]\,.
 \end{equation}
@@
 \begin{equation}
-\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))]\label{eq:dataexp}
+\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))]\label{eq:dataexp}\,.
 \end{equation}
@@
 \end{equation}
-
 Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., use cross-validation.

### 2023-07-21 manuscript.Rmd
include pdfx latex package

@@
     - titling
+    - pdfx
   apt:

### 2023-08-08 manuscript.Rmd
use ghostscript to generate PDF/A

@@
 
-\includepdf[pages=-, offset=7mm 0mm,frame=false]{papers/workflow.pdf}
+\includepdf[pages=-, width=\paperwidth - 12mm, offset=12mm 0]{papers/workflow.pdf}
 
@@
 
-\includepdf[pages=-, offset=7mm 0mm]{papers/tutorial.pdf}
+\includepdf[pages=-, width=\paperwidth - 12mm, offset=12mm 0]{papers/tutorial.pdf}
 
@@
 
-\includepdf[pages=-, offset=7mm 0mm]{papers/prereg.pdf}
+\includepdf[pages=-, width=\paperwidth - 12mm, offset=12mm 0]{papers/prereg.pdf}
